\section{Strongly Connected Components Handling}

Steps:
\begin{enumerate}
\item Convert input graph to graph for 2-Dyck querying.
\item Convert graph to one strongly connected component by adding edges with new unique label from 
synks to sources.
\item Convert 2-Dyck grammar to grammar which can accept arbitrary path with 2-Dyck subpaths.
\item Execute modified Rytter for one arbitrary selected vertex and its output edge.
\end{enumerate}

In strongly connected components each vetex is reachcbel from another, but not each path should match 
required constraints.
The idea is to extend grammar by the such way, that it accepts arbitrary path and provide information 
about parts which satisfie to original constraints.
As far as we can reduce any CFPQ to 2-Dyck query, we can fix grammar as follows.
\begin{align*}
S & \rightarrow A \ S_1 \ | \ C \ S_2  \ | \ S \ S \ | \ A \ B \ | \ C \ D
\\
S_1 & \rightarrow S \ B 
\\
S_2 & \rightarrow S \ D 
\\
A & \rightarrow a 
\\
B & \rightarrow b 
\\
C & \rightarrow c 
\\
D & \rightarrow d
\end{align*}

Let label of new edges which added in order to convert graph to single SCC is $E$.
Arbitrary path consists of 2-Dyck subpaths connected by unbalanced parts.
We can specify grammar for these paths.

\begin{align*}
S'  \rightarrow & a \ | \ b \ | \ c \ | \ d \ |  \ e \ |
\\
  & \ A \ S' \ | \ B \ S' \ | \ C \ S' \ | \ D \ S' \ | \ E \ S' \ | \ S' \ S' \ |  
 \\
  & \ A \ S_1 \ | \ C \ S_2 \ | \ S \ S \ | \ A \ B \ | \ C \ D \   
\\
S   \rightarrow & A \ S_1 \ | \ C \ S_2  \ | \ S \ S \ | \ A \ B \ | \ C \ D
\\
S_1  \rightarrow & S \ B 
\\
S_2  \rightarrow & S \ D 
\\
A  \rightarrow & a 
\\
B  \rightarrow & b 
\\
C  \rightarrow & c 
\\
D  \rightarrow & d
\\
E  \rightarrow & e
\end{align*}

Now we can start processing from one arbitrary selected vertex.

Scheme of proof.
\begin{enumerate}
\item Convertion to 2-Dyck path querying. Look at ection !!!. 
\item Convertion to single SCC. In worst case we should add $|V|$ outgoing edges for each vertes. 
So, time complexity is $O(|V|^2)$.
\item Why can we select arbitrary edge for start?
We can select arbitrarry vertex just because we handle SCC, so all other vertices should be reachcble from selected one.
We should choose outgoing edge from selected vertex in order to fix source vertex in Rytter graph.
\item Rytter
\end{enumerate}