\documentclass[12pt]{article}  % standard LaTeX, 12 point type

\usepackage{geometry}

\usepackage{amsmath}
\usepackage{amsfonts,latexsym}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage[utf8]{inputenc} % Кодировка
\usepackage[english,russian]{babel} % Многоязычность
\usepackage{verbatim}
\usepackage{longtable}
\usepackage{csvsimple}
\usepackage[toc,page]{appendix}
\usepackage{booktabs}

\usepackage{float}
\usepackage{array}
\usepackage{multirow}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{ucs}
\usepackage{rotating}
\usepackage{pdflscape}
\usepackage{afterpage}
\usepackage{capt-of}% or use the larger `caption` package
\usepackage{url}

% unnumbered environments:

\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{notation}{Notation}
\newtheorem*{note}{Note}

\setlength{\parskip}{5pt plus 2pt minus 1pt}
\newcolumntype{C}{>{\centering\arraybackslash}p{1.3cm}}
\renewcommand\appendixpagename{Приложение}
\graphicspath{{pics/}}

\title{Обзор проблемы КС-достижимости}
\author{Семён Григорьев}
\date{\today}

\begin{document}

\newgeometry{left=0.8in,right=0.8in,top=1in,bottom=1in}

\maketitle 

\section{Введение}

~\cite{Bradford2018}

\begin{thebibliography}{9}

\bibitem{Bradford2018}
   Bradford P. G. Efficient exact paths for Dyck and semi-Dyck labeled path reachability //arXiv preprint arXiv:1802.05239. --- 2018.

\bibitem{Rytter}
   Rytter W. Context-free recognition via shortest paths computation: a version of Valiant's algorithm //Theoretical Computer Science. --- 1995. --- Т. 143. --- \textnumero. 2. --- С. 343-352.

\bibitem{Lee}
   Lee L. Fast context-free grammar parsing requires fast boolean matrix multiplication //Journal of the ACM (JACM). --- 2002. --- Т. 49. --- \textnumero. 1. --- С. 1-15.

\bibitem{WilliamsWilliams}
   Williams V. V., Williams R. Subcubic equivalences between path, matrix and triangle problems //Foundations of Computer Science (FOCS), 2010 51st Annual IEEE Symposium on. --- IEEE, 2010. --- С. 645-654.

\bibitem{POPL2018}
   Chatterjee K., Choudhary B., Pavlogiannis A. Optimal Dyck reachability for data-dependence and alias analysis //Proceedings of the ACM on Programming Languages. --- 2017. --- Т. 2. --- \textnumero. POPL. --- С. 30.

\bibitem{Yannakakis}
   Yannakakis M. Graph-theoretic methods in database theory //Proceedings of the ninth ACM SIGACT-SIGMOD-SIGART symposium on Principles of database systems. --- ACM, 1990. --- С. 230-242.

\bibitem{Valiant}
   Valiant L. General context-free recognition in less than cubic time. --- 1974.

\bibitem{Reps}
   David Melski and Thomas Reps. 1997. Interconvertbility of set constraints and context-free language reachability. 
   In Proceedings of the 1997 ACM SIGPLAN symposium on Partial evaluation and semantics-based program manipulation (PEPM '97), A. Michael Berman (Ed.). ACM, New York, NY, USA, 74-89. 
   DOI=http://dx.doi.org/10.1145/258993.259006

\bibitem{YARD}
YARD language description. Online. \url{http://yaccconstructor.github.io/YaccConstructor/yard.html}

\bibitem{cons16s}
Rehakova K. et al. Variation in secondary structure of the 16S rRNA molecule in cyanobacteria with implications for phylogenetic analysis //Fottea. – 2014. – Т. 14. – С. 161-178.

\end{thebibliography}

\end{document}