% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Context-Free Shuffle Languages Parsing via Boolean Satisfiability Problem Solving\thanks{The research was supported by the Russian Science Foundation grant 18-11-00100 and a grant from JetBrains Research.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Princeton University, Princeton NJ 08544, USA \and
Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
\email{lncs@springer.com}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
    Verification of concurrent systems is important and nontrivial problem.
    One of directions in this area is modeling of sequential subsystems with push-down automata (PDA) and investigating its communication.
    PDA is equal to context-free languages and ``communication'' may be expressed as shuffle of them.
    In this paper we consider the problem of concurrent programs' model checking from the side of context-free languages shuffle:
    in order to check correctness of system we should check emptiness of intersection of shuffled context-free languages (which describe behavior of the system) with regular language (which describe set of ``bad'' behaviors).
    Even in simple case, when regular language is finite, it leads to NP-complete problem and we show how it can be solved by using SAT-solvers.
    Our reduction is very native and use classical parsing techniques, such as Shared Packed Parse Forest and Generalized LL parsing algorithm, and some ideas from Context-Free Language reachability framework.
    We do not propose solution for arbitrary regular language (existence of which looks an open problem) but we show a some possible directions of research and hope that ever for restricted case proposed solution may be useful.


\keywords{Formal languages  \and Context-free languages \and !!!.}
\end{abstract}
%
%
%

\input{introduction}
\input{preliminaries}
\input{shuffle_to_SAT}
\input{related_work}
\input{conclusion}

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
\bibitem{ref_article1}
Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

\bibitem{ref_lncs1}
Author, F., Author, S.: Title of a proceedings paper. In: Editor,
F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
Springer, Heidelberg (2016). \doi{10.10007/1234567890}

\bibitem{ref_book1}
Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
Location (1999)

\bibitem{ref_proc1}
Author, A.-B.: Contribution title. In: 9th International Proceedings
on Proceedings, pp. 1--2. Publisher, Location (2010)

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017
\end{thebibliography}
\end{document}
