% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{multirow} 
\usepackage{fancyvrb}
\usepackage[misc,geometry]{ifsym} 
\usepackage{subcaption}
\usepackage{caption}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{booktabs}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\setlength{\tabcolsep}{0.7em}

\lstloadlanguages{Python}
\lstset{
  language=Python,
  basicstyle=\scriptsize\sffamily,
  numberstyle=\color{gray},
  stringstyle=\color[HTML]{933797},
  commentstyle=\color[HTML]{228B22}\sffamily,
  emph={[2]from,import,pass,return}, emphstyle={[2]\color[HTML]{DD52F0}},
  emph={[3]range}, emphstyle={[3]\color[HTML]{D17032}},
  emph={[4]for,in,def}, emphstyle={[4]\color{blue}},
  showstringspaces=false,
  breaklines=true,
  prebreak=\mbox{{\color{gray}\tiny$\searrow$}},
  numbers=left,
  xleftmargin=15pt
}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage{hyperref}
\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Genegram: RNA Secondary Structure Prediction Using Formal Grammars and Residual Neural Networks\thanks{Supported by the Russian Science Foundation grant 18-11-00100}}
%
\titlerunning{Genegram: CFG + ResNet}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Polina Lunina\inst{1,2}\orcidID{0000-0002-7172-2647} \Letter \and
Semyon Grigorev\inst{1,2}\orcidID{0000-0002-7966-0698} \and
Vadim Abzalov\inst{1,2}\orcidID{0000-0002-0805-0315}}
%
\authorrunning{Polina Lunina, Semyon Grigorev, Vadim Abzalov}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Saint Petersburg State University, 7/9 Universitetskaya nab., St. Petersburg, 199034, Russia \and
JetBrains Research, Primorskiy prospekt 68-70, Building 1, St. Petersburg 197374, Russia\\
\email{lunina\_polina@mail.ru \Letter, semyon.grigorev@jetbrains.com, vadim.i.abzalov@gmail.com}\\
}
%
\maketitle             % typeset the header of the contribution
%
\begin{abstract}
Improvement in RNA secondary structure prediction accuracy is one of the key focuses in computational genomics due to its crucial role in the functional analysis of RNA molecules. We propose a new approach to solve this problem. Simple context-free grammar describes the most common types of stems and the parsing matrix for some sequence represents all the theoretically possible folds for each subsequence. This information is excessive because only one combination of stems will be presented in a real secondary structure. Moreover, the parsing matrix can be incomplete due to inevitable limitations in grammar. So, we process parsing matrices with residual neural networks in order to generate a valid secondary structure. This approach allows to process pseudoknots, wobble base pairs, and multiple contacts due to the flexible nature of the neural network training process. Our solution was implemented as a Python console tool Genegram.

\keywords{RNA \and Secondary Structure \and Genomic Sequences \and CNN \and ResNet \and Machine Learning  \and Formal Grammars \and Parsing.}
\end{abstract}
%
%
%
\section{Introduction}
\input{introduction}

\section{Proposed Approach Overview}
\input{solution}

\section{Experiments}
\input{experiments}

\section{Conclusion}
\input{conclusion}

%
\bibliography{main} 
\bibliographystyle{ieeetr}
\end{document}
