%4 or 8 pages
%
% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf]{acmart}

\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{verbatim}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{subcaption}

\usepackage{tikz}
\usetikzlibrary{automata,positioning}


%
% defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

% Rights management information.
% This information is sent to you when you complete the rights form.
% These commands have SAMPLE values in them; it is your responsibility as an author to replace
% the commands and values with those provided to you when you complete the rights form.
%
% These commands are for a PROCEEDINGS abstract or paper.
\copyrightyear{2018}
\acmYear{2018}
\setcopyright{acmlicensed}
\acmConference[GRADES-NDA 2019]{GRADES-NDA 2019: the 2nd Joint International Workshop on Graph Data Management Experiences \& Systems (GRADES) and Network Data Analytics (NDA) 2019}{June 30, 2019}{Amsterdam, Netherlands}
%!!\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection, June 03--05, 2018, Woodstock, NY}
\acmPrice{15.00}
\acmDOI{10.1145/1122445.1122456}
\acmISBN{978-1-4503-9999-9/18/06}

%
% These commands are for a JOURNAL article.
%\setcopyright{acmcopyright}
%\acmJournal{TOG}
%\acmYear{2018}\acmVolume{37}\acmNumber{4}\acmArticle{111}\acmMonth{8}
%\acmDOI{10.1145/1122445.1122456}

%
% Submission ID.
% Use this when submitting an article to a sponsored event. You'll receive a unique submission ID from the organizers
% of the event, and this ID should be used as the parameter to this command.
%\acmSubmissionID{123-A56-BU3}

%
% The majority of ACM publications use numbered citations and references. If you are preparing content for an event
% sponsored by ACM SIGGRAPH, you must use the "author year" style of citations and references. Uncommenting
% the next command will enable that style.
%\citestyle{acmauthoryear}

\newcommand{\ltz}{$<0.001$}

%
% end of the preamble, start of the body of the document source.
\begin{document}

%
% The "title" command has an optional parameter, allowing the author to define a "short title" to be used in page headers.
\title[Evaluation of the CFPQ Algorithm Based on Matrix Multiplication]{Evaluation of the Context-Free Path Querying Algorithm Based~on~Matrix Multiplication}

%\and JetBrains Research, Universitetskaya emb., 7-9-11/5A \\ St.Petersburg, Russia \\


%
% The "author" command and its associated commands are used to define the authors and their affiliations.
% Of note is the shared affiliation of the first two authors, and the "authornote" and "authornotemark" commands
% used to denote shared contribution to the research.
\author{Nikita Mishin}
\email{mishinnikitam@gmail.com}
\author{Iaroslav Sokolov}
\email{sokolov.yas@gmail.com}
\author{Egor Spirin}
\email{egor@spirin.tech}
\affiliation{%
  \institution{Saint Petersburg State University}
  \streetaddress{7/9 Universitetskaya nab.}
  \city{St. Petersburg}
  \country{Russia}
  \postcode{199034}
}


\author{Vladimir Kutuev}
\email{vladimir.kutuev@gmail.com}
\author{Egor Nemchinov}
\email{nemchegor@gmail.com}
\author{Sergey Gorbatyuk}
\email{sergeygorbatyuk171@gmail.com}
\affiliation{%
  \institution{Saint Petersburg State University}
  \streetaddress{7/9 Universitetskaya nab.}
  \city{St. Petersburg}
  \country{Russia}
  \postcode{199034}
}


\author{Semyon Grigorev}
\email{s.v.grigoriev@spbu.ru}
\email{semen.grigorev@jetbrains.com}
\orcid{0000-0002-7966-0698}
\affiliation{
  \institution{Saint Petersburg State University}
  \streetaddress{7/9 Universitetskaya nab.}
  \city{St. Petersburg}
  \country{Russia}
  \postcode{199034}
}
\affiliation{
  \institution{JetBrains Research}
  \streetaddress{Universitetskaya emb., 7-9-11/5A}
  \city{St. Petersburg}
  \country{Russia}
  \postcode{199034}
}





%
% By default, the full list of authors will be used in the page headers. Often, this list is too long, and will overlap
% other information printed in the page headers. This command allows the author to define a more concise list
% of authors' names for this purpose.
\renewcommand{\shortauthors}{Mishin and Sokolov, et al.}

%
% The abstract is a short summary of the work to be presented in the article.
\begin{abstract}
Recently proposed matrix multiplication based algorithm for context-free path querying (CFPQ) offloads the most performance-critical parts onto boolean matrices multiplication. Thus, it is possible to utilize modern parallel hardware and software to achieve high performance of CFPQ easily. In this work, we provide results of empirical performance comparison of different implementations of this algorithm on both real data and synthetic data for the worst cases.  
 
%We collect and publish a dataset which includes both real data and syntatica data for worst cases.
%We provide a number of implementations of matrix-based CFPQ algorithm which utilize GPGPUs and high-perfprmance libraries for BMM, and compare performance of them.
%Our evaluation shows that !!! more then 10 times faster then even by naive !!!
\end{abstract}

%
% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002951.10002952.10003197.10010825</concept_id>
<concept_desc>Information systems~Query languages for non-relational engines</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003752.10003766.10003771</concept_id>
<concept_desc>Theory of computation~Grammars and context-free languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010147.10010169.10010170.10010174</concept_id>
<concept_desc>Computing methodologies~Massively parallel algorithms</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003752.10003753.10003761.10003762</concept_id>
<concept_desc>Theory of computation~Parallel computing models</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10010520.10010521.10010528.10010534</concept_id>
<concept_desc>Computer systems organization~Single instruction, multiple data</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Query languages for non-relational engines}
\ccsdesc[500]{Theory of computation~Grammars and context-free languages}
\ccsdesc[500]{Computing methodologies~Massively parallel algorithms}
\ccsdesc[300]{Theory of computation~Parallel computing models}
\ccsdesc[300]{Computer systems organization~Single instruction, multiple data}
%
% Keywords. The author(s) should pick words that accurately describe the work being
% presented. Separate the keywords with commas.
\keywords{Context-free path querying, transitive closure, graph databases, context-free grammar, GPGPU, CUDA, matrix multiplication, boolean matrix}


%
% This command processes the author and affiliation and title information and builds
% the first part of the formatted document.
\maketitle

\input{introduction}
\input{matrix_based_algorithm}
\input{implementation}
\input{dataset}
\input{evaluation}
\input{conclusion}

\begin{acks}
The research was supported by the Russian Science Foundation grant 18-11-00100 and a grant from JetBrains Research.
\end{acks}

%
% The next two lines define the bibliography style to be used, and the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{cfpq_on_gpgpu}

%
% If your work has an appendix, this is the place to put it.
%\appendix

%\section{Research Methods}

%\subsection{Part One}


\end{document}
