% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads,table]{llncs}
%
\usepackage{graphicx}

% For algorithmic attachments and graphics
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{subcaption}
\captionsetup{compatibility=false}

\usepackage{tikz}
\usepackage{dsfont}
\usepackage{multicol, multirow}
%\usepackage{booktabs}
\usetikzlibrary{fit,calc,automata,positioning}


% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Context-Free Path Querying by Using Kronecker Product\thanks{Supported by organization x.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Egor Orachev\inst{1} \and
Ilya Epelbaum\inst{1} \and
Rustam Azimov\inst{1,2} \and
Semyon Grigorev\inst{1,2}\orcidID{0000-0002-7966-0698}}
%
\authorrunning{E. Orachev et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{St.Petersburg State University, 7/9 Universitetskaya nab., St. Petersburg, Russia, 199034 \\
\email{semyon.grigorev@jetbrains.com, iliyepelbaun@gmail.com, rustam.azimov19021995@gmail.com, s.v.grigoriev@spbu.ru} \and
JetBrains Research, Primorskiy prospekt 68-70, Building 1, St. Petersburg, Russia, 197374\\
\email{semyon.grigorev@jetbrains.com}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Context-free path queries (CFPQ) extend regular path queries (RPQ) and allow one to use context-free grammars as constraints for paths. Algorithms for CFPQ are actively developed, but J. Kuijpers recently concludes, that existing algorithms are not performant enough to be used in real-world applications. Thus the development of new algorithms for CFPQ is required. In this paper we provide a new CFPQ algorithm which is based on such linear algebra operations as Kronecker product and transitive closure, and handles grammars presented as recursive state machines. Thus, the proposed algorithm can be implemented by using high-performance libraries and modern parallel hardware, and avoids grammar growth, consequently, it provides the ability for queries optimization.


\keywords{Context-free path querying \and Graph database \and Context-free grammars \and CFPQ \and Kronecker product \and Recursive state machines.}
\end{abstract}
%
%
%

\input{introduction}
\input{recursive_state_machines}
\input{kronecker_cfpq}
\input{evaluation}
\input{conclusion}

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
 \bibliographystyle{splncs04}
 \bibliography{tensor_product_cfpq}
\end{document}
