\section{Error Recovery to CFL Reachcbility}

Editing distance: to find string which is in the given lznguage, such that editing distance is minimal.

\subsection{Preliminaries}

Token is a triple (aka edge).

\subsection{Example}

We start with example.
Suppose, the given language is a simple arithmetic language which is specified by grammar $G_1:$
$$
s \to \text{'('} \ s \ \text{')'} \ | \ \textit{NUM} \ | \ s \ \text{'+'} \ \textit{NUM}.
$$

The string to parse is ''.
Firstly we convert this string into linear graph and fix start and final vertices.

After that we add edges for all possible atomic editing steps (replacement, deletion, insertion).

Now we can solve CFL reachcbility problem and get structural representation of result, or derivation tree for the string which are correst and closest to the input string.

\subsection{Algorithm}

Now we explain how it should works in general case.
Additional edges with error markers goes forward and with all tokens, goes in the its start vertex
(as a result we have loops, but in they can be limited).

\begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,auto]
\node[state] (q_1)   {$1$};
\node[state] (q_2) [above=of q_1] {$2$};
\node[state] (q_3) [above right=of q_1, below right=of q_2] {$0$};
\node[state] (q_4) [right=of q_3] {$3$};
\path[->]
(q_1) edge  node {A} (q_2)
(q_2) edge  node {A} (q_3)
(q_3) edge  node {A} (q_1)
(q_3) edge[bend left, above]  node {B} (q_4)
(q_4) edge[bend left, below]  node {B} (q_3);
\end{tikzpicture}


It is not necessary to create this graph explicetely.
We just can specify generation function which should generate appropriate set of tokens for the given position.

Number of edges may be optimized by filtering with FIRST(k), it can reduce false attempts.

One can customize weight function and maximal length of insertions.

GLL operates woth queue of descriptors.
To optimize order of descriptors processing we introduce priority queue for descriptors.
How to choose priority function? --- ordered tuples!

In the general case we can get SPPF rather single trrr, so it may be necessary to choose the best tree from SPPF after parsing finish.

Priority is a number of additional edges (not from the original input) in processed prefix.
Suffix length.
