% Use pdflatex.exe -synctex=1 -interaction=nonstopmode "graphparsing".tex
% For bibliography use BibTex
\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}


% DOI
\acmDOI{}

% ISBN
\acmISBN{}

%Conference
\acmConference[]{}
\acmYear{}
\copyrightyear{}


\acmArticle{}
\acmPrice{}

\usepackage{graphicx}

\usepackage{caption}
\usepackage{subcaption}
%\usepackage{gnuplottex}
\usepackage{tikz}
\usepackage{mathtools}

\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{textcomp}

\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{algorithmicx}

\settopmatter{printacmref=false, printccs=false, printfolios=false}

\pagestyle{plain} % removes running headers


\begin{document}

\title{Context-Free Path Querying by Matrix Multiplication}

\author{Rustam Azimov}
\affiliation{%
	\institution{Saint Petersburg State University}
	\streetaddress{7/9 Universitetskaya nab.}
	\city{St. Petersburg} 
	\state{Russia} 
	\postcode{199034}
}
\email{rustam.azimov19021995@gmail.com}

\author{Semyon Grigorev}
\affiliation{%
	\institution{Saint Petersburg State University}
	\streetaddress{7/9 Universitetskaya nab.}
	\city{St. Petersburg}
	\state{Russia}  
	\postcode{199034}
}
\email{Semen.Grigorev@jetbrains.com}

% The default list of authors is too long for headers.
\renewcommand{\shortauthors}{B. Trovato et al.}	
		
\newtheorem{mytheorem}{Theorem}
%\newtheorem{lemma}{Lemma}
\newtheorem{mydef}{Definition}

\algrenewcommand\algorithmicindent{0.5em}
\algnewcommand\algorithmicswitch{\textbf{switch}}
\algnewcommand\algorithmiccase{\textbf{case}}
\algnewcommand\algorithmicassert{\texttt{assert}}
\algnewcommand\Assert[1]{\State \algorithmicassert(#1)}
% New "environments"
\algdef{SE}[SWITCH]{Switch}{EndSwitch}[1]{\algorithmicswitch\ #1\ \algorithmicdo}{\algorithmicend\ \algorithmicswitch}
\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1}{\algorithmicend\ \algorithmiccase}

\algtext*{EndSwitch}
\algtext*{EndCase}
\algtext*{EndWhile}% Remove "end while" text
\algtext*{EndIf}% Remove "end if" text
\algtext*{EndFor}% Remove "end for" text
\algtext*{EndFunction}% Remove "end function" text

\newif\ifboldnumber
\newcommand{\boldnext}{\global\boldnumbertrue}




\begin{abstract}
Graph data models are widely used in many areas, for example, graph databases, bioinformatics. In these areas, it is often required to process queries for large graphs. Some of the most common graph queries are navigational queries. The result of query evaluation is a set of implicit relations between nodes of the graph, i.e. paths in the graph. A natural way to specify these relations is by specifying paths using formal grammars over the alphabet of edge labels. An answer to a context-free path query in this approach is usually a set of triples $(A, m, n)$ such that there is a path from the node $m$ to the node $n$, whose labeling is derived from a non-terminal $A$ of the given context-free grammar. This type of queries is evaluated using the \textit{relational query semantics}. Another example of path query semantics is the \textit{single-path query semantics} which requires presenting such a single path for all triples $(A, m, n)$ evaluated using the relational query semantics. There is a number of algorithms for query evaluation which use these semantics but all of them perform poorly on large graphs. One of the open problems in this area is to generalize a well-known matrix-based context-free recognition solution (Valiant 1975) to the context-free path query evaluation. The advantage of the matrix-based algorithm in this area is the possibility of efficient using of a graphics processing unit (GPU) to perform query evaluation on large graphs. The existing algorithms in this area do not allow to use this technique efficiently. In this paper, we show how the context-free path query evaluation using these query semantics can be reduced to the calculation of the matrix transitive closure. Also, we propose a matrix-based algorithm for context-free path query evaluation which uses relational query semantics and makes it possible to speed up computations by using a GPU. Furthermore, we evaluate the performance of our algorithm on some popular ontologies.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
%\begin{CCSXML}
%	<ccs2012>
%	<concept>
%	<concept_id>10010520.10010553.10010562</concept_id>
%	<concept_desc>Computer systems organization~Embedded systems</concept_desc>
%	<concept_significance>500</concept_significance>
%	</concept>
%	<concept>
%	<concept_id>10010520.10010575.10010755</concept_id>
%	<concept_desc>Computer systems organization~Redundancy</concept_desc>
%	<concept_significance>300</concept_significance>
%	</concept>
%	<concept>
%	<concept_id>10010520.10010553.10010554</concept_id>
%	<concept_desc>Computer systems organization~Robotics</concept_desc>
%	<concept_significance>100</concept_significance>
%	</concept>
%	<concept>
%	<concept_id>10003033.10003083.10003095</concept_id>
%	<concept_desc>Networks~Network reliability</concept_desc>
%	<concept_significance>100</concept_significance>
%	</concept>
%	</ccs2012>
%\end{CCSXML}
%

\begin{CCSXML}
	<ccs2012>
	<concept>
	<concept_id>10002951.10002952.10003197.10010825</concept_id>
	<concept_desc>Information systems~Query languages for non-relational engines</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10003752.10003766.10003771</concept_id>
	<concept_desc>Theory of computation~Grammars and context-free languages</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Query languages for non-relational engines}
\ccsdesc[500]{Theory of computation~Grammars and context-free languages}

\keywords{Transitive closure, context-free path querying, graph databases, context-free grammar, GPGPU, matrix multiplication}

\maketitle

\input{introduction.tex}
\input{preliminaries.tex}
\input{related_works.tex}
\input{algorithm.tex}
\input{singlepath.tex}
\input{evaluation.tex}
\input{conclusion.tex}
\input{acknowledgments.tex}

\bibliographystyle{ACM-Reference-Format}
\bibliography{graphparsing}

\end{document}
