\documentclass[12pt]{article}  % standard LaTeX, 12 point type
\usepackage{amsfonts,latexsym}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage[utf8]{inputenc} % Кодировка
\usepackage[english]{babel} % Многоязычность

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{definition}
\newtheorem{definition}{Определение}[section]
\newtheorem{example}{Example}[section]

% unnumbered environments:

\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{notation}{Notation}
\newtheorem*{note}{Note}

\setlength{\parskip}{5pt plus 2pt minus 1pt}
%\setlength{\parindent}{0pt}

\usepackage{color}
\usepackage{listings}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{ucs}

\newcommand{\tab}[1][0.3cm]{\ensuremath{\hspace*{#1}}}
% A generalized view on parsing and translation
% http://dl.acm.org/citation.cfm?id=2206331
\title{Generalized LL For CF Languages Intersection}
\author{Semyon Grigorev}
%\date{\today}

\begin{document}

\maketitle

Let we have two CF grammars $G_1$ and $G_2$. 
Both this grammars can be presented as deterministic recursive finite automaton.
Position in grammar is number of automata state.

We will use two different GSS for construction of $G_1$ and $G_2$ intersection.

Let descriptor is $(g_1,p_1,g_2,p_2)$ where $g_1$ and $g_2$ are GSS nodes, $p_1$ and $p_2$ are 
positions in grammars. Each triple $(g_1, p_1, p_2)$ and $(g_2, p_2, p_1)$ can be considered as 
descriptors in origins GLL (GSS node, position in grammar, position in input). 

Let we have a descriptor of form $(g_1,p_1,g_2,p_2)$. Possible cases.
\begin{itemize}
\item There are token edges from $p_1$ and $p_2$ which are labelled with same tokens. For each pair 
of edges $(p_1,t,q_1)(p_2,t,q_2)$ create a descriptor $(g_1,q_1,g_2,q_2)$
\item Other cases can be processed independently by the similar to classical GLL way.
\end{itemize}

Is it possible to store $U$ and $P$ in GSS-s?

Complexity?

Correctness?

Formalize in Coq?


\end{document}