\section{Background}\label{sec:background}

In this section we briefly discuss GraphBLAS and introduce tree-based matrix representation.
We also provide the basics of interaction nets, discuss implementations of interaction nets, and motivate our choice of which one we use for evaluation.

\subsection{GraphBLAS}

The goal of the GraphBLAS API is to provide basic building blocks for constructing high-performance graph analysis algorithms.
To achieve this goal, GraphBLAS utilizes \textbf{sparse generic} linear algebra.
Operations over sparse generic matrices and vectors parameterized by semiring-like structures provide a powerful basis for expressing a wide range of graph analysis algorithms~\cite{8778338}.

GraphBLAS API: C API, C++ API, implementations in Java, Julia, etc.

Challenges on this path include the implementation of high-performance generic operations over sparse data structures.
Imbalanced subtasks together with irregular data access patterns make load balancing — and thus efficient parallel implementation — hard.
Moreover, graph analysis algorithms are not single operations but nontrivial chains of operations over matrices and vectors.
This leads to the problem of intermediate data structures.
Kernel fusion could solve these problems, but it is too difficult to implement automatic optimizations for sparse generic computations, so it is done manually in the most frequent cases.
Attempts to use MLIR (GraphBLAS dialect).!!!
Poor utilization on GPUs. Nontrivial challenge to create generic kernels for GPU and FPGA.!!!
There are also other problems discussed in ``What Did GraphBLAS Get Wrong?'' by John Gilbert\footnote{Talk by John Gilbert, UC Santa Barbara, GraphBLAS BoF at HPEC, September 2022. Details can be found here: \url{https://sites.cs.ucsb.edu/~gilbert/talks/talks.htm}.}

SuiteSparse:GraphBLAS serves as the reference implementation of this API, offering a highly optimized, multi-threaded C library that supports various sparse matrix formats.
Building on this foundation, LAGraph provides a collection of production-ready graph algorithms (e.g., PageRank, connected components, triangle counting) implemented using SuiteSparse:GraphBLAS.
Currently it is one of the state-of-the-art (SOTA) solutions for high-performance graph analysis, thus we use it as a baseline for evaluation.

The long-term goal of our work is to investigate how interaction nets can help solve challenges with irregular parallelism and, possibly, with kernel fusion.
In this work, we do not focus on the GraphBLAS API implementation, but we are inspired by it: we implemented a minimal set of generic data structures and operations sufficient to implement several classic graph analysis algorithms.
 

\subsection{Tree-based Representation for Sparse Matrices and Vectors}

To store sparse matrices and vectors efficiently, we employ a quad-tree data structure. The matrix is recursively subdivided into four quadrants~--- northwest (NW), northeast (NE), southwest (SW), and southeast (SE)~--- until each block becomes either uniform or empty.
The Figure~\ref{fig:quadtree_example} shows an example of a matrix in the quad-tree form.

\begin{figure}[htbp]
    \centering
    \caption{Example of the matrix and its quad-tree representation}
\begin{subfigure}[b]{0.9\textwidth}
    \centering
    \caption{Block decomposition of the matrix}
\begin{tikzpicture}[scale=0.8, font=\sffamily]

  \draw[step=1cm, black!70, thick] (0,0) grid (8,8);
  
  \foreach \x/\y/\val in {
    0/0/1, 1/0/1, 2/0/1, 3/0/1, 4/0/5, 5/0/5, 6/0/7, 7/0/7,
    0/1/1, 1/1/1, 2/1/1, 3/1/1, 4/1/5, 5/1/5, 6/1/7, 7/1/7,
    0/2/1, 1/2/1, 2/2/1, 3/2/1, 4/2/6, 5/2/6, 6/2/8, 7/2/8,
    0/3/1, 1/3/1, 2/3/1, 3/3/1, 4/3/6, 5/3/6, 6/3/8, 7/3/8,
    0/4/2, 1/4/2, 2/4/2, 3/4/2, 4/4/3, 5/4/3, 6/4/4, 7/4/9,
    0/5/2, 1/5/2, 2/5/2, 3/5/2, 4/5/3, 5/5/3, 6/5/10, 7/5/11,
    0/6/2, 1/6/2, 2/6/2, 3/6/2, 4/6/12, 5/6/12, 6/6/13, 7/6/13,
    0/7/2, 1/7/2, 2/7/2, 3/7/2, 4/7/12, 5/7/12, 6/7/13, 7/7/13
  } {
    \node at (\x+0.5, \y+0.5) {\val};
  }

\node at (9.5, 4) {$=$};

\begin{scope}[shift={(11,0)}]
    \draw[step=1cm, gray!10, very thin] (0,0) grid (8,8);
    
    \draw[draw=black, thick] (0,4) rectangle ++(4,4) node[midway, font=\Large\bfseries] {2 (NW)};
    \draw[draw=black, thick] (0,0) rectangle ++(4,4) node[midway, font=\Large\bfseries] {1 (SW)};
    
    \draw[draw=black, thick] (4,6) rectangle ++(2,2) node[midway, font=\bfseries] {12 (NW)};
    \draw[draw=black, thick] (6,6) rectangle ++(2,2) node[midway, font=\bfseries] {13 (NE)};
    \draw[draw=black, thick] (4,4) rectangle ++(2,2) node[midway, font=\bfseries] {3 (SW)};
    
    \draw[draw=black, thick] (6,4) rectangle ++(1,1) node[midway, font=\tiny\bfseries] {4(SW)};
    \draw[draw=black, thick] (7,4) rectangle ++(1,1) node[midway, font=\tiny\bfseries] {9(SE)};
    \draw[draw=black, thick] (6,5) rectangle ++(1,1) node[midway, font=\tiny\bfseries] {10(NW)};
    \draw[draw=black, thick] (7,5) rectangle ++(1,1) node[midway, font=\tiny\bfseries] {11(NE)};
    
    \draw[draw=black, thick] (4,2) rectangle ++(2,2) node[midway, font=\bfseries] {6 (NW)};
    \draw[draw=black, thick] (6,2) rectangle ++(2,2) node[midway, font=\bfseries] {8 (NE)};
    \draw[draw=black, thick] (4,0) rectangle ++(2,2) node[midway, font=\bfseries] {5 (SW)};
    \draw[draw=black, thick] (6,0) rectangle ++(2,2) node[midway, font=\bfseries] {7 (SE)};
    

  \end{scope}
\end{tikzpicture}

\label{fig:quadtree_blocks}
\end{subfigure}

\vspace{1cm}

\begin{subfigure}[b]{0.9\textwidth}
    \centering
    \caption{Quadtree representation of the blocks in~\ref{fig:quadtree_blocks}}
\begin{tikzpicture}[
    grow=down,
    level 1/.style={sibling distance=3.5cm, level distance=2cm},
    level 2/.style={sibling distance=1.2cm, level distance=2cm},
    level 3/.style={sibling distance=1.1cm, level distance=2cm},
    edge from parent/.style={draw, -latex, thick},
    int/.style={draw, circle, fill=black, font=\small\bfseries, minimum size=0.5cm, inner sep=0pt},
    leaf/.style={draw, rectangle, fill=gray!20, font=\large\bfseries, minimum size=0.8cm, inner sep=2pt, rounded corners=2pt},
    q/.style={font=\scriptsize\itshape, pos=0.5, fill=white, inner sep=1pt, rounded corners=1pt}
]

\node[int] { }
    child { node[leaf] {2} edge from parent node[q, left] {NW} }
    child {
        node[int] { }
        child { node[leaf] {5} edge from parent node[q, left] {NW} }
        child { node[leaf] {6} edge from parent node[q, left] {SW} }
        child { node[leaf] {7} edge from parent node[q, right] {NE} }
        child {
            node[int] {}
            child { node[leaf] {10} edge from parent node[q, left] {NW} }
            child { node[leaf] {12} edge from parent node[q, left] {SW} }
            child { node[leaf] {11} edge from parent node[q, right] {NE} }
            child { node[leaf] {13} edge from parent node[q, right] {SE} }
            edge from parent node[q, right] {SE}
            }
        edge from parent node[q, left] {NE}
    }
    child { node[leaf] {1} edge from parent node[q, right] {SW} }
    child {
        node[int] { }
        child { node[leaf] {3} edge from parent node[q, left] {NW} }
        child { node[leaf] {8} edge from parent node[q, left] {SW} }
        child { node[leaf] {9} edge from parent node[q, right] {NE} }
        child { node[leaf] {4} edge from parent node[q, right] {SE} }
        edge from parent node[q, right] {SE}
    };
\end{tikzpicture}
\label{fig:quadtree_tree}
\end{subfigure}
\label{fig:quadtree_example}
\end{figure}



This recursive partitioning allows large uniform blocks to be represented implicitly, thereby reducing the memory required to store the matrix.
Unlike flat array formats such as CSR or CSC, which are standard in libraries like SuiteSparse:GraphBLAS, the quad-tree naturally supports a divide-and-conquer execution model and adapts dynamically to irregular sparsity patterns.
This structure simplifies parallelization of element-wise, matrix-matrix, and matrix-vector operations, as independent sub-blocks can be processed recursively without global synchronization.

However, this flexibility comes at a cost: for dense or highly regular matrices, the overhead of maintaining tree metadata and pointer indirection can outweigh the benefits of sparse storage, making traditional linear formats more efficient in such cases.


\subsection{Interaction Nets}

\label{sec:interaction_nets}

We utilize the variant of definitions provided by Y. Lafont in 1989~\cite{10.1145/96709.96718}. Interaction Nets is a graph-rewriting model where computation proceeds through local rewriting rules applied to pairs of connected nodes.

\textit{(Agent)}. An \textit{agent} is a symbol from a fixed alphabet $\Sigma$, for which:
\begin{itemize}
    \item an arity is defined --- a function $Ar: \Sigma \to \mathbb{N} \cup \{0\}$;
    \item there is one \textit{principal port} and a number of \textit{auxiliary ports} equal to its arity.
\end{itemize}

\textit{(Net)}. A \textit{net} over the alphabet $\Sigma$ is a labeled undirected multi-graph in which:
\begin{itemize}
    \item each node is labeled by an agent;
    \item edges can only exist between ports;
    \item at most one edge enters each port.
\end{itemize}

Let $\mathcal{N}_\Sigma$ denote the set of all possible nets over the alphabet $\Sigma$.

We say that a port is \textit{free} if no edge originates from it. A net can also consist simply of a set of edges, in which case each end of an edge is considered a free port.

\textit{(Interface)}. The \textit{interface} $I(n)$ is the set of all free ports of a net $n$.

We say that port $p_a$ is connected to port $p_b$ if there is an edge between them.

Computation in Interaction Nets occurs via \textit{reductions}: rewriting an \textit{active pair} into a net. An active pair is defined as two agents connected via their principal ports.

\textit{(Reduction Rule)}. A \textit{reduction rule} is a graph rewriting rule. Let $R$ be the set of reduction rules. More formally, a reduction rule is a triplet $r \in \Sigma \times \Sigma \times \mathcal{N}_\Sigma$, for which the following properties hold:
\begin{itemize}
    \item \textbf{Symmetry:} $(l_1, l_2, n) \in R \Rightarrow (l_2, l_1, n) \in R$
    \item \textbf{Determinism:} $(l_1, l_2, n_1) \in R \land (m_1, m_2, n_2) \in R \Rightarrow (l_1, l_2) \neq (m_1, m_2)$
    \item \textbf{Interface Preservation:} $\forall (l_1, l_2, n) \in R, Ar(l_1) + Ar(l_2) = \|I(n)\|$
\end{itemize}

An example of the computation of the net encoding of the term $(0 + 1) + (0 + 1)$ is shown in Figure~\ref{fig:net_example}.
The set of agents is $\Sigma=\{0, S, Add\}$ with arities $Ar(0)=0$, $Ar(S)=1$ and $Ar(Add)=2$ and with the rules shown in Figure~\ref{fig:peano_rules}.
The left auxiliary ports of the active $Add$-agents are connected to the net representing $1$, encoded in Peano numerals as $S(0)$, and the right auxiliary ports are connected to the third $Add$-agent.
There is one free port~--- one of the auxiliary ports of the $Add$-agent.
This constitutes the interface.

\begin{figure}[htbp]
\centering
\caption{Example of computation in interaction nets: reduction of the term $(1+0) + (1+0)$ encoded in Peano numerals }
\begin{tikzpicture}[>=Stealth, node distance=0.5cm]

\begin{scope}[local bounding box=group1, baseline=(group1.center)]
    \inetcell[draw=darkred](add1){$Add$}[U]
    \inetcell[above=of add1.pal, draw=darkred](zero1){$0$}[D]
    \inetcell[below=of add1.left pax](s1){$S$}[U]
    \inetcell[below=of s1.middle pax](zero2){$0$}[U]
    
    \inetwire[darkred](add1.pal)(zero1.pal)
    \inetwire(add1.left pax)(s1.pal)
    \inetwire(s1.middle pax)(zero2.pal)
    
    \inetcell[below right=5em and 1em of add1](addMain){$Add$}[L]
    \inetwire(add1.right pax)(addMain.pal)
    \inetwirefree(addMain.right pax)
    
    \inetcell[right=6em of add1, draw=darkred](add2){$Add$}[U]
    \inetcell[above=of add2.pal, draw=darkred](zero3){$0$}[D]
    \inetcell[below=of add2.left pax](s2){$S$}[U]
    \inetcell[below=of s2.middle pax](zero4){$0$}[U]
    
    \inetwire[darkred](add2.pal)(zero3.pal)
    \inetwire(add2.left pax)(s2.pal)
    \inetwire(s2.middle pax)(zero4.pal)
    
    \draw[thick] (add2.right pax) .. controls +(down:30mm) and +(right:20mm) .. (addMain.left pax);
\end{scope}

\begin{scope}[shift={(group1.east)}, xshift=2.5cm, local bounding box=group2, baseline=(group2.center)]
    \inetcell[draw=darkred](addRed1){$Add$}[U]
    \inetcell[above=of addRed1.pal, draw=darkred](sRed1){$S$}[D]
    \inetcell[above = of sRed1.middle pax](zeroRed1){$0$}[D]
    \inetcell[below= of addRed1.left pax](sRed2){$S$}[U]
    \inetcell[below= of sRed2.middle pax](zeroRed2){$0$}[U]
    
    \inetwire(zeroRed1.pal)(sRed1.middle pax)
    \inetwire(zeroRed2.pal)(sRed2.middle pax)
    \inetwire(sRed2.pal)(addRed1.left pax)
    \inetwire[darkred](sRed1.pal)(addRed1.pal)
    \inetwirefree(addRed1.right pax)
\end{scope}


\begin{scope}[shift={(group2.east)}, xshift=2.5cm, local bounding box=group3, baseline=(group3.center)]
    \inetcell[draw=darkred](addRed1){$Add$}[U]
    \inetcell[below=of addRed1.right pax](sRed1){$S$}[D]
    \inetcell[above=of addRed1.pal, draw=darkred](zeroRed1){$0$}[D]
    \inetcell[below= of addRed1.left pax](sRed2){$S$}[U]
    \inetcell[below= of sRed2.middle pax](zeroRed2){$0$}[U]
    
    \inetwire[darkred](zeroRed1.pal)(addRed1.pal)
    \inetwire(zeroRed2.pal)(sRed2.middle pax)
    \inetwire(sRed2.pal)(addRed1.left pax)
    \inetwire(sRed1.middle pax)(addRed1.right pax)
    \inetwirefree(sRed1.pal)
\end{scope}

\begin{scope}[shift={(group3.east)}, xshift=2.5cm]
    \inetcell(sRed1){$S$}[D]
    \inetcell[above = of sRed1.middle pax](zeroRed1){$0$}[D]
    \inetcell[below = of sRed1.pal](sRed2){$S$}[D]
    
    \inetwire(zeroRed1.pal)(sRed1.middle pax)
    \inetwire(sRed1.pal)(sRed2.middle pax)
    \inetwirefree(sRed2.pal)
\end{scope}


\coordinate (arrowY) at (group1.east);

\draw[->, thick] ([xshift=0.5cm]group1.east) -- ++(0.9,0);
\draw[->, thick] ([xshift=0.5cm]group2.east |- arrowY) -- ++(0.9,0);
\draw[->, thick] ([xshift=0.5cm]group3.east |- arrowY) -- ++(0.9,0);

\end{tikzpicture}
\label{fig:net_example}
\end{figure}


\begin{figure}[htbp]
    \centering
    \caption{Reduction rules for Peano numerals}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
    
    
\begin{tikzpicture}[>=Stealth, node distance=0.5cm]
    \begin{scope}[local bounding box=group1]
        
        \inetcell(Add){$\text{Add}$}[U]
        \inetcell[above = of Add.pal](0){$0$}[D]
        % \node (p) [right=of 0.pal] {$\phantom x$};
        % \inetwirefree(Add.right pax)
        % \inetwirefree(Add.left pax)
        \inetwire(0.pal)(Add.pal)
        \node (l) [below = of Add.left pax, draw, circle, pattern color=darkblue, pattern=north east lines] {};
        \node (r) [below = of Add.right pax, draw, circle, pattern color=darkred, pattern=crosshatch] {};
        \draw[thick] (Add.left pax) -- (l);
        \draw[thick] (Add.right pax) -- (r);
    \end{scope}

    \draw[->, thick] (group1.east) ++(0.5,0) -- ++(0.9,0);

    \begin{scope}[shift={(group1.east)}, xshift=2.5cm]
        \matrix[row sep=1em]{
            \node (t) [draw, circle, pattern color=darkblue, pattern=north east lines] {};  \\
            \node (p1) {$\phantom x$}; \\
            \node (b1) [draw, circle, pattern color=darkred, pattern=crosshatch] {}; \\};
            \inetwirecoords(t)(b1)
    \end{scope}
\end{tikzpicture}
\caption{ Rule for $0 + x = x$}
\end{subfigure}
\begin{subfigure}[b]{0.45\textwidth}
    \centering

\begin{tikzpicture}[>=Stealth, node distance=0.5cm]
    \begin{scope}[local bounding box=group1]
        \matrix[row sep=1em]{
            \inetcell(S){$S$}[D] \\
            \inetcell(Add){$\text{Add}$}[U] \\};
            \node (p) [right=of S.pal] {$\phantom x$};
            % \inetwirefree(S.middle pax)
            % \inetwirefree(Add.left pax)
            % \inetwirefree(Add.right pax)
            \inetwire(S.pal)(Add.pal)
            \node (la) [below = of Add.left pax, draw, circle, pattern color=darkred, pattern=crosshatch] {};
            \node (ra) [below = of Add.right pax, draw, circle, pattern color=darkblue, pattern=north east lines] {};
            \node (sm) [above = of S.middle pax, draw, circle, pattern color=darkgreen, pattern=horizontal lines] {};

            \draw[thick] (Add.right pax) -- (ra);
            \draw[thick] (Add.left pax) -- (la);
            \draw[thick] (S.middle pax) -- (sm);
    \end{scope}

    \draw[->, thick] (group1.east) ++(0.5,0) -- ++(0.9,0);

    \begin{scope}[shift={(group1.east)}, xshift=2.5cm]
        \inetcell(Add){$\text{Add}$}[U]
        \inetcell[node distance=1em, below=of Add.right pax](S){$S$}[D]
        % \node (p) [right=of Add.middle pax] {$\phantom x$};
        % \inetwirefree(Add.pal)
        % \inetwirefree(Add.left pax)
        % \inetwirefree(S.pal)
        \inetwire(Add.right pax)(S.middle pax)

        \node (ap) [above = of Add.pal, draw, circle, pattern color=darkgreen, pattern=horizontal lines] {};
        \node (la) [below = of Add.left pax, draw, circle, pattern color=darkred, pattern=crosshatch] {};
        \node (sp) [below = of S.pal, draw, circle, pattern color=darkblue, pattern=north east lines] {};

        \draw[thick] (Add.pal) -- (ap);
        \draw[thick] (Add.left pax) -- (la);
        \draw[thick] (S.pal) -- (sp);
    \end{scope}
\end{tikzpicture}
\caption{ Rule for $S(x) + y = S(x + y)$}
\end{subfigure}
\label{fig:peano_rules}
\end{figure}

As shown in Lafont's work~\cite{10.1145/96709.96718}, $\lambda$-calculus can be expressed through this formalism.
Additionally, reductions can be performed independently of each other and it has been proven that the order of reductions doesn't matter. This serves as the primary reason for the natural suitability of Interaction Nets for parallel computations.

\subsection{Implementations of Interaction Nets}

 Proposed by Lafont, interaction nets can serve as a foundation for irregular parallel computations.

 Notable implementations include Inpla, HVM, and Vine. In OCaml~\cite{Huber_2025} (Comparison, Inpla fastest)

    Vine --- high-level language (Rust-like) + interaction-nets-based intermediate representation + parallel virtual machine.

    HVM --- high-level language (JS-like) + interaction-nets-based intermediate representation + parallel virtual machine (+GPGPU?).

    Inpla --- mature enough, demonstrates good performance and scalability. Interaction-nets-based language + parallel VM.

No information is available on the evaluation or comparison of Vine and HVM.
Typical benchmarks for these systems include Fibonacci, Ackermann, sorting, and list traversal. 
No graph analysis algorithms while such algorithms looks grate candidates for interaction nets.
Other projects implement interaction nets in languages such as OCaml and Haskell.


We are focused on languages and runtimes. 

In Haskell and OCaml

Inpla is a low-level interaction nets programming language, equipped with a parallel runtime. Programming in Inpla requires defining a custom set of rewriting rules for a chosen set of agents. Computation begins when two agents are connected via their primary ports.
\begin{listing}[H]
\caption{Peano numerals addition in Inpla}
\begin{verbatim}
// Defining custom rules
Add(r, x) >< Z =>
    r ~ x;
Add(r, x) >< S(y) =>
    Add(r, S(x)) ~ y;

// 1 + 1: connecting Add and S
Add(r, S(Z)) ~ S(Z);
// Printing the result and freeing the `r` name
r; free r;
\end{verbatim}
\end{listing}
Vine

IVY, IVM. Differences with Inpla: fixed set of agents (and respective rewriting rules).


HVM: fixed set of agents (and respective rewriting rules). Bend.  Acceleration in Cuda. \footnote{Issues, problems, eg with QTree}