\section{Conclusion and Future Work}\label{sec:conclusion}

We have demonstrated a feasible mapping of GraphBLAS concepts to interaction nets using quadtrees, provided concrete implementations in Inpla and Vine, and evaluated their scalability.
The performance gap compared to state-of-the-art parallel linear-algebra-based frameworks remains substantial, and scalability is limited.

Several bottlenecks and limitations were identified.
The naive implementation suffers from a duplication problem: a huge number of duplicated nodes limits both performance and memory efficiency.
Reordering (RCM) is crucial for performance and memory consumption but is not handled automatically.
Currently it is a separated preprocessing step.
Moreover, only a limited set of operations and algorithms has been evaluated.

Future work includes investigating improved duplication strategies.
The set of implemented operations and algorithms should be extended to include indexed operations, slices, and such graph analysis algorithms as PageRank, minimum spanning forest (MSF), regular path queries, context-free path queries (CFPQ), and maximal independent set (MIS).

A systematic evaluation of other interaction-net-based implementations, such as HVM, is also required.
Alternative sparse matrix representations, such as the coordinate list (COO) format, rely on sorting---an operation that parallelizes well in interaction nets and therefore warrants further investigation.

Automatic kernel fusion presents another direction.
Advanced optimization techniques natural to functional programming, such as supercompilation~\cite{10.1007/3-540-62064-8_20} or distillation~\cite{10.1007/978-3-030-98869-2_7}, could be applied to automatically fuse sequences of operations, analogous to manual fusion for masks and other constructs in classic GraphBLAS.
Partial evaluation for interaction nets~\cite{DBLP:conf/sas/Bechet92}
Whether these techniques can be adapted for interaction nets remains an open question.

Finally, hardware acceleration remains an open opportunity.
Designing an FPGA specifically optimized for interaction net reduction could exploit the fine-grained parallelism inherent to the model and yield substantial performance gains.