%% GCM 2026
%% https://conf.researchr.org/home/staf-2026/gcm-2026#Call-for-Papers
%% Regular papers of at most 16 pages describing innovative contributions.
%% Position papers, system descriptions or work in progress of 6 to 12 pages.



%% The first command in your LaTeX source must be the \documentclass command.
%%
%% Options:
%% twocolumn : Two column layout. Do not use twocolumn for papers submitted to CEUR-WS!
%% hf: enable header and footer.
\documentclass[
% twocolumn,
% hf,
]{ceurart}

%%
%% One can fix some overfulls
\sloppy

%%
%% Minted listings support 
%% Need pygment <http://pygments.org/> <http://pypi.python.org/pypi/Pygments>
\usepackage{listings}
%% auto break lines
\lstset{breaklines=true}

\usepackage{minted}

\usepackage{tikz}
\usepackage{tikz-inet}
\usetikzlibrary{positioning, arrows.meta, patterns}
\usepackage{subcaption}
\usepackage{xcolor}
\colorlet{darkred}{red!60!black}
\colorlet{darkblue}{blue!60!black}
\colorlet{darkgreen}{green!60!black}


%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% Rights management information.
%% CC-BY is default license.
\copyrightyear{2022}
\copyrightclause{Copyright for this paper by its authors.
  Use permitted under Creative Commons License Attribution 4.0
  International (CC BY 4.0).}

%%
%% This command is for the conference information
\conference{Woodstock'22: Symposium on the irreproducible science,
  June 07--11, 2022, Woodstock, NY}

%%
%% The "title" command
\title{Graph Analysis via Graph Rewriting: Sparse Linear Algebra Based Graph Analysis in Interaction Nets}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
\author[1]{Danil Zaripov}[%
orcid=0000-0002-0877-7063,
email=kulyabov-ds@rudn.ru,
]
\cormark[1]

\author[1]{Efim Kubyshkin}[%
orcid=0009-0009-5531-1758,
email=e.kubyshkin@spbu.ru,
]

\author[1]{Nikolai Ponomarev}[%
orcid=0009-0000-2382-5687,
email=n.ponomarev@spbu.ru,
]

\author[1]{Georgii Sichkar}[%
orcid=0000-0002-9421-8566,
email=Manfred.Jeusfeld@acm.org,
]

\author[1]{Vladimir Zaikin}[%
orcid=0000-0002-9421-8566,
email=Manfred.Jeusfeld@acm.org,
]

\author[1]{Semyon Grigorev}[%
orcid=0000-0002-7966-0698,
email=s.v.grigoriev@mail.spbu.ru,
]


\address[1]{Saint Petersburg State University,
  7-9 Universitetskaya Embankment, St. Petersburg, 199034, Russia}



%% Footnotes
\cortext[1]{Corresponding author.}

%%
%% The abstract is a short summary of the work to be presented in the
\begin{abstract}
%% article.
  A clear and well-documented \LaTeX{} document is presented as an
  article formatted for publication by CEUR-WS in a conference
  proceedings. Based on the ``ceurart'' document class, this article
  presents and explains many of the common variations, as well as many
  of the formatting elements an author may use in the preparation of
  the documentation of their work.
\end{abstract}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\begin{keywords}
  Interaction Nets \sep
  graph rewriting \sep
  graph analysis algorithms \sep
  GraphBLAS \sep
  sparse linear algebra \sep
  quad-tree
\end{keywords}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\input{Introduction.tex}
\input{Background.tex}
\input{Implementation_details.tex}
\input{Evaluation.tex}
\input{Conclusion.tex}
\input{Code_availability.tex}

%  \section{Appendices}
%  
%  If your work needs an appendix, add it before the
%  ``\verb|\end{document}|'' command at the conclusion of your source
%  document.
%  
%  Start the appendix with the ``\verb|\appendix|'' command:
%  \begin{lstlisting}
%  \appendix
%  \end{lstlisting}
%  and note that in the appendix, sections are lettered, not
%  numbered. 

%%
%% The acknowledgments section is defined using the "acknowledgments" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acknowledgments}
  Thanks to the developers of ACM consolidated LaTeX styles
  \url{https://github.com/borisveytsman/acmart} and to the developers
  of Elsevier updated \LaTeX{} templates
  \url{https://www.ctan.org/tex-archive/macros/latex/contrib/els-cas-templates}.  
\end{acknowledgments}

%% The declaration on generative AI comes in effect
%% in Janary 2025. See also
%% https://ceur-ws.org/GenAI/Policy.html
\section*{Declaration on Generative AI}
 %(by using the activity taxonomy in ceur-ws.org/genai-tax.html)
 
 During the preparation of this work, the author(s) used X-GPT-4 and Gramby in order to: Grammar and spelling check. Further, the author(s) used X-AI-IMG for figures 3 and 4 in order to: Generate images. After using these tool(s)/service(s), the author(s) reviewed and edited the content as needed and take(s) full responsibility for the publication’s content. 

%%
%% Define the bibliography file to be used
\bibliography{GraphBLAS_in_Interaction_Nets}

%%
%% If your work has an appendix, this is the place to put it.
% \appendix
% 
% \section{Online Resources}
% 
% 
% The sources for the ceur-art style are available via
% \begin{itemize}
% \item \href{https://github.com/yamadharma/ceurart}{GitHub},
% % \item \href{https://www.overleaf.com/project/5e76702c4acae70001d3bc87}{Overleaf},
% \item
%   \href{https://www.overleaf.com/latex/templates/template-for-submissions-to-ceur-workshop-proceedings-ceur-ws-dot-org/pkfscdkgkhcq}{Overleaf
%     template}.
% \end{itemize}

\end{document}

%%
%% End of file

