\section{Implementation Details}\label{sec:implementation_details}

In this section we describe our implementations: in F\# (used as a prototype and reference for other implementations), in Inpla (a low-level language), and in Vine (a high-level language).
We present the data structures and functions, list graph analysis algorithms (with short descriptions), and discuss details related to the implementation of the described structures and functions in selected languages.

\subsection{Basic Data Structures and Functions}\label{sec:basic_structures_and_functions}

Matrix representation, as described in the F\# code, is shown in Listing~\ref{lst:matrix_representation}.
Our representation is generic, meaning it is parameterized by the matrix cell type.
The structure does not only work for matrices with sizes that are powers of two.
A user-defined matrix of arbitrary size will be aligned to a square storage with size equal to a power of two using \verb|Dummy| blocks.
Since the representation is compressed, the \verb|Storage| stores the real size (as a power of two) together with the respective tree. Finally, the \verb|SparseMatrix| stores the matrix dimensions and the number of non-zero entries (nnz) together with data represented as \verb|Storage|.
Data structures for a sparse vector are described in a similar way.

We propose to use \verb|Option<'T>| to explicitly represent cells that have no values.
This approach is natural for graph representation: a cell of the adjacency matrix of a graph with edge labels of type \verb|'T| has type \verb|Option<'T>|: either there exists an edge with a label of the respective type (\verb|Some(l)|), or there is no edge at all (\verb|None|).
Thus, for a sparse matrix with filled cells of type \verb|'T|, the storage has type \verb|Option<'T>|.

\begin{listing}
\begin{minted}{ocaml}
type treeValue<'value> = | Dummy | UserValue of 'value

(*
| x1 | x2 |
----------
| x3 | x4 |
*)
type qtree<'value> =
    | Node of qtree<'value> * qtree<'value> * qtree<'value> * qtree<'value>
    | Leaf of treeValue<'value>

[<Struct>]
type Storage<'value> =
    // Storage is always size-x-size square.
    val size: uint64<storageSize>
    val data: qtree<'value>
    new(_size, _data) = { size = _size; data = _data }

[<Struct>]
type SparseMatrix<'value> =
    val nrows: uint64<nrows>
    val ncols: uint64<ncols>
    val nvals: uint64<nvals>
    val storage: Storage<Option<'value>>
\end{minted}
\caption{Matrix representation}
\label{lst:matrix_representation}
\end{listing}

We implemented a minimal set of operations over matrices and vectors required to implement selected graph analysis algorithms.
The list of basic functions includes \verb|map|, \verb|fold|, \verb|reduce|.
Element-wise operations are based on \verb|map2|.
Particularly, \verb|mask| is implemented as a case of \verb|map2|.
Linear algebra functions include: \verb|vxm| (generic vector-matrix multiplication) and \verb|mxm| (generic matrix-matrix multiplication).
We also support transpose and extracting the triangular part of a matrix (required for the triangle counting algorithm).
We did not implement fused versions.
Conversion to and from coordinate list format is provided to simplify input graph building.
All operations take sparsity into account and preserve compression.
No manual parallelization to !!!


Exception handling is unlikely to be possible in interaction nets in the traditional way due to their highly parallel and irregular nature.
Moreover, the selected languages (Inpla and Vine) do not provide any specific mechanisms for exception handling.
We focus on realistic code, so we should handle not only correct cases but also erroneous ones (e.g., inconsistent matrix or vector sizes).
For this, we use the \verb|Result| type (even in the F\# version)\footnote{We do this to be as close to interaction-net-based implementations as possible. The F\# language has traditional exception mechanisms.} that explicitly represents two cases: \verb|Ok<'R>| for successfully completed computations and \verb|Err<'E>| for errors. 
This way of handling errors is represented in many languages such as Rust, F\#, etc.

\subsection{Implemented Graph Analysis Algorithms}\label{sec:graph_algorithms}

We implemented three basic graph analysis algorithms.
\begin{itemize}
    \item Level-based BFS: for the given graph and start vertex, it returns a sparse vector \verb|result| such that if \verb|result[i] = Some(k)|, then vertex \verb|i| was visited at step \verb|k|.
    If \verb|result[i] = None|, then \verb|i| is unreachable.
    \item SSSP: for the given graph and start vertex, it returns a sparse vector \verb|result| such that if \verb|result[i] = Some(k)|, then the shortest path from the start vertex to \verb|i| has length \verb|k|.
    If \verb|result[i] = None|, then \verb|i| is unreachable. 
    To prevent an infinite loop in the case of negative cycles, we check that the number of iterations should not be greater than the number of vertices in the input graph.
    \item Triangle counting algorithm.
\end{itemize}

All algorithms were implemented using operations over matrices and vectors similarly to how they are described in GraphBLAS-related materials.
We provide basic implementations without any optimizations such as the \emph{push-pull} optimization for BFS or SSSP.

\subsection{Interaction Nets Related Details}

The translation is almost ``as-is'' without any application of interaction-net-specific optimizations or code templates.

Inpla has no I/O, so data structures are generated. There is no error handling (hard to implement in a low-level language).

Vine supports reading. Execution is parallel, so separating concerns is difficult.
Error handling uses the standard \verb|Result| type from the Vine language.

Inpla does not support floats. 
Vine supports floats but without comparison (possibly due to problems with NaN comparison).
Therefore, weights in SSSP are represented as integers, not floats, in both Vine and Inpla implementations.