\section{Related Work}\label{sec:related_work}

 Proposed by Lafont, interaction nets can serve as a foundation for irregular parallel computations.
 Notable implementations include Inpla, HVM, and Vine. In OCaml~\cite{Huber_2025} (Comparison, Inpla fastest)

    Vine --- high-level language (Rust-like) + interaction-nets-based intermediate representation + parallel virtual machine.

    HVM --- high-level language (JS-like) + interaction-nets-based intermediate representation + parallel virtual machine (+GPGPU?).

    Inpla --- mature enough, demonstrates good performance and scalability. Interaction-nets-based language + parallel VM. Equipped with a high-level language.

No information is available on the evaluation or comparison of Vine and HVM.
Typical benchmarks for these systems include Fibonacci, Ackermann, sorting, and list traversal. 
No graph analysis algorithms while such algorithms looks grate candidates for interaction nets.
Other projects implement interaction nets in languages such as OCaml and Haskell.

Sparse linear algebra and graphs.
GraphBLAS API: C API, C++ API, implementations in Java, Julia, etc.
Discussion of GraphBLAS problems (“What did GraphBLAS Get Wrong?”, John Gilbert, UC Santa Barbara, GraphBLAS BoF at HPEC, September 2022).
SuiteSparse:GraphBLAS, LAGraph, GraphBLAS on GPUs. GraphBLAST, GraphLily (FPGA).
Manual kernel fusion. Attempts to use MLIR (GraphBLAS dialect).
Poor utilization on GPUs. Nontrivial challenge to create generic kernels for GPU and FPGA.

Special hardware for graph analysis.
"A Survey on Graph Processing Accelerators: Challenges and Opportunities", "Graph Processing on FPGAs: Taxonomy, Survey, Challenges"

Functional languages for parallel computing.
LIFT, AnyDSL, Futhark.

Parallel processors for functional programming languages.
Haflang/Heron, Reduceron.