\section{The Chomsky Normal Form}

The important part of our proof is that any context-free language can be described with grammar in Chomsky Normal Form (CNF) or, equally, any context-free grammar can be converted to the grammar in CNF which specifies the same language.
Let us recall the definition of CNF and the algorithm for CFG to CNF conversion.
\definition[Chomsky Normal Form]{Context-free grammar is in CNF if:
\begin{itemize}
\item start nonterminal does not occur in the right-hand side of rules,
\item all rules are of the form: $N_i \rightarrow t_i$, $N_i \rightarrow N_j N_k$ or $S \rightarrow \varepsilon$ where $N_i,N_j,N_k$ are nonterminals, $t_i$ is a terminal and $S$ is the start nonterminal.
\end{itemize} }

Transformation algorithm has the following steps.
\begin{enumerate}
\item Eliminate the start symbol from the right-hand sides of the rules.
\item Eliminate rules with nonsolary terminals.
\item Eliminate rules which right-hand side contains more than two nonterminals.
\item Delete $\varepsilon$-rules.
\item Eliminate unit rules.
\end{enumerate}

As far as Bar-Hillel theorem operates with arbitrary context-free languages, but the proof requires grammar in nCNF, it is necessary to implement a certified algorithm for conversion of arbitrary CFG to CNF.
We wanted to reuse existing proof of conversion of arbitrary context-free grammar to CNF.
We chose Smolka version which proves conversion to CNF correctness in the following way.
\begin{listing}[h]
    \begin{pyglist}[language=coq, numbers=none, numbersep=5pt]
  CNF!!!
    \end{pyglist}
    \caption{TODO}
    \label{lst:verbments1}
\end{listing}



