\section{Related Works}
\label{sec:rel-work}

There is a big number of works in mechanization of different parts of formal languages theory and certified implementations of parsing algorithms and algorithms for graph data base querying.
These works use different tools, such s Coq, Agda, Isabelle/HOL, and aimed to different problems such as theory mechanization or executable algorithm certification.
We discuss only small part which is close enough to the scope of this work.

\subsection{Formal Language Theory in Coq}

Huge amount of work was done by Ruy de Queiroz who formalize different parts of formal language theory, such as pumping lemma~\cite{ramos2015formalizationPumping}, context-free grammar simplification~\cite{ramos2015formalization} and closure properties~\cite{ramos2015formalizationClosure} in Coq.
The work on closure properties contains mechanization of such properties as closure under union, Klenee star, but it does not contains mechanization of intersection with regular language.
All these results are summarized in~\cite{ramos2016formalization}. 

Gert Smolka et.al. also provide big set of works on regular and context-free languages formalization in Coq~\cite{smolka2017regular,smolka2013regular,kaiser2012constructive,smolkaHofmann2016}.
The work~\cite{smolkaHofmann2016} contains certified transformation of arbitrary context-free grammar to Chomsky normal form which is required for our proof of the Bar-Hillel theorem. 
Initially we hope to use these both parts because Bar-Hillel theorem is about both context-free and regular languages, and it was the reason to choose results of Gert Smolka as base for our work.
But works on regular languages and on context-free languages are independent and we face with  problems of reusing and integration and in current proof we use only results on context-free languages.

\subsection{Formal Language Theory in Other Languages}

In the parallel with works in Coq there exist works on formal languages mechanization in other languages and tools such as Agda or Isabelle/HOL.

First part is works of Denis Firsov who implements in Agda some parts of formal language theory and parsing algorithms.
CYK parsing algorithm~\cite{firsov2014certified,firsov2016cfl} and Chomsky Normal Form~\cite{firsov2015certified}, and some results on regular languages~\cite{10.1007/978-3-319-03545-1_7}.

Another part is formal language theory mechanization in Isabelle/HOL~\cite{1885-16399,barthwal2010formalisation,10.1007/978-3-642-13824-9_11} by Aditi Bartwall and Michael Norrish.
This work contains basic definitions and big number of theoretical results, such as Chomsky normal form and Greibach normal form for context-free grammars. 
As an application of mechanized theory authors provide certified implementation of SLR parsing algorithm~\cite{10.1007/978-3-642-00590-9_12}.

\subsection{Certified Algorithms}

Additionally we want to mention some works on certified applied algorithms based on formal language theory.
Certification are required in different areas for various reasons and it is a reason to work on theory mechanization.

The first area where languages intersection may be applied is language constrained path querying in structured data (for example in graphs or XML).
There exist works on certification of core of XQuery~\cite{10.1007/978-3-642-25379-9_21}. 
XQuery is a W3C standard for path querying in XML, extended for graph querying. 
Another result is a work on certified Regular Datalog querying in Coq~\cite{certifiedPrologGraphQuerying}. 
Inspired by these results, our work may be a base for certified context-free path querying algorithm.

Another area which is growth fast is certified parsers and parser generators based on different algorithms for different language classes~\cite{10.1007/978-3-642-28869-2_20,bernardy2016certified,Lopes2016CertifiedDP,Gross2015ParsingPA}. 
