\documentclass[sigconf]{acmart}

\usepackage{calc}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{pstricks}

\makeatletter
\newcommand*{\da@rightarrow}{\mathchar"0\hexnumber@\symAMSa 4B }
\newcommand*{\da@leftarrow}{\mathchar"0\hexnumber@\symAMSa 4C }
\newcommand*{\xdashrightarrow}[2][]{%
  \mathrel{%
    \mathpalette{\da@xarrow{#1}{#2}{}\da@rightarrow{\,}{}}{}%
  }%
}
\newcommand{\xdashleftarrow}[2][]{%
  \mathrel{%
    \mathpalette{\da@xarrow{#1}{#2}\da@leftarrow{}{}{\,}}{}%
  }%
}
\newcommand*{\da@xarrow}[7]{%
  % #1: below
  % #2: above
  % #3: arrow left
  % #4: arrow right
  % #5: space left
  % #6: space right
  % #7: math style
  \sbox0{$\ifx#7\scriptstyle\scriptscriptstyle\else\scriptstyle\fi#5#1#6\m@th$}%
  \sbox2{$\ifx#7\scriptstyle\scriptscriptstyle\else\scriptstyle\fi#5#2#6\m@th$}%
  \sbox4{$#7\dabar@\m@th$}%
  \dimen@=\wd0 %
  \ifdim\wd2 >\dimen@
    \dimen@=\wd2 %
  \fi
  \count@=2 %
  \def\da@bars{\dabar@\dabar@}%
  \@whiledim\count@\wd4<\dimen@\do{%
    \advance\count@\@ne
    \expandafter\def\expandafter\da@bars\expandafter{%
      \da@bars
      \dabar@
    }%
  }%
  \mathrel{#3}%
  \mathrel{%
    \mathop{\da@bars}\limits
    \ifx\\#1\\%
    \else
      _{\copy0}%
    \fi
    \ifx\\#2\\%
    \else
      ^{\copy2}%
    \fi
  }%
  \mathrel{#4}%
}
\makeatother

\copyrightyear{2019}
\acmYear{2019}
\setcopyright{acmcopyright}
\acmConference[CEE-SECR'19]{Central and Eastern European Software Engineering Conference Russia}{November 14--15(16), 2019}{St.Petersburg, Russia}
%\acmBooktitle{2nd Joint International Workshop on Graph Data Management
%Experiences \& Systems (GRADES) and Network Data Analytics (NDA)
%(GRADES-NDA'19), June 30-July 5, 2019, Amsterdam, Netherlands}
%\acmPrice{15.00}
%\acmDOI{10.1145/3327964.3328503}
%\acmISBN{978-1-4503-6789-9/19/06}

\begin{document}

\title[IDEs-Friendly Interprocedural Analyser]{IDEs-Friendly Interprocedural Analyser}

\author{Ilya Nozhkin}
\affiliation{
  \institution{Saint Petersburg State University}
  \streetaddress{7/9 Universitetskaya nab.}
  \city{St. Petersburg}
  \country{Russia}
  \postcode{199034}
}
\email{nozhkin.ii@gmail.com}

\author{Semyon Grigorev}
\orcid{0000-0002-7966-0698}
\affiliation{
  \institution{Saint Petersburg State University}
  \streetaddress{7/9 Universitetskaya nab.}
  \city{St. Petersburg}
  \country{Russia}
  \postcode{199034}
}
\affiliation{
  \institution{JetBrains Research}
  \streetaddress{Universitetskaya nab., 7-9-11/5A}
  \city{St. Petersburg}
  \country{Russia}
  \postcode{199034}
}
\email{s.v.grigoriev@spbu.ru}
\email{semen.grigorev@jetbrains.com}


\begin{abstract}
We propose an extensible framework for interprocedural static code analysis implementation.
Our solution is based on CFL-reachability: analysis is formulated in terms of context-free constrained reachability in the interprocedural graph.
Extensible architecture allows one to implement new analysis and integrate it into the IDE of choice or static code analysis tool.
To demonstrate the abilities of our solution, we implement the plugin which provides basic taint analysis and variable flow analysis upon ReSharper infrastructure.
We demonstrate its applicability for real-world problems.
\end{abstract}


\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10010940.10010992.10010998.10011000</concept_id>
<concept_desc>Software and its engineering~Automated static analysis</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011006.10011066.10011069</concept_id>
<concept_desc>Software and its engineering~Integrated and visual development environments</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011074.10011099.10011102</concept_id>
<concept_desc>Software and its engineering~Software defect analysis</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10010940.10011003.10011114</concept_id>
<concept_desc>Software and its engineering~Software safety</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10003752.10003766.10003771</concept_id>
<concept_desc>Theory of computation~Grammars and context-free languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Automated static analysis}
\ccsdesc[500]{Software and its engineering~Integrated and visual development environments}
\ccsdesc[500]{Software and its engineering~Software defect analysis}
\ccsdesc[300]{Software and its engineering~Software safety}
\ccsdesc[300]{Theory of computation~Grammars and context-free languages}

\keywords{Static code analysis, interprocedural analysis, CFL-reachability, taint analysis, IDE, plugin, context-free languages, PDA}

%\acmBadgeR{artifacts_available}

\maketitle

\section{Introduction}
\input{Introduction.tex}

\section{Analysis definition}
\input{Analysis.tex}

\input{Solution.tex}

\section{Evaluation}
\input{Evaluation.tex}

\input{Conclusion.tex}

\bibliographystyle{ACM-Reference-Format}
\bibliography{IDE_Friendly_Interprocedural_Analyser}

\end{document}
