% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads,table]{llncs}
%
\usepackage{graphicx}

% For algorithmic attachments and graphics
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage[noend]{algpseudocode}
\usepackage{subcaption}
\captionsetup{compatibility=false}

\usepackage{tikz}
\usepackage{dsfont}
\usepackage{multicol, multirow}
\usepackage{textcomp}
%\usepackage{booktabs}
\usetikzlibrary{fit,calc,automata,positioning}

\usepackage{kbordermatrix}
\usepackage{extarrows}
\usepackage{balance}
\usepackage{colortbl}
\usepackage{mathtools}
\usepackage{verbatim}
\definecolor{lightgray}{gray}{0.9}



% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\newtheorem{mytheorem}{Theorem}
\newtheorem{myproposition}{Proposition}

\newcommand{\ltz}{$< 1$}

\algtext*{EndWhile}% Remove "end while" text
\algtext*{EndIf}% Remove "end if" text
\algtext*{EndFor}% Remove "end for" text
\algtext*{EndFunction}% Remove "end function" text

\begin{document}
%
\title{Multiple Context-Free Path Querying by Matrix Multiplication\thanks{The research was supported by the Russian Science Foundation, grant \textnumero 18-11-00100}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Ilya Epelbaum\inst{1} \and
Rustam Azimov\inst{1,2} \and
Semyon~Grigorev\inst{1,2}\orcidID{0000-0002-7966-0698}}
%
\authorrunning{I. Epelbaum et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{St.Petersburg State University, 7/9 Universitetskaya nab., St. Petersburg, Russia, 199034 \\
\email{iliyepelbaun@gmail.com, rustam.azimov19021995@gmail.com, s.v.grigoriev@spbu.ru} \and
JetBrains Research, Primorskiy prospekt 68-70, Building 1, St. Petersburg, Russia, 197374}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Many graph analysis problems can be formulated as formal language-constrained path querying problems where the formal languages are used as constraints for
navigational path queries. Recently, the context-free language (CFL) reachability formulation has become very popular and can be used in many areas, for example, querying graph databases, RDF analysis, static code analysis, biological data analysis, graph segmentation. However, the generative capacity of context-free grammars (CFGs) is too weak to generate some complex queries, for example, from natural languages, and the various extensions of CFGs have been proposed. Multiple context-free grammar (MCFG) is one of such extensions of CFGs. Despite the fact that, to the best of our knowledge, there is no algorithm for MCFL-reachability, this problem is known to be decidable. The creation of the MCFL-reachability algorithms allows one to use more complex graph queries that may find application in various areas, for example, in static code analysis. In this paper, we provide the first MCFL-reachability algorithm that also is based on Boolean matrix multiplication. Thus, the proposed algorithm can be implemented by using high-performance libraries and modern parallel hardware.


\keywords{Multiple context-free path querying \and Graph database \and RDF \and Multiple context-free grammars \and Boolean matrix multiplication \and GraphBLAS API.}
\end{abstract}
%
%
%

\input{introduction}
\input{preliminaries}
\input{algorithm}
\input{evaluation}
\input{conclusion}

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
 \bibliographystyle{splncs04}
 \bibliography{mcfg}
\end{document}