\section{The platform}

Any language can be used both as string-embedded or host language. For each combination it could be necessary to solve different tasks: from errors detection to additional information extraction and metrics computation. As there are substantial number of languages combinations and different types of tasks, our research is not aiming to create a tool that handles all possible scenarios. The main goal is to create the platform that simplifies the process of building endpoint tools for the certain languages and tasks. The similar approach can be seen in the tools for compilers development. Usually such tools include lexical and syntax generators and utility functions libraries to simplify the process of  specific compiler creating.

Being complex and self-contained task, the host language analysis is assumed to be performed by some external tool. Moreover, this tool is assumed to return a source code's AST which contains the information needed for further steps as a result of it's work.

The platform is to perform the whole process of string-embedded languages handling. It includes the following steps:
\begin{itemize}
\item dynamically generated expressions can have a set of possible values so the approximation of the set is to be constructed;
\item lexical analysis is to be applied to the FSA that has been obtained after approximation;
\item syntax analysis is to be performed resulting in the parse forest;
\item parse forest analysis and semantics calculation is to be performed.
\end{itemize}

\input{04_01_Architecture}

\input{04_02_Approximation}
\input{04_03_Lexing_and_parsing}
\input{04_04_Semantics}