\section{Conclusion and future work}
In this paper, we shown how the context-free path query evaluation w.r.t. relational and single-path query semantics can be reduced to the calculation of matrix transitive closure. Also, we provided a formal proof of the correctness of the proposed reduction. In addition, we introduced an algorithm for computing this transitive closure, which allows us to efficiently apply GPGPU computing techniques. Finally, we shown the practical applicability of the proposed algorithm by running different implementations of our algorithm on real-world data.

We can identify several open problems for further research. In this paper we have considered only two semantics of context-free path querying but there are other important semantics, such as \textit{all-path} semantics~\cite{hellingsPathQuerying} which requires to present all paths for all triples $(A,m,n)$. Context-free path querying implemented with algorithm~\cite{GLL} can answer the queries in all-path query semantics by constructing a parse forest. It is possible to construct a parse forest for a linear input by matrix multiplication~\cite{okhotin_cyk}. Whether it is possible to generalize this approach for a graph input is an open question.

In our algorithm, we calculate the matrix transitive closure naively, but there are algorithms for the transitive closure calculation, which are asymptotically more efficient. Therefore, the question is whether it is possible to apply these algorithms for the matrix transitive closure calculation to the problem of context-free path querying.

Also, there are Boolean grammars~\cite{okhotinBoolean}, which have more expressive power than context-free grammars. Boolean path querying is an undecidable problem~\cite{hellingsRelational} but our algorithm can be trivially generalized to work on boolean grammars because parsing with boolean grammars can be expressed by matrix multiplication~\cite{okhotin_cyk}. It is not clear what a result of our algorithm applied to Boolean grammars would look like. Our hypothesis is that it would produce the upper approximation of a solution.

From a practical point of view, matrix multiplication in the main loop of the proposed algorithm may be performed on different GPGPU independently. It can help to utilize the power of multi-GPU systems and increase the performance of context-free path querying.

There is an algorithm~\cite{apspGPU} for transitive closure calculation on directed graphs which generalized to handle graph sizes inherently larger then the DRAM memory available on the GPU. Therefore, the question is whether it is possible to apply this approach to the matrix transitive closure calculation in the problem of context-free path querying.