\documentclass[sigconf,edbt]{acmart-edbt2018}

\usepackage{booktabs} % For formal tables
\usepackage{graphicx}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)

\usepackage{caption}
\usepackage{subcaption}
%\usepackage{gnuplottex}
\usepackage{tikz}
\usepackage{mathtools}

\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{textcomp}

\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{algorithmicx}


% Copyright
\setcopyright{rightsretained}

% DOI
\acmDOI{}

% ISBN
\acmISBN{978-3-89318-078-3}

%Conference
\acmConference[EDBT 2018]{21st International Conference on Extending Database Technology (EDBT)}{March 26-29, 2018}{Vienna, Austria} 
\acmYear{2018}

\settopmatter{printacmref=false, printccs=false, printfolios=false}

\pagestyle{plain} % removes running headers


\begin{document}
	
\newtheorem{mytheorem}{Theorem}
%\newtheorem{lemma}{Lemma}
\newtheorem{mydef}{Definition}

\algrenewcommand\algorithmicindent{0.5em}
\algnewcommand\algorithmicswitch{\textbf{switch}}
\algnewcommand\algorithmiccase{\textbf{case}}
\algnewcommand\algorithmicassert{\texttt{assert}}
\algnewcommand\Assert[1]{\State \algorithmicassert(#1)}
% New "environments"
\algdef{SE}[SWITCH]{Switch}{EndSwitch}[1]{\algorithmicswitch\ #1\ \algorithmicdo}{\algorithmicend\ \algorithmicswitch}
\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1}{\algorithmicend\ \algorithmiccase}

\algtext*{EndSwitch}
\algtext*{EndCase}
\algtext*{EndWhile}% Remove "end while" text
\algtext*{EndIf}% Remove "end if" text
\algtext*{EndFor}% Remove "end for" text
\algtext*{EndFunction}% Remove "end function" text

\newif\ifboldnumber
\newcommand{\boldnext}{\global\boldnumbertrue}


\title{Context-Free Path Querying by Matrix Multiplication}

\author{Rustam Azimov}
\affiliation{%
  \institution{Saint Petersburg State University}
  \streetaddress{7/9 Universitetskaya nab.}
  \city{St. Petersburg} 
  \state{} 
  \postcode{199034 Russia}
}
\email{rustam.azimov19021995@gmail.com}

\author{Semyon Grigorev}
\affiliation{%
	\institution{Saint Petersburg State University}
	\streetaddress{7/9 Universitetskaya nab.}
	\city{St. Petersburg}
	\state{}  
	\postcode{199034 Russia}
}
\email{Semen.Grigorev@jetbrains.com}

% The default list of authors is too long for headers}
% \renewcommand{\shortauthors}{B. Trovato et al.}
\renewcommand{\shortauthors}{}


\begin{abstract}
Graph data model is widely used in many areas, for example, bioinformatics, graph databases, RDF. One of the most common graph queries are navigational queries. The result of query evaluation are implicit relations between nodes of the graph, i.e. paths in the graph. A natural way to specify these relations is by specifying paths using formal grammars over edge labels. The answer to the context-free path queries in this approach is usually a set of triples $(A, m, n)$ such that there is a path from node $m$ to node $n$ whose labeling is derived from a non-terminal $A$ of the given context-free grammar. This type of queries is evaluated using the relational query semantics. There is a number of algorithms for query evaluation which use such semantics but they have computational problems with big data. One of the most common technique for efficient big data processing is GPGPU, but these algorithms do not allow to use this technique effectively. In this paper we propose an algorithm for context-free path query evaluation which use relational query semantics and is based on matrix operations which allows to speed up computations by means of GPGPU.
\end{abstract}

%\keywords{}

\maketitle

\section{Introduction}
Graph data model is widely used in many areas, for example, bioinformatics~\cite{Bio}, graph databases~\cite{graphDB}, RDF~\cite{RDF}. In these areas, it is often required to process large graphs. Most common among graph queries are navigational queries. The result of query evaluation is implicit relations between nodes of the graph, i.e. paths in the graph. A natural way to specify these relations is by specifying paths using formal grammars (regular expressions, context-free grammars) over edge labels. Context-free grammars are actively used in graphs queries because of the limited expressive power of regular expressions. 

The result of context-free path query evaluation is usually a set of triples $(A, m, n)$ such that there is a path from node $m$ to node $n$ whose labeling is derived from a non-terminal $A$ of the given context-free grammar. This type of queries is evaluated using the \textit{relational query semantics}~\cite{hellingsRelational}. There is a number of algorithms for context-free path query evaluation using this semantics~\cite{GLL, hellingsRelational, RDF}.

Existing algorithms for ontext-free path query evaluation using this semantics have computational problems with big data. One of the most common technique for efficient big data processing is \textit{GPGPU} (General-Purpose computing on Graphics Processing Units), but these algorithms do not allow to use this technique effectively. The algorithms for context-free recognizing had a similar problem until Valiant~\cite{valiant} proposed a parsing algorithm that computes a recognition table by computing a matrix transitive closure. Thus, the active use of matrix operations (such as matrix multiplication) in the process of computing a transitive closure makes it possible to effectively apply GPGPU computing techniques~\cite{matricesOnGPGPU}.

Therefore the question is whether it is possible to create an algorithm for context-free path query evaluation using the relational query semantics which allows to speed up computations with GPGPU by using the matrix operations.

The main contribution of this paper can be summarized as follows:
\begin{itemize}
	\item We show how the context-free path query evaluation using the relational query semantics can be reduced to the calculation of the matrix transitive closure.
	\item We provide a formal proof of correctness of the proposed reduction.
	\item We introduce an algorithm for context-free path query evaluation which use the relational query semantics and is based on matrix operations which allows to speed up computations by means of GPGPU.
\end{itemize}

\input{preliminaries.tex}
\input{related_works.tex}
\input{algorithm.tex}
\input{evaluation.tex}
\input{conclusion.tex}
\input{acknowledgments.tex}

\bibliographystyle{ACM-Reference-Format}
\bibliography{graphparsing} 

\end{document}
