\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables


% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}


% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[FTfJP'18]{Formal Techniques for Java-like Programs}{July 2018}{Amsterdam, Netherlands}
\acmYear{2018}
\copyrightyear{2018}


\acmArticle{4}
\acmPrice{15.00}

% These commands are optional
%\acmBooktitle{Transactions of the ACM Woodstock conference}
%\editor{Jennifer B. Sartor}
%\editor{Theo D'Hondt}
%\editor{Wolfgang De Meuter}


\begin{document}
\title{Context-Free Shuffle Languages Parsing via Boolean Satisfiability Problem Solving}
%\titlenote{Produces the permission block, and
%   copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as
%  \texttt{acmart.pdf} document}


\author{Artem Gorokhov}
\affiliation{
  \institution{Saint Petersburg State University}
  \streetaddress{7/9 Universitetskaya nab.}
  \city{Saint Petersburg}
  \state{Russia}
  \postcode{199034}
}
\email{gorohov.art@gmail.com}

\author{Semyon Grigorev}
\affiliation{
    \institution{Saint Petersburg State University}
    \streetaddress{7/9 Universitetskaya nab.}
    \city{Saint Petersburg}
    \state{Russia}
    \postcode{199034}
}
\email{s.v.grigoriev@spbu.ru}

% The default list of authors is too long for headers.
%\renewcommand{\shortauthors}{B. Trovato et al.}


\begin{abstract}
    Verification of concurrent systems is important and nontrivial problem.
    One of directions in this area is modeling of sequential subsystems with push-down automata (PDA) and investigating its communication.
    PDA is equal to context-free languages and ``communication'' may be expressed as shuffle of them.
    In this paper we consider the problem of concurrent programs' model checking from the side of context-free languages shuffle:
    in order to check correctness of system we should check emptiness of intersection of shuffled context-free languages (which describe behavior of the system) with regular language (which describe set of ``bad'' behaviors).
    Even in simple case, when regular language is finite, it leads to NP-complete problem and we show how it can be solved by using SAT-solvers.
    Our reduction is very native and use classical parsing techniques, such as Shared Packed Parse Forest and Generalized LL parsing algorithm, and some ideas from Context-Free Language reachability framework.
    We do not propose solution for arbitrary regular language (existence of which looks an open problem) but we show a some possible directions of research and hope that ever for restricted case proposed solution may be useful.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%

%\ccsdesc[500]{Computer systems organization~Embedded systems}
%\ccsdesc[300]{Computer systems organization~Redundancy}
%\ccsdesc{Computer systems organization~Robotics}
%\ccsdesc[100]{Networks~Network reliability}
\ccsdesc[500]{Theory of computation~Grammars and context-free languages}
\ccsdesc[500]{Software and its engineering~Software reliability}
\keywords{Model checking, static analysis, concurrency, shuffle, formal languages, language intersection, context-free languages}


\maketitle

\input{body}

\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-bibliography}

\end{document}
