% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration
%\usepackage[T2A]{fontenc} 
%\usepackage[utf8]{inputenc}
%\usepackage[english,russian]{babel}
\usepackage{graphicx}
\usepackage{indentfirst}
\usepackage{hyperref}
\usepackage{verbatim}
%
\begin{document}

\sloppy
%
%\frontmatter          % for the preliminaries
%
%\pagestyle{headings}  % switches on printing of running heads
%\addtocmark{Hamiltonian Mechanics} % additional mark in the TOC
%
%
%\tableofcontents
%
%\mainmatter              % start of the contributions
%
\title{From Abstract Parsing to Abstract Translation}
%
%\titlerunning{Hamiltonian Mechanics}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Semyon Grigorev\inst{1} 
\and Iakov Kirilenko\inst{2}
}
%
%\authorrunning{Ivar Ekeland et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Semyon Grigorev
, Iakov Kirilenko
}


\institute{St. Petersburg State University \\
198504, Universitetsky prospekt 28, Peterhof, St. Petersburg, Russia. \\
\email{rsdpisuy@gmail.com}
\and
St. Petersburg State University \\
198504, Universitetsky prospekt 28, Peterhof, St. Petersburg, Russia. \\
\email{jake@math.spbu.ru}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
String-embedded language transformation is one of the problems which can be faced during 
database and information system migration. The conventional solution which
is provided by a number of tools is based on run-time translation. We present a static 
\emph{abstract translation} approach which orignates from the \emph{abstract parsing} 
technique~\cite{AbstrParsing} initially developed for syntax analysis of string-embedded 
languages. We present abstract translation algorithm and some optimization techniques, and
discuss the results of its evaluation on a real-world industrial application.

%There are two approaches to 
%solve this problem: static translation or translation at run time. Our research deals 
%with the first one since solution for embedded language static translation problem do 
%not still exists. There is an abstract parsing algorithm~\cite{} for parsing of 
%string-embedded language. Parsing is close problem and in this article we propose abstract 
%parsing algorithm modification for application to string-embedded language translation. 
%We describe modifications of the basic algorithm which allow to translate embedded languages. 
%An algorithm of parsing forest minimization is also presented. We discuss some issues of 
%real-world dynamic SQL queries static translation and provide possible solutions. Finally, 
%we present results of implemented algorithm evaluation.

\keywords{Embedded languages, two-staged languages, string-embedded languages, abstract 
parsing,  abstract translation, LALR, dynamic SQL}
\end{abstract}

\input{introduction}
\input{dags}
\input{abstract_translation}
\input{evaluation}
\input{conclusion}

%
% ---- Bibliography ----
%
\begin{thebibliography}{}
  
\bibitem{RelaxedLALR}
Ефимов А.А., Кириленко Я.А. Построение ослабленного LALR-транслятора на основе анализа грамматики на избыточность // Системное программирование. Т. 4, вып. 1, 2009. С. 79–103.  

\bibitem{mart}
Мартыненко Б.К. Языки и трансляции. Издательство Санкт-Петербургского университета, 2008. 257 с. 

\bibitem{TiunovaUIInt}
Мосиенко М.А., Тиунова А.Е. Интеграция программной логики с пользовательским интерфейсом при реинжиниринге приложений // Системное программирование. Т. 1, вып. 1, 2004. С. 199–224.

\bibitem{NetDbTransform}
Трошин С.Л. Преобразование сетевых баз данных в реляционные: задачи и подходы // Системное программирование. Т. 1, вып. 1. 2004. С. 282–310.

\bibitem{OpenSystemsDBMS}
Shapot M., Popov E. Database reengineeing // Open Systems.DBMS. Number 4. 2004.    

\bibitem{ALVOR1}
Annamaa A., Breslav A., Kabanov J. e.a. An interactive tool for analyzing embedded SQL queries. Programming Languages and Systems. LNCS, vol. 6461. Springer: Berlin; Heidelberg. 2010. P. 131–138.

\bibitem{ALVOR2}
Annamaa A., Breslav A., Vene V. Using abstract lexical analysis and parsing to detect errors in string-embedded DSL statements // Proceedings of the 22nd Nordic Workshop on Programming Theory. Marina Walden and Luigia Petre, editors. 2010. P. 20-22.

\bibitem{StringExpr}
Aske Simon Christensen, Mller A., Michael I. Schwartzbach. Precise analysis of string expressions // Proc. 10th International Static Analysis Symposium (SAS), Vol. 2694 of LNCS. Springer-Verlag: Berlin; Heidelberg, June, 2003. P. 1–18.

\bibitem{Grune}
Grune D., Ceriel J. H. Jacobs. Parsing techniques: a practical guide. Ellis Horwood, Upper Saddle River, NJ, USA, 1990. P. 322.

\bibitem{SAofStrVal}
Costantini G., Ferrara P., Cortesi F. Static analysis of string values // Proceedings of the 13th international conference on Formal methods and software engineering, ICFEM’11. Springer-Verlag: Berlin; Heidelberg, 2011. P. 505–521.

\bibitem{ISO}
ISO. ISO/IEC 9075:1992: Title: Information technology — Database languages — SQL. 1992. P. 668.

\bibitem{JSA}
Java String Analyzer. URL: \href{http://www.brics.dk/JSA/}{http://www.brics.dk/JSA/}

\bibitem{AbstrParsing}
Kyung-Goo Doh, Hyunha Kim, David A. Schmidt. Abstract parsing: Static analysis of dynamically generated string output using LR-parsing technology // Proceedings of the 16th International Symposium on Static Analysis, SAS’09. Springer-Verlag: Berlin; Heidelberg, 2009. P. 256–272.

\bibitem{PHPSA}
PHP String Analyzer. URL: \href{http://www.score.is.tsukuba.ac.jp/~minamide/phpsa/}{http://www.score.is.tsukuba.ac.jp/$\sim$minamide/phpsa/}

\bibitem{PLSQL}
PL/SQL Developer. URL: \href{http://www.allroundautomations.com/plsqldev.html}{http://www.allroundautomations.com/plsqldev.html}

\bibitem{SQLWays}
SQL Ways. URL: \href{http://www.ispirer.com/products}{http://www.ispirer.com/products}

\bibitem{SwissSQL}
SwissSQL. URL: \href{http://www.swissql.com/}{http://www.swissql.com/}

\bibitem{SAForInject}
Xiang Fu, Xin Lu, Peltsverger B. e.a. A static analysis framework for detecting SQL injection vulnerabilities // Proceedings of the 31st Annual International Computer Software and Applications Conference. Vol. 01, COMPSAC’07, Washington, DC, USA, IEEE Computer Society, 2007. P. 87–96.

\end{thebibliography}


%\clearpage
%\addtocmark[2]{Author Index} % additional numbered TOC entry
%\renewcommand{\indexname}{Author Index}
%\printindex
%\clearpage
%\addtocmark[2]{Subject Index} % additional numbered TOC entry
%\markboth{Subject Index}{Subject Index}
%\renewcommand{\indexname}{Subject Index}
%\input{subjidx.ind}
\end{document}
