\section{Introduction}

% Commented references
% CuSha~\cite{10.1145/2600212.2600227}
% MapGraph~\cite{10.1145/2621934.2621936/MapGraph}
% Medusa~\cite{6497047/Medusa}
% Gunrock~\cite{7967137}

Scalable high-performance graph analysis is a challenge.
CuSha~\cite{10.1145/2600212.2600227} and Gunrock~\cite{7967137} show that the utilization of GPUs can improve the performance of graph analysis. But the low flexibility and high complexity of API are problems of these solutions. 
A linear algebra based model, formalized in GraphBLAS~\cite{7761646}, provides a user-friendly API. 
While the reference CPU-based implementation for this API, SuiteSparse~\cite{10.1145/3322125}, demonstrates good performance on real-world tasks, GPU-based implementation is challenging due to required operations generalization, data sparsity, and hardware programming complexity. 
Such well-known libraries as cuSPARSE, clSPARSE, bhSPARSE cannot be reused, since almost all of them are specified for operations over floats. 
GraphBLAST~\cite{yang2019graphblast} and GBTL~\cite{7529957} show promising GPU performance of GraphBLAS-based graph analysis solutions. 
But these solutions are not portable because they are based on Nvidia Cuda. 
Portable solution is a challenge due to intricate GPU programming. 
To address this problem, we developed the GraphBLAS-inspired \textit{Spla}\footnote{Source code and project page: \url{github.com/SparseLinearAlgebra/spla}} library, which features portable OpenCL-based acceleration and shows performance comparable with GraphBLAST, achieving speedup of up to 36 times.