\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{caption}
\usepackage{subcaption}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

    
\begin{document}

\title{Spla: Generalized Sparse Linear Algebra Library with Vendor-Agnostic GPUs Acceleration}

\author{\IEEEauthorblockN{1\textsuperscript{st} Egor Orachev}
\IEEEauthorblockA{\textit{St. Petersburg State University} \\
St. Petersburg, Russia \\
egor.orachev@gmail.com \\
0000-0002-0424-4059}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Semyon Grigorev}
\IEEEauthorblockA{\textit{St. Petersburg State University} \\
St. Petersburg, Russia \\
s.v.grigoriev@spbu.ru \\
0000-0002-7966-0698}
}

\maketitle

\begin{abstract}
    Scalable high-performance graph analysis is an actual nontrivial challenge. Usage of sparse linear algebra operations as building blocks for graph analysis algorithms, which is a core idea of GraphBLAS standard, is a promising way to attack it.
    While it is known that sparse linear algebra operations can be efficiently implemented on GPU, full GraphBLAS implementation on GPU is a nontrivial task that is almost solved by GraphBLAST project. Though it is shown that utilization of GPUs for GraphBLAS implementation significantly improves performance, portability problem is not solved yet: GraphBLAST uses Nvidia Cuda stack.
    In this work we propose Spla library which aims to solve this problem using OpenCL API for vendor-agnostic GPUs computations.
    Evaluation shows that the proposed solution demonstrates performance comparable with GraphBLAST, outperforming it up to 36 times in some cases, remaining portable across different GPUs vendors. However there is still a space for further optimizations.
\end{abstract}

\begin{IEEEkeywords}
graphs, algorithms, graph analysis, sparse linear algebra, GraphBLAS, GPGPU, OpenCL
\end{IEEEkeywords}

\input{introduction}
\input{solution_description}
\input{evaluation}
\input{conclusion}

\bibliographystyle{IEEEtran}
\bibliography{spla_hpec}

\end{document}
