\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage{multirow}
\usepackage{listings}   

\lstset{
  frame=none,
  language=scala,
  basicstyle={\small\ttfamily}
}

\sloppy

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}


% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[GRADES-NDA'18]{Joint International Workshop on Graph Data Management Experiences \& 
Systems (GRADES) and Network Data Analytics (NDA) 2018}{June 2018}{Houston, Texas USA}
\acmYear{2018}
\copyrightyear{2018}


\acmArticle{4}
\acmPrice{15.00}

% These commands are optional
%\acmBooktitle{Transactions of the ACM Woodstock conference}
%\editor{Jennifer B. Sartor}
%\editor{Theo D'Hondt}
%\editor{Wolfgang De Meuter}


\begin{document}
\title{Parser-Combinators for Contex-Free Path Querying}
\titlenote{This work is supported by grant from JetBrains Research.}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as
%  \texttt{acmart.pdf} document}


\author{Sophia Smolina}
\affiliation{%
  \institution{Electrotechnical University}
  \streetaddress{ul. Professora Popova 5,}
  \city{St. Petersburg}
  \country{Russia}
  \postcode{197376}
}
\email{sofysmol@gmail.com}

\author{Ekaterina Verbitskaia}
\affiliation{%
  \institution{Saint Petersburg State University}
  \streetaddress{7/9 Universitetskaya nab.}
  \city{St. Petersburg}
  \country{Russia}
  \postcode{199034}
}
\email{kajigor@gmail.com}

\author{Ilya Kirillov}
\affiliation{%
  \institution{Saint Petersburg State University}
  \streetaddress{7/9 Universitetskaya nab.}
  \city{St. Petersburg}
  \country{Russia}
  \postcode{199034}
}
\email{kirillov.ilija@gmail.com}

\author{Semyon Grigorev}
\orcid{0000-0002-7966-0698}
\affiliation{%
  \institution{Saint Petersburg State University}
  \streetaddress{7/9 Universitetskaya nab.}
  \city{St. Petersburg}
  \country{Russia}
  \postcode{199034}
}
\email{semen.grigorev@jetbrains.com}

% The default list of authors is too long for headers.
\renewcommand{\shortauthors}{Smolina et al.}

\begin{abstract}
Context-free path queries (CFPQs) express path queries in terms of context-free grammars and are applied in biological data analysis and static code analysis. 
A transparent integration of a domain-specific language for specification of CFPQs into a general-purpose programming languge as well as static checking of errors in queries may greatly simplify the development of applications utilizing CFPQs.  
Such techniques as LINQ and ORM can be used for the integration, but they have issues with flexibility: query decomposition and reusing of subqueries are a challenge.

Parser combinators is a native and flexible language specification mechanism. 
An adaptation of this technique can be the desired means of integration. 
Conventional parser combinators process linear input and only the Trails library is known to apply this technique for the path querying.
Trails suffers the common parser combinators issue: it does not support left-recursive grammars and also experiences problems in cycles handling.
We demostrate that it is possible to create general parser combinators for CFPQ which support arbitrary context-free grammars and arbitrary input graphs.
We implement a library of such parser combinators and show that it is applicable for realistic problems.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002951.10002952.10002953.10010146</concept_id>
<concept_desc>Information systems~Graph-based database models</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10002951.10002952.10003197.10010825</concept_id>
<concept_desc>Information systems~Query languages for non-relational engines</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011006.10011008.10011009.10011012</concept_id>
<concept_desc>Software and its engineering~Functional languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10003752.10003766.10003771</concept_id>
<concept_desc>Theory of computation~Grammars and context-free languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Graph-based database models}
\ccsdesc[500]{Information systems~Query languages for non-relational engines}
\ccsdesc[300]{Software and its engineering~Functional languages}
\ccsdesc[300]{Theory of computation~Grammars and context-free languages}

\keywords{Graph data bases, Language-constrained path problem, Context-Free path querying, Parser Combinators, Domain Specific Language, Generalized LL, GLL, Neo4J, Scala}

\maketitle

\input{introduction}
\input{combinators}
\input{evaluation}
\input{conclusion}

%\section*{Acknowledgments}


\bibliographystyle{ACM-Reference-Format}
\bibliography{combinators_for_graph_querying}

\end{document}
