\section{Introduction}


Language-constrained path querying~\cite{barrett2000formal} is one of the techniques for graph navigation querying.
This technique allows one to use formal languages as constraints on paths in edge-labeled graphs: path satisfies constraints if labels along it form a word from the specified language.

The utilization of regular languages as constraints, or \textit{Regular Path Querying} (RPQ), is most well-studied and widely spread.
Different aspects of RPQs are actively studied in graph databases~\cite{10.1145/2463664.2465216, 10.1145/3104031,10.1145/2850413}, and support of regular constraints is implemented in such popular query languages as PGQL~\cite{10.1145/2960414.2960421}, or SPARQL\footnote{Specification of regular constraints in SPARQL property paths: \url{https://www.w3.org/TR/sparql11-property-paths/}. Access date: 07.07.2020.}~\cite{10.1007/978-3-319-25007-6_1} (property paths).
Even that, the improvement of RPQ algorithms efficiency on huge graphs is an actual problem nowadays, and new algorithms and solutions are being created~\cite{Wang2019,10.1145/2949689.2949711}.

At the same time, the utilization of more powerful languages, namely context-free languages, gains popularity in the last few years. 
\textit{Context-Free Path Querying} problem (CFPQ) was introduced by Mihalis Yannakakis in 1990 in~\cite{Yannakakis}.
A number of different algorithms were proposed since that time, but recently, in~\cite{Kuijpers:2019:ESC:3335783.3335791} Jochem Kuijpers et al. show that state-of-the-art CFPQ algorithms are not performant enough to be used in practice.
This fact motivates us to find new algorithms for CFPQ.

One of the promising ways to achieve high-performance solutions for graph analysis problems is to reduce graph problems to linear algebra operations.
This way, the description of basic linear algebra primitives, the GraphBLAS~\cite{7761646} API, was proposed.
Solutions that use libraries that implement this API, such as SuiteSparce~\cite{10.1145/3322125} and CombBLAS~\cite{10.1177/1094342011403516}, show that reduction to linear algebra is a way to utilize high-performance parallel and distributed computations for graph analysis.

Rustam Azimov in~\cite{Azimov:2018:CPQ:3210259.3210264} shows how to reduce CFPQ to matrix multiplication.
Late, in~\cite{Mishin:2019:ECP:3327964.3328503} and~\cite{10.1145/3398682.3399163}, it was shown that utilization of appropriate libraries for linear algebra for Azimov's algorithm implementation allows one to get practical solution for CFPQ.
However Azimov's algorithm requires transforming the input grammar to Chomsky Normal Form, which leads to the grammar size increase, and hence worsens performance especially for regular queries and complex context-free queries.

To solve these problems, recently, an algorithm based on automata intersection was proposed~\cite{10.1007/978-3-030-54832-2_6}.
This algorithm is based on linear algebra and does not require the input grammar transformation.
In this work, we improve it.
First of all, we reduce the above mentioned solution to operations over Boolean matrices, thus simplify its description and implementation.
Also, we show that this algorithm is performant enough for regular queries, so it is a good candidate for integration with real-world query languages: we can use one algorithm to evaluate both regular and context-free queries. 

Moreover, we show that this algorithm opens the way to attack a long-standing problem whether there is a truly-subcubic $O(n^{3-\epsilon})$ CFPQ algorithm ~\cite{10.1145/1328438.1328460, Yannakakis}. 
The best-known result is an $O(n^3/\log{n})$ algorithm of Swarat Chaudhuri~\cite{10.1145/1328438.1328460}. 
Also, there are truly subcubic solutions using fast matrix multiplication for some fixed subclasses of context-free languages~\cite{8249039}. 
Unfortunately, this solutions cannot be generalized to arbitrary CFPQs. 
In this work, we show that incremental transitive closure is a bottleneck on the way to get subcubic time complexity for CFPQ.

To summarize, we make the following contributions in this paper.
\begin{enumerate}
	\item We rethink and improve the tensor-product-based algorithm for CFPQ of Orachev et al. ~\cite{10.1007/978-3-030-54832-2_6}. First of all, we reduce this algorithm to operations over Boolean matrices. This way all-path query semantics is handled. Notice that the previous matrix-based solution handles only single-path semantics. Also one can formulate query using both regular and context-free grammars. We prove the correctness and time complexity for the proposed algorithm.
	\item We demonstrate the interconnection between CFPQ and incremental transitive closure. We show that incremental transitive closure is a bottleneck on the way to get faster CFPQ algorithm for general case of arbitrary graphs as well as for special families of graphs, like planar graphs.
	\item We implement the described algorithm and evaluate it on real-world data. RPQ, CFPQ. Results show that !!!
\end{enumerate}