\documentclass[conference]{IEEEtran}

\usepackage{makeidx}  % allows for indexgeneration
\usepackage{graphicx}
\usepackage{indentfirst}
\usepackage{hyperref}
\usepackage{verbatim}
\usepackage{balance}
\usepackage{amsmath}

\begin{document}
\sloppy
\IEEEoverridecommandlockouts
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{From Abstract Parsing to Abstract Translation {\thanks{This work is licensed under the Creative Commons Attribution License.}}}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Semen Grigorev}
\IEEEauthorblockA{St. Petersburg State University\\
198504, Universitetsky prospekt 28\\
Peterhof, St. Petersburg, Russia\\
Email: rsdpisuy@gmail.com}
\and
\IEEEauthorblockN{Iakov Kirilenko}
\IEEEauthorblockA{St. Petersburg State University\\
198504, Universitetsky prospekt 28\\
Peterhof, St. Petersburg, Russia\\
Email: jake@math.spbu.ru}}

\maketitle

\begin{abstract}
String-embedded language transformation is one of the problems which can be faced during 
database and information system migration. The conventional solution which
is provided by a number of tools is based on run-time translation. We present a static 
\emph{abstract translation} approach which originates from the \emph{abstract parsing} 
technique~\cite{AbstrParsing} initially developed for syntax analysis of string-embedded 
languages. We present abstract translation algorithm and some optimization techniques, and
discuss the results of its evaluation on a real-world industrial application.

\end{abstract}

\IEEEpeerreviewmaketitle



\input{introduction}
\input{dags}
\input{abstract_translation}
\input{evaluation}
\input{conclusion}




\begin{thebibliography}{1}

\bibitem{OpenSystemsDBMS}
Shapot M., Popov E. Database reengineeing // Open Systems.DBMS. Number 4. 2004.    

\bibitem{ALVOR1}
Annamaa A., Breslav A., Kabanov J. e.a. An interactive tool for analyzing embedded SQL queries. Programming Languages and Systems. LNCS, vol. 6461. Springer: Berlin; Heidelberg. 2010. P. 131-138.

\bibitem{ALVOR2}
Annamaa A., Breslav A., Vene V. Using abstract lexical analysis and parsing to detect errors in string-embedded DSL statements // Proceedings of the 22nd Nordic Workshop on Programming Theory. Marina Walden and Luigia Petre, editors. 2010. P. 20-22.

\bibitem{StringExpr}
Aske Simon Christensen, Mller A., Michael I. Schwartzbach. Precise analysis of string expressions // Proc. 10th International Static Analysis Symposium (SAS), Vol. 2694 of LNCS. Springer-Verlag: Berlin; Heidelberg, June, 2003. P. 1-18.

\bibitem{Grune}
Grune D., Ceriel J. H. Jacobs. Parsing techniques: a practical guide. Ellis Horwood, Upper Saddle River, NJ, USA, 1990. P. 322.

\bibitem{SAofStrVal}
Costantini G., Ferrara P., Cortesi F. Static analysis of string values // Proceedings of the 13th international conference on Formal methods and software engineering, ICFEM’11. Springer-Verlag: Berlin; Heidelberg, 2011. P. 505-521.

\bibitem{ISO}
ISO. ISO/IEC 9075:1992: Title: Information technology — Database languages — SQL. 1992. P. 668.

\bibitem{JSA}
Java String Analyzer. URL: \href{http://www.brics.dk/JSA/}{http://www.brics.dk/JSA/}

\bibitem{AbstrParsing}
Kyung-Goo Doh, Hyunha Kim, David A. Schmidt. Abstract parsing: Static analysis of dynamically generated string output using LR-parsing technology // Proceedings of the 16th International Symposium on Static Analysis, SAS’09. Springer-Verlag: Berlin; Heidelberg, 2009. P. 256-272.

\bibitem{PHPSA}
PHP String Analyzer. URL: \href{http://www.score.is.tsukuba.ac.jp/~minamide/phpsa/}{http://www.score.is.tsukuba.ac.jp/$\sim$minamide/phpsa/}

\bibitem{PLSQL}
PL/SQL Developer. URL: \href{http://www.allroundautomations.com/plsqldev.html}{http://www.allroundautomations.com/plsqldev.html}

\bibitem{SQLWays}
SQL Ways. URL: \href{http://www.ispirer.com/products}{http://www.ispirer.com/products}

\bibitem{SwissSQL}
SwissSQL. URL: \href{http://www.swissql.com/}{http://www.swissql.com/}

\bibitem{SAForInject}
Xiang Fu, Xin Lu, Peltsverger B. e.a. A static analysis framework for detecting SQL injection vulnerabilities // Proceedings of the 31st Annual International Computer Software and Applications Conference. Vol. 01, COMPSAC’07, Washington, DC, USA, IEEE Computer Society, 2007. P. 87-96.

\end{thebibliography}
\balance
\end{document}

