\section{Introduction}

%Graph database management systems have become increasingly popular across diverse domains, including knowledge base analysis \cite{!!!}, biological data processing \cite{!!!}, semantic web applications \cite{!!!}, and source code analysis \cite{!!!}. Edge-labeled directed graphs provide a robust framework for data representation, where vertices correspond to objects and edges represent relations between them. This versatile approach is formalized in several data model standards, most notably the W3C Resource Description Framework (RDF) \cite{!!!}.

Language-constrained path querying (FLPQ)~\cite{doi:10.1137/S0097539798337716} provides a mechanism for searching paths in edge-labeled graphs where constraints are specified in terms of formal languages.
A path is considered valid if the word formed by concatenating the labels of its edges belongs to the specified language.
When constraints are regular languages, such queries are called regular path queries (RPQs). In practice, it is often useful to allow traversal opposite to edge directions, leading to two-way regular path queries (2-RPQs)~\cite{!!!,!!!}.
Queries specifying only a start or an end vertex are known as single-source or single-destination queries, respectively.
Both have been adopted in graph query languages such as GQL, SQL/PGQ~\cite{!!!}, and SPARQL~\cite{!!!}.

Graph database systems support different path query semantics, as the choice significantly affects both result interpretation and computational complexity.
Common semantics include: reachability (all vertices reachable via any valid path), all-shortest-paths (one shortest path per reachable vertex), all-simple-paths (all simple paths satisfying the query), and all-trails (all trails satisfying the query)~\cite{!!!}.
Each semantics serves different use cases, and the query engine must select the appropriate algorithm accordingly.

Efficient evaluation of RPQs and 2-RPQs remains an active research area \cite{!!!,!!!}.
Sparse linear algebra based methods have shown promise for RPQ evaluation \cite{!!!}.
In particular, the LARPQ (Linear Algebra-based Regular Path Query) algorithm \cite{!!!} employs a breadth-first traversal that simultaneously processes both the graph and the finite automaton representing the regular language.
While LARPQ demonstrates promising performance, it was designed solely for reachability, and its application to queries with fixed destination vertices remains unexplored.

The contributions of this paper are as follows.
\begin{itemize}
	\item A formal generalization of the LARPQ algorithm to a semiring-based framework, enabling the solution of various path-aggregation problems within a unified algorithmic structure.
	\item Instantiation of the framework with specific semirings to solve distinct problems: (i) Boolean semiring for reachability, and (ii) path semiring with index unary operators for all-simple-paths and all-trails semantics.
	\item ??? Proof of correctness and computational complexity analysis for the generalized algorithm.
	%\item Implementation within the LAGraph project using SuiteSparse:GraphBLAS for high-performance sparse matrix operations.
	\item Experimental evaluation on real-world knowledge graphs demonstrating significant performance improvements over state-of-the-art alternatives.
\end{itemize}

%The remainder of this paper is organized as follows. Section~\ref{sec:preliminaries} introduces necessary preliminaries, including edge-labeled graphs, formal language concepts, and matrix operations. Section~\ref{sec:la-graphs} presents the linear algebra foundations and describes the original LARPQ algorithm for reachability. Section~\ref{sec:semiring-framework} introduces our semiring-based generalization and its instantiations for different path semantics. Section~\ref{sec:implementation} discusses implementation details. Section~\ref{sec:evaluation} presents the experimental evaluation. Section~\ref{sec:related-work} discusses related work, and Section~\ref{sec:conclusion} concludes with directions for future research.
