\section{Related Work}\label{sec:related-work}

Regular path queries have been extensively studied in the context of graph databases. In this section, we discuss related work on linear algebra-based graph analytics, RPQ evaluation techniques, and semiring-based path problems.

\paragraph{Linear Algebra-Based Graph Analytics.}
The GraphBLAS initiative \cite{!!!,!!!} established a standardized interface for expressing graph algorithms using linear algebra operations.
LAGraph \cite{!!!} provides a collection of graph algorithms built on GraphBLAS, including BFS, PageRank, and triangle counting.
Our work extends this ecosystem with RPQ algorithms.

The use of sparse matrices for graph representation has proven effective for large-scale graphs.
Real-world graphs exhibit sparsity, allowing for memory-efficient storage and fast computation using formats like CSR/CSC \cite{!!!}.
The GraphBLAS specification supports arbitrary semirings, enabling the expression of algorithms that go beyond traditional numeric computations.

Recent work has explored linear algebra approaches for various graph problems.
However, RPQ evaluation in this paradigm remains underexplored, with only a few prior works \cite{!!!,!!!} addressing this problem.

\paragraph{RPQ Evaluation Methods.}
Traditional approaches to RPQ evaluation include:

\begin{itemize}
	\item \textbf{Automaton-based methods:} Construct the product automaton of the graph and the query NFA, then perform traversal \cite{!!!}. These methods can be combined with indexing techniques for improved performance.

	\item \textbf{BFS-based methods:} Perform breadth-first search guided by the query automaton \cite{!!!}. The original LARPQ algorithm \cite{!!!} falls into this category.

	\item \textbf{Matrix-based methods:} Translate the regular expression into matrix operations. The RPQ-matrix algorithm \cite{!!!} uses this approach, representing adjacency matrices in compressed formats like k2-trees or CSR/CSC.
\end{itemize}

Our approach combines elements from both BFS and matrix-based methods, using the simultaneous traversal of graph and automaton structures while leveraging sparse matrix operations for efficiency.

Different query semantics have been studied in the literature. The all-shortest-paths semantics is addressed in \cite{!!!}, while all-simple-paths and all-trails semantics are more challenging due to the potential for exponential result sizes. Our framework provides a unified approach for handling these different semantics.

\paragraph{Semiring-Based Path Problems.}

The algebraic approach to path problems has a long history. Kleene's algorithm \cite{!!!} computes reachability using matrix operations over the Boolean semiring. This approach has been generalized to arbitrary semirings for solving various algebraic path problems \cite{!!!,!!!}.

The connection between semirings and path enumeration is well-established. The tropical semiring (min-plus) corresponds to shortest path problems, while the Boolean semiring corresponds to reachability. Our work extends this connection to include the path semiring for explicit path enumeration.

Provenance semirings \cite{!!!} provide a framework for tracking the origin of query results. Our path semiring can be seen as a form of provenance tracking, where each result path is explicitly stored.

Context-free path querying (CFPQ) has also been approached using linear algebra \cite{!!!}. While CFPQ is more expressive than RPQ, our work focuses on regular languages and different evaluation semantics.

\paragraph{Graph Database Systems.}

Modern graph database systems vary significantly in their RPQ support:

\begin{itemize}
	\item \textbf{MillenniumDB} \cite{!!!}: A persistent graph database with state-of-the-art RPQ performance, supporting various path semantics.

	\item \textbf{FalkorDB} \cite{!!!} (formerly RedisGraph): An in-memory property-graph database that uses GraphBLAS for query evaluation. Supports a subset of regular path queries through Cypher.

	\item \textbf{Neo4j:} The most widely used graph database, supporting Cypher queries including limited forms of RPQ.

	\item \textbf{Blazegraph:} An RDF database with SPARQL support including property paths.
\end{itemize}

Our implementation differs from these systems in its focus on the linear algebra paradigm and its flexibility in supporting different semirings for various path semantics.

Table~\ref{tab:comparison} summarizes the comparison of our approach with related systems.

\begin{table}[h]
	\centering
	\MissingTable{Comparison of RPQ approaches}
	\caption{Comparison of RPQ evaluation approaches. LA = Linear Algebra, AB = Automaton-based, BFS = Breadth-first search.}
	\label{tab:comparison}
\end{table}

Our contributions differ from prior work in several key aspects:
\begin{itemize}
	\item We provide a \textbf{unified framework} that can solve multiple path semantics using different semiring instantiations;
	\item We extend the original LARPQ algorithm beyond reachability to path enumeration;
	\item We leverage the \textbf{GraphBLAS standard} for high-performance sparse matrix operations.
\end{itemize}
