%%  Full (research or experimental) papers: 16 LNCS style pages, including references
%%  Short (early findings or visionary) papers: up to 10 LNCS style pages, including references


% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%

% Placeholder for figures to be added later
\newcommand{\MISSINGFIGURE}{[Figure: Placeholder]}
\newcommand{\MissingTable}{[Table: Placeholder]}

\begin{document}
%
\title{Semiring-Based Linear Algebra Framework for Two-Way Regular Path Queries}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%

\author{Georgiy Belyanin\inst{1}\orcidID{0009-0004-7527-4281} \and Rodion Suvorov\inst{1}\orcidID{0009-00007-4934-4911} \and Iakov Kirilenko\inst{1}\orcidID{0000-0003-4384-8274} \and
	Semyon Grigoriev\inst{1}\orcidID{0000-0002-7966-0698}}

\authorrunning{G. Belyanin, R. Suvorov, I.Kirilenko and S. Grigoriev}
%
\institute{Saint Petersburg State University, Saint Petersburg, Russia
	\email{belyaningeorge@ya.ru, suvorov.53245324@gmail.com, y.kirilenko@spbu.ru, s.v.grigoriev@mail.spbu.ru}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Two-way regular path queries (2-RPQs) allow the use of regular languages over edges and their inverses in edge-labeled graphs to constrain paths of interest.
Adopted in modern graph query languages such as GQL, SQL/PGQ, and Cypher, 2-RPQs have become part of the GQL ISO standard.
However, their efficient evaluation on large-scale graphs remains challenging. 
In this paper, we propose a semiring-based generalization of the linear algebra approach for evaluating single-source 2-RPQs.
The original LARPQ algorithm solves the reachability problem using Boolean matrices.
We generalize it by replacing Boolean algebra with arbitrary semirings, enabling a unified solution for various path-aggregation problems.
Specifically, we employ the path semiring to find all simple paths and all trails.
The algorithms are implemented using the SuiteSparse:GraphBLAS library within the LAGraph infrastructure.
Experimental evaluation on real-world knowledge graphs demonstrates significant performance improvements over state-of-the-art solutions.

	\keywords{Regular path queries \and Two-way RPQ \and Linear algebra \and Semirings \and Graph databases \and Sparse matrices.}
\end{abstract}
%
%
%
\input{Introduction.tex}
\input{Preliminaries.tex}
\input{LA_Graps_Relations.tex}
\input{LA_RPQ_Paths_theory.tex}
%\input{Implementation_details.tex}
\input{Evaluation.tex}
\input{Related_work.tex}
\input{Conclusion.tex}


\begin{credits}
\subsubsection{\ackname} This research has been supported by the St. Petersburg State University, grant number 116636233.

\subsubsection{\discintname}
It is now necessary to declare any competing interests or to specifically
state that the authors have no competing interests. Please place the
statement with a bold run-in heading in small font size beneath the
(optional) acknowledgments\footnote{If EquinOCS, our proceedings submission
system, is used, then the disclaimer can be provided directly in the system.},
for example: The authors have no competing interests to declare that are
relevant to the content of this article. Or: Author A has received research
grants from Company W. Author B has received a speaker honorarium from
Company X and owns stock in Company Y. Author C is a member of committee Z.
\end{credits}
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
 \bibliographystyle{splncs04}
 \bibliography{RPQ_paths_LA}
%
\end{document}
