\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage[table]{xcolor}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{The Library of GPGPU-Powered Sparse Boolean Linear Algebra Operations
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{
\IEEEauthorblockN{Egor Orachev}
\IEEEauthorblockA{\textit{Saint Petersburg State University} \\
%\textit{name of organization (of Aff.)}\\
St. Petersburg, Russia \\
egor.orachev@gmail.com}
\and

\IEEEauthorblockN{Maria Karpenko}
\IEEEauthorblockA{\textit{ITMO University} \\
%\textit{name of organization (of Aff.)}\\
St. Petersburg, Russia \\
mkarpenko.spb@gmail.com}
\and

% \IEEEauthorblockN{Vasily Kuporosov}
% \IEEEauthorblockA{\textit{HSE University} \\
% %\textit{name of organization (of Aff.)}\\
% St. Petersburg, Russia \\
% vvkuporosov@edu.hse.ru}
% \and

% Не бейте меня. Просто хотел места больше свободного сделать

\IEEEauthorblockN{Artem Khoroshev}
\IEEEauthorblockA{\textit{Computation Biology} \\
\textit{Department} \\
\textit{BIOCAD}\\
St. Petersburg, Russia \\
arthoroshev@gmail.com}
\and

\IEEEauthorblockN{Semyon Grigorev}
\IEEEauthorblockA{\textit{Saint Petersburg State University, }\\
                          %{7/9 Universitetskaya nab.}\\
                  \textit{JetBrains Research,} \\
                          %{Primorskiy prospekt 68-70, Building 1,}
                  {St. Petersburg, Russia}\\
s.v.grigoriev@spbu.ru, \\ semyon.grigorev@jetbrains.com}
}

\maketitle

\begin{abstract}
Sparse matrices are widely applicable in data analysis while the theory of matrix processing is well-established.
There are a wide range of algorithms for basic operations such as matrix-matrix and matrix-vector multiplication, factorization, etc.
To facilitate data analysis, GraphBLAS API provides a set of building blocks and allows for reducing algorithms to sparse linear algebra operations.
While GPGPU utilization for high-performance linear algebra is common, the high complexity of GPGPU programming makes the implementation of GraphBLAS API on GPGPU challenging.
In this work, we present a GPGPU library of sparse operations for an important case --- Boolean algebra.
The library is based on modern algorithms for sparse matrix processing.
We provide a Python wrapper for the library to simplify its use in applied solutions.
Our evaluation shows that operations specialized for Boolean matrices can be up to 2 times faster and consume up to YYY times less memory than generic operations from modern libraries.
We hope that our results help to move the development of a GPGPU version of GraphBLAS API forward.
\end{abstract}

\begin{IEEEkeywords}
sparse linear algebra, GPGPU, boolean semiring, sparse boolean matrix
\end{IEEEkeywords}


\input{introduction}
% \input{related_work}
\input{about_libraries}
\input{implementation_details}
\input{evaluation}
\input{conclusion}

%\section*{Acknowledgment}
%
%The preferred spelling of the word ``acknowledgment'' in America is without
%an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B.
%G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor
%acknowledgments in the unnumbered footnote on the first page.


\bibliographystyle{./IEEEtran}
\bibliography{./Sparse_Boolean_Algebra_on_GPGPU}

\end{document}
