\section{Introduction}

Problems in many areas can be reduced to one of the formal-languages-constrained path problems~\cite{barrett2000formal}. In these areas, it is often required to process queries for large graphs using formal languages to describe paths to find. For example, various problems of static code analysis~\cite{bastani2015specification,xu2009scaling} can be formulated in terms of the context-free language reachability~\cite{reps1998program} or in terms of the linear conjunctive language reachability~\cite{zhang2017context}.

The result of a context-free path query evaluation is usually a set of triples \mbox{$(A, m, n)$}, such that there is a path from the node $m$ to the node $n$, whose labeling is derived from a non-terminal $A$ of the given context-free grammar. This type of query is evaluated using the \emph{relational query semantics}~\cite{hellingsRelational}. There is a number of algorithms for context-free path query evaluation using this semantics~\cite{azimov2018context,GLL,hellingsRelational,GraphQueryWithEarley,RDF}.

In~\cite{azimov2018context}, the matrix-based algorithm for context-free path query evaluation w.r.t. relational query semantics is proposed. This algorithm computes the same parsing table as the CYK algorithm but does this by offloading the most intensive computations into calls to a Boolean matrix multiplication procedure. This approach allows us to effectively apply \emph{GPGPU} (General-Purpose computing on Graphics Processing Units) computing techniques and other optimizations for matrix operations.

Also, there are conjunctive grammars~\cite{okhotinConjAndBool}, which have more expressive power than context-free grammars. For example, conjunctive grammars are used in static analysis to describe an interleaved matched-parentheses language~\cite{zhang2017context}, which is not context-free. Path querying with conjunctive grammars is known to be undecidable~\cite{hellingsRelational}. Although there is an algorithm~\cite{zhang2017context} for path querying with linear conjunctive grammars~\cite{okhotinConjAndBool} which provides an over-approximation of the result. However, there is no algorithm for path querying with conjunctive grammars of an arbitrary form.

The purpose of this work is to develop the algorithm
for path querying with arbitrary conjunctive grammars using the matrix-based approach from~\cite{azimov2018context}. It will also be shown that the proposed algorithm allows us to effectively apply GPGPU computing techniques.

This paper is structured as follows: the section~\ref{section_preliminaries} defines some notions, used later on; in the section~\ref{section_related} the overview of related works is presented; the section~\ref{section_main} discusses our matrix-based algorithm for path querying with arbitrary conjunctive grammar and provides a step-by-step demonstration for a small example;  we evaluate the performance of our algorithm in the section~\ref{section_evaluation}, and provide some concluding remarks in the section~\ref{section_conclusion}.
