% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}

\usepackage{amssymb}

\usepackage{caption}
\usepackage{subcaption}
%\usepackage{gnuplottex}
\usepackage{tikz}
\usepackage{mathtools}

\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{textcomp}

\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{algorithmicx}

\usepackage{booktabs}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
\newtheorem{mytheorem}{Theorem}
%\newtheorem{lemma}{Lemma}
\newtheorem{mydef}{Definition}

\algrenewcommand\algorithmicindent{0.5em}
\algnewcommand\algorithmicswitch{\textbf{switch}}
\algnewcommand\algorithmiccase{\textbf{case}}
\algnewcommand\algorithmicassert{\texttt{assert}}
\algnewcommand\Assert[1]{\State \algorithmicassert(#1)}
% New "environments"
\algdef{SE}[SWITCH]{Switch}{EndSwitch}[1]{\algorithmicswitch\ #1\ \algorithmicdo}{\algorithmicend\ \algorithmicswitch}
\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1}{\algorithmicend\ \algorithmiccase}

\algtext*{EndSwitch}
\algtext*{EndCase}
\algtext*{EndWhile}% Remove "end while" text
\algtext*{EndIf}% Remove "end if" text
\algtext*{EndFor}% Remove "end for" text
\algtext*{EndFunction}% Remove "end function" text

\newif\ifboldnumber
\newcommand{\boldnext}{\global\boldnumbertrue}	
	
%
\title{Path Querying with Conjunctive Grammars by Matrix Multiplication\thanks{Supported by the Russian Science Foundation grant 18-11-00100 and a grant from JetBrains Research.}}

\author{Rustam Azimov \and
	Semyon Grigorev}
%
\authorrunning{Azimov R.,Grigorev S.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Saint Petersburg State University, St. Petersburg, Russia\\
	\email{st013567@student.spbu.ru}\\
	\email{s.v.grigoriev@spbu.ru}}

\maketitle

\begin{abstract}
Problems in many areas can be reduced to one of the formal-languages-constrained path problems. In these areas, it is often required to process queries for large graphs using formal languages to describe paths to find. For example, various problems of static code analysis can be formulated in terms of the context-free language reachability or in terms of the linear conjunctive language reachability.

Conjunctive grammars have more expressive power than context-free grammars and are used, for example, in static analysis to describe an interleaved matched-parentheses language, which is not context-free. Path querying with conjunctive grammars is known to be undecidable. Although there is an algorithm for path querying with linear conjunctive grammars which provides an over-approximation of the result. However, there is no algorithm for path querying with arbitrary conjunctive grammars.

We propose the first algorithm for path querying with arbitrary conjunctive grammars. The proposed algorithm is matrix-based and allows us to effectively apply GPGPU (General-Purpose computing on Graphics Processing Units) computing techniques and other optimizations for matrix operations.

\keywords{Conjunctive grammars \and Path querying \and Graphs \and Transitive closure \and GPGPU \and Static analysis \and Matrix multiplication}

\end{abstract}


\input{introduction.tex}
\input{preliminaries.tex}
\input{related_works.tex}
\input{algorithm_conj.tex}
\input{evaluation.tex}
\input{conclusion.tex}

\bibliographystyle{splncs04}
\bibliography{graphparsing}

\end{document}
