\section{Related works} \label{section_related}
Problems in many areas can be reduced to one of the formal-languages-constrained path problems~\cite{barrett2000formal}. For example, various problems of static code analysis~\cite{bastani2015specification,xu2009scaling} can be formulated in terms of the context-free language reachability~\cite{reps1998program} or in terms of the linear conjunctive language reachability~\cite{zhang2017context}. 

%One of the well-known problems in the area of graph database analysis is the language-constrained path querying. For example, the regular language constrained path querying~\cite{reutter2017regular, fan2011adding, abiteboul1997regular, nole2016regular}, and the context-free language constrained path querying.

There are a number of solutions~\cite{hellingsRelational, GraphQueryWithEarley, RDF, azimov2018context} for context-free path query evaluation w.r.t. the relational query semantics, which employ such parsing algorithms as CYK~\cite{kasami, younger} or Earley~\cite{Grune}. Other examples of path query semantics are \textit{single-path} and \textit{all-path query semantics}. The all-path query semantics requires presenting all possible paths from node $m$ to node $n$ whose labeling is derived from a non-terminal $A$ for all triples $(A, m, n)$ evaluated using the relational query semantics. While the single-path query semantics requires presenting only one such a path for all the node-pairs $(m, n)$. Hellings~\cite{hellingsPathQuerying} presented algorithms for the context-free path query evaluation using the single-path and the all-path query semantics. If a context-free path query w.r.t. the all-path query semantics is evaluated on cyclic graphs, then the query result can be an infinite set of paths. For this reason, in~\cite{hellingsPathQuerying}, annotated grammars are proposed as a possible solution.

In~\cite{GLL}, the algorithm for context-free path query evaluation w.r.t. the all-path query semantics is proposed. This algorithm is based on the generalized top-down parsing algorithm~---~GLL~\cite{scott2010gll}. This solution uses derivation trees for the result representation which is more native for grammar-based analysis. The algorithms in~\cite{GLL, hellingsPathQuerying} for the context-free path query evaluation w.r.t. the all-path query semantics can also be used for query evaluation using the relational and the single-path semantics.

Hellings~\cite{hellingsRelational} presented an algorithm for the context-free path query evaluation using the relational query semantics. According to Hellings, for a given graph $D = (V, E)$ and a grammar $G = (N, \Sigma, P)$ the context-free path query evaluation w.r.t. the relational query semantics reduces to a calculation of the context-free relations $R_A$. Thus, in this work, we focus on the calculation of conjunctive relations which are similar to the context-free relations.

In~\cite{azimov2018context}, the matrix-based algorithm for context-free path query evaluation w.r.t. relational query semantics is proposed. This algorithm computes the same parsing table as the CYK algorithm but does this by offloading the most intensive computations into calls to a Boolean matrix multiplication procedure. This approach allows us to effectively apply GPGPU computing techniques and other optimizations for matrix operations.

Also, there is an algorithm~\cite{zhang2017context} for path querying with linear conjunctive grammars and relational query semantics. This grammars have no more than one nonterminal in each conjunct of the rule. The possibility of creating an algorithm for path query evaluation w.r.t. conjunctive grammars of an arbitrary form is an open problem since the linear conjunctive grammars are known to be strictly less powerful than the arbitrary conjunctive grammars~\cite{okhotinConjAndBool}.