% Use pdflatex.exe -synctex=1 -interaction=nonstopmode "graphparsing".tex
% For bibliography use BibTex
\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}


% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[GRADES-NDA'18]{The 1st Joint International Workshop on Graph Data Management Experiences \& Systems (GRADES) and Network Data Analytics (NDA) 2018}{June 10, 2018}{Houston, Texas, USA}
\acmYear{2017}
\copyrightyear{2018}


\acmArticle{4}
\acmPrice{15.00}

\usepackage{graphicx}

\usepackage{caption}
\usepackage{subcaption}
%\usepackage{gnuplottex}
\usepackage{tikz}
\usepackage{mathtools}

\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{textcomp}

\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{algorithmicx}

\settopmatter{printacmref=false, printccs=true, printfolios=false}

%\pagestyle{plain} % removes running headers


\begin{document}

\title{Context-Free Path Querying by Matrix Multiplication}

\author{Rustam Azimov}
\affiliation{%
	\institution{Saint Petersburg State University}
	\streetaddress{7/9 Universitetskaya nab.}
	\city{St. Petersburg} 
	\state{Russia} 
	\postcode{199034}
}
\email{rustam.azimov19021995@gmail.com}

\author{Semyon Grigorev}
\affiliation{%
	\institution{Saint Petersburg State University}
	\streetaddress{7/9 Universitetskaya nab.}
	\city{St. Petersburg}
	\state{Russia}  
	\postcode{199034}
}
\email{Semen.Grigorev@jetbrains.com}
		
\newtheorem{mytheorem}{Theorem}
%\newtheorem{lemma}{Lemma}
\newtheorem{mydef}{Definition}

\algrenewcommand\algorithmicindent{0.5em}
\algnewcommand\algorithmicswitch{\textbf{switch}}
\algnewcommand\algorithmiccase{\textbf{case}}
\algnewcommand\algorithmicassert{\texttt{assert}}
\algnewcommand\Assert[1]{\State \algorithmicassert(#1)}
% New "environments"
\algdef{SE}[SWITCH]{Switch}{EndSwitch}[1]{\algorithmicswitch\ #1\ \algorithmicdo}{\algorithmicend\ \algorithmicswitch}
\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1}{\algorithmicend\ \algorithmiccase}

\algtext*{EndSwitch}
\algtext*{EndCase}
\algtext*{EndWhile}% Remove "end while" text
\algtext*{EndIf}% Remove "end if" text
\algtext*{EndFor}% Remove "end for" text
\algtext*{EndFunction}% Remove "end function" text

\newif\ifboldnumber
\newcommand{\boldnext}{\global\boldnumbertrue}




\begin{abstract}
The algorithms for context-free path query evaluation are widely used in many areas, for example, graph databases, bioinformatics. In these areas, it is often required to process queries for large graphs. The existing algorithms for context-free path query evaluation perform poorly on large graphs. One of the open problem in this area is to create a matrix-based algorithm for context-free path query evaluation. The active use of matrix operations (such as matrix multiplication) in the process of the context-free path query evaluation makes it possible to efficiently apply such computing techniques as \textit{GPGPU} (General-Purpose computing on Graphics Processing Units) and parallel computation. From a practical point of view, matrix multiplication may be performed on different GPU independently. It can help to utilize the power of multi-GPU systems and increase the performance of the context-free path querying. Also, the algorithms for distributed-memory matrix multiplication allow us to handle graph sizes inherently larger than the DRAM memory available on the GPU. In this paper, we show how the context-free path query evaluation can be reduced to the calculation of the matrix transitive closure. Also, we propose the first matrix-based algorithm for context-free path query evaluation which makes it possible to speed up computations by using a GPU. Furthermore, we evaluate the performance of our algorithm on some popular ontologies.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.

\begin{CCSXML}
	<ccs2012>
	<concept>
	<concept_id>10002951.10002952.10003197.10010825</concept_id>
	<concept_desc>Information systems~Query languages for non-relational engines</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10003752.10003766.10003771</concept_id>
	<concept_desc>Theory of computation~Grammars and context-free languages</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Query languages for non-relational engines}
\ccsdesc[500]{Theory of computation~Grammars and context-free languages}

\keywords{Transitive closure, context-free path querying, graph databases, context-free grammar, GPGPU, matrix multiplication}

\maketitle

\input{introduction.tex}
\input{motivating.tex}
\input{preliminaries.tex}
\input{related_works.tex}
\input{algorithm.tex}
\input{evaluation.tex}
\input{conclusion.tex}
\input{appendices.tex}
\input{acknowledgments.tex}

\bibliographystyle{ACM-Reference-Format}
\bibliography{graphparsing}

\end{document}
