\section{Conclusion}

We propose a native way to integrate a language for context-free path querying into a general-purpose programming language.
Our solution can handle arbitrary context-free grammars and arbitrary input graphs.
The proposed approach is language-independent and may be implemented for closely all general-purpose programming languages.
We implement it in the Scala programming language and show that our implementation can be applied to the real world problems.

We can propose some possible directions for the future work.
First of all, it is necessary to formulate the creation of a user-friendly interface for SPPF processing.
We can just extract reachability information, but SPPF contains mach more useful information.
One such representation may be a set of paths with additional information about their structure.
This may simplify debugging and query result processing.

In order to improve performance and investigate scalability of proposer solution it is necessary to try to implement parallel single machne and distributed GLL.
It is not only algorithmic problem: to get practical solution we should choose appropriate tools, libraryes for parallel and distributet computing for Scala.

Another direction is a semantic actions computation, otherwise known as attributed grammars handling.
It increases the expressiveness of queries by means of the specification of user-defined actions, such as filters, over subqueries result. 
Although it is impossible in general, techniques such as lazy evaluation can provide a technically adequate solution.
Anoter possible direction is utilizatuin of relational programming (minikanren) which is aimed to search~\cite{DB}.
For what class of semantic actions it is possible to provide a precise general solution is a theoretical question to be answered. 

Some important problems in static code analysis reqwuire languages more expressive than contex-free one.
For example, context-sensitive data-dependence analysis may be precisely expressed in terms of linear-conjunctive language~\cite{Okhotin2003LCL} reachability, but not context-free~\cite{LCLReachability}.
While problem formulation is precise, it is possible to get only approximated solution, because emptyness problem for linear-conjunctive languages is undersidable.
It would be an interesting task to support not only linear-cinjunctive grammars, but arbitrary conjunctive grammars~\cite{okhotin2001conjunctive} in te library and investigate nature of approximation.

All these improvements may provide !!! Declarative code analysis tools and languages Flix~\cite{Flix}. Reps~\cite{Reps:1995}

Improved version of OpenCypher~\cite{OpenCypherPR}, which is the one of the most popular graph query languages, provides context-free path querying mechanism.
Detailed comparison with it may provide more information for direcion of future work.
