\section{Introduction}

Graph querying is finding all paths in the graph which satisfy some constraints. 
If the constraints are specified with some language formalism, i.e. a grammar, it is called a language-constrained path query. 
For example, a grammar $S \rightarrow a \ S \ b \mid a \ b$ can be regarded as a query for the paths of the form $a^n b^n$, where $n \geq 1$.
Querying the graph in the fig.~\ref{fig:exmplInputGraph} returns the set of paths one of which starts and ends in the vertex $3$ and goes around the cycles in the graph the appropriate number of times: $3,1,2,3,1,2,3,4,3,4,3,4,3$. 

% One type of graph queries is language-constrained path queries~\cite{FLCpathProblem} which mean that constraints are formulated by using formal languages formalisms such as grammars.
% For example, suppose that one have a graph presented in fig.~\ref{fig:exmplInputGraph} and want to find all paths which have a form $a^nb^n, n > 0$.
% It is possible to specify this constraint in terms of context-free grammar $G$, which is presented in fig~\ref{fig:anbnGrammar}.

% \begin{figure}[h]
% \begin{align*}
% S & \rightarrow a \ S \ b \
% \\
% S & \rightarrow a \ b
% \end{align*}
% \caption{Context-free grammar $G$ for the language $L=\{a^nb^n| n > 0\}$}
% \label{fig:anbnGrammar}
% \end{figure}


\begin{figure}[h]
\begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,auto]
   \node[state] (q_1)   {$1$};
   \node[state] (q_2) [above=of q_1] {$2$};
   \node[state] (q_3) [above right=of q_1, below right=of q_2] {$3$};
   \node[state] (q_4) [right=of q_3] {$4$};
    \path[->]
    (q_1) edge  node {a} (q_2)
    (q_2) edge  node {a} (q_3)
    (q_3) edge  node {a} (q_1)
    (q_3) edge[bend left, above]  node {b} (q_4)
    (q_4) edge[bend left, below]  node {b} (q_3);
\end{tikzpicture}
\caption{An example of input graph}
\label{fig:exmplInputGraph}
\end{figure}

Most existing graph traversing/querying languages, including SPARQL~\cite{sparql}, Cypher~\footnote{Cypher language web page: \url{https://neo4j.com/developer/cypher-query-language/}. Access date: 16.01.2018}, and Gremlin~\cite{gremlin} support only regular languages as constrains. 
For some applications regular languages are not expressive enough. 
Context-free path queries (CFPQ), on which we focus this paper, employ context-free languages for constraints specification. 
CFPQs are used in bioinformatics~\cite{GraphQueryWithEarley}, static code analysis~\cite{Reps, Zheng, LabelFlowCFLReachability, specificationCFLReachability, JavaCFL}, and RDF processing~\cite{CFGonRDF}. 
Although there is a lot of problem-specific solutions and theoretical research on CFPQs~\cite{Yannakakis, ConjCFPathQuery, Hellings16, GrigorevR16, QueryGraphWithData, RegularDBQuery, GraphQueryWithEarley, graphDB}, cfSPARQL~\cite{CFGonRDF} is the single known graph query language to support CF constraints. 
Generic solution for the integration of CFPQs into general-purpose languages is not discussed enough. 

%There are several languages for graph traversing/querying which support constraints formulated in terms of regular languages.
%For example SPARQL~\cite{sparql}, Cypher~\footnote{Cypher language web page: \url{https://neo4j.com/developer/cypher-query-language/}. Access date: 16.01.2018}, and Gremlin~\cite{gremlin}.
%In this work, we are focused on context-free path queries (CFPQs) which use context-free languages for constraints specification and are used in bioinformatics~\cite{GraphQueryWithEarley}, static code analysis~\cite{Reps, Zheng, LabelFlowCFLReachability, specificationCFLReachability, JavaCFL}, and RDF processing~\cite{CFGonRDF}.
%There are a lot of problem-specific solutions and theoretical research on CFPQs~\cite{Yannakakis, ConjCFPathQuery, Hellings16, GrigorevR16, QueryGraphWithData, RegularDBQuery, GraphQueryWithEarley, graphDB}.
%Among them, cfSPARQL~\cite{CFGonRDF} is a single known graph query language to support CF constraints. Generic solution for the integration of CFPQs into general-purpose languages is not discussed enough.

When developing a data-centric application, one wants to use a general-purpose programming language and also to have a transparent and native access to data sources. 
One way to achieve this is to use string-embedded DSLs. 
In this approach, a query is written as a string, then passed on to a dedicated driver which executes it and returns a possibly untyped result.
Despite the simplicity, string-embedded DSLs have serious drawbacks. 
First of all, they require the developer to learn the language itself, its features, runtime and how the integration between the languages is implemented. 
DSLs are also a source of possible errors and vulnerabilities, static detection of which is a serious challenge~\cite{stringEmbeddedLanguagesProblem}. 
Such techniques as the Object Relationship Mapping (ORM) or Language Integrated Query~(LINQ)~\cite{LINQ1, LINQ2, LinqRDF} partially solve these problems, but they still have issues with flexibility: both query decomposition and  reusing of subqueries are a struggle. 
In this paper, we propose a transparent and natural integration of CFPQs into a general-purpose programming language.



%When one develops a data-centric application, one wants to use a general purpose programming language and have a transparent and native access to data sources.
%String-embedded DSLs is one way to do it.
%It utilizes a driver to execute a query written as a string and to return a possibly untyped result.
%This approach has serious drawbacks.
%First of all, a DSL may require additional knowledge from a developer.
%Moreover, a string-embedded language itself is a source of possible errors and vulnerabilities, static detection of which is very difficult~\cite{stringEmbeddedLanguagesProblem}.
%In trying to solve these issues, such special techniques as Object Relationship Mapping (ORM) or Language Integrated Query~(LINQ)~\cite{LINQ1, LINQ2, LinqRDF} were created. Unfortunately, they still experience difficulties with flexibility: for example with the query decomposition and the reusing of subqueries.
%In this paper, we propose a transparent and natural integration of CFPQs into a general-purpose language.

Context-free path queries are known in various domains under different names. The \emph{context-free language reachability framework} or  \emph{IFDS framework} is how they are called in the area of static code analysis. 
In~\cite{Reps:1995, Reps} Thomas Reps shows that the wide range of static code analysis problems can be formulated in terms of CFL-reachability in the graph. 
This framework is used for such problems as the taint analysis~\cite{CFLTaint}, the alias analysis~\cite{JavaCFL, Zheng, CFLGraspan}, the label flow analysis~\cite{LabelFlowCFLReachability}, and the fix locations problem~\cite{CFLfinding}.
What we propose in the paper can be viewed as a core of such framework since it provides both problem and domain independent mechanism for CFPQ evaluation. 

%Note that CFPQs is applicable not only for graph data base querying but also for static code analysis (with name \emph{context-free language reachability framework} or  \emph{IFDS framework}).
%CFPQ and context-free language reachability (CFL-reachability) is a different names of one thing: the first used mostely in database community and the second one in static code analysis.
%In 1995 Thomas Reps shows that the wide range of static code analysis problems can be formulated in terms of CFL-reachability in graph~\cite{Reps:1995, Reps}.
%This framework is widely used for solving different problems~\cite{JavaCFL, Zheng, LabelFlowCFLReachability, CFLfinding, CFLTaint, CFLGraspan} and in domain specific languages, like the Flix~\cite{Flix}.
%For softaware development it is more more convinient way to have an generic integrated framework (library) for general-purpose prgramming language to avoid integration problems.
%Our solurion may be use as a core of such framework because provide generic (problem and domain independent) mechanism for CFPQs evaluation.

We view parser combinators as the best way to integrate context-free language specifications into a general-purpose programming language. Parser combinators not only provide a transparent integration but also compile-time checks of correctness and high-level techniques for generalization. An idea to use combinators for graph traversing has already been proposed in~\cite{ScalaGraphParsing}. Unfortunately, the solution presented processes cycles in the input graph only approximately and is unable to handle left-recursive combinators, which is the most common issue of the approach. Authors pointed out that the idea described is similar to the classical parser combinators, but the language class supported or restrictions are not discussed.

%It is necessary to find an appropriate technique for integration of context-free language specification into general-purpose programming languages.
%One natural way to specify a language is to specify its formal grammar which can be done by using a special DSL based, for example, on EBNF-like notation~\cite{EBNFISO}.
%The classical alternative way is parser combinators~\cite{MonadicPArserCombinators} which provide all required features, including transparent integration, compile-time checks of correctness, high-level techniques for generalization.

%An idea to use combinators for graph traversing has already been proposed in~\cite{ScalaGraphParsing}, but the solution presented provides only approximated handling of cycles in the input graph and does not support left-recursive grammars.
%Authors pointed out that the idea described is very similar to the classical parser combinators, but the language class supported or restrictions are not discussed.
%This point is very important, because conventional combinators implement top-down parsing and cannot handle left-recursive and ambiguous grammars.

Parser combinators are known to handle only a subset of context-free grammars: left recursion and ambiguity of the grammars are problematic. 
In~\cite{Meerkat}, authors demonstrate a set of parser combinators which handles arbitrary context-free grammars by using ideas of the Generalized LL~\cite{scott2010gll} algorithm (GLL). 
Meerkat~\footnote{Meerkat project repository:~\url{https://github.com/meerkat-parser/Meerkat}. Access date: 16.01.2018} parser combinators library implements the ideas from the paper~\cite{Meerkat} and provides the parsing result in a compact form as a Shared Packed Parse Forest~\cite{SPPF}~(SPPF).
SPPF is a suitable finite structural representation of a CFPQ result, even when the set of paths is infinite~\cite{GrigorevR16}. 
All the paths can be extracted from the SPPF---as the corresponding derivation trees---and further analysis can be done. 
It is also possible to run some further processing over the SPPF itself---without explicit paths extraction.

%In~\cite{Meerkat}, authors demonstrate a set of parser combinators which can handle arbitrary context-free grammars by using ideas of Generalized LL~\cite{scott2010gll} (GLL).
%Meerkat~\footnote{Meerkat project repository:~\url{https://github.com/meerkat-parser/Meerkat}. Access date: 16.01.2018} parser combinators library is based on~\cite{Meerkat} and provides result of parsing in a compact form as Shared Packed Parse Forest~\cite{SPPF}~(SPPF).
%It is showed that SPPF is a suitable finite structural representation of a CFPQ query result, even if the set of paths is infinite~\cite{GrigorevR16}, which can be used for paths extraction, queries debugging and processing of result.

In this paper, we compose these ideas and present a set of parser-combinators for context-free path querying which handle arbitrary context-free grammars and provide a structural representation of the result. 
We make the following contributions in the paper.

%In this paper, we show how to compose these ideas and present the parser combinators for CFPQ which can handle arbitrary context-free grammars and provide a structural representation of the result.
%We make the following contributions in this paper.

\begin{enumerate}
\item We show that it is possible to create a set of parser combinators for context-free path querying which work on both arbitrary context-free grammars and arbitrary graphs and provide a finite structural representation of the query result.
\item We implement the parser combinators library in Scala. This library provides an integration to Neo4j~\footnote{Neo4j graph database site: \url{https://neo4j.com/}. Access date: 16.01.2018} graph database. The source code is available on GitHub: \url{https://github.com/YaccConstructor/Meerkat}.
\item We perform an evaluation on realistic data and compare the performance of our library with another GLL-based CFPQ tool and with the Trails library.
We conclude that our solution is expressive and performant enough to be applied to the real-world problems.
\end{enumerate}

%\begin{enumerate}
%\item We show that it is possible to create parser combinators for context-free path querying which work on both arbitrary context-free grammars and arbitrary graphs and provide a finite structural representation of the query result.
%\item We provide the implementation of the parser combinators library in Scala. This library provides an integration to Neo4j~\footnote{Neo4j graph database site: \url{https://neo4j.com/}. Access date: 16.01.2018} graph database. The source code is available on GitHub:\url{https://github.com/YaccConstructor/Meerkat}.
%\item We perform an evaluation on realistic data.
%Also, we compare the performance of our library with another GLL-based CFPQ tool and with the Trails library.
%We conclude that our solution is expressive and performant enough to be applied to the real-world problems.
%\end{enumerate}

This paper is organized as follows.
We introduce a formal definition of the CFPQ problem in section~\ref{sec:CFPQ} and we provide a basic description of the Meerkat library and SPPF data structure in section~\ref{sec:GLL}.
We describe our solution in section~\ref{sec:combinators}.
In section~\ref{sec:examples} we present and discuss a set of classical queries (the same generation query, the queries to a movie dataset~\footnote{The movie database is a traditional dataset for graph databases. Detailed description is available here: \url{https://neo4j.com/developer/movie-database/}. Access date: 16.01.2018}) 
formulated in terms of our library.
Evaluation of the library is described in section~\ref{sec:evaluation}.
Finally, In section~\ref{sec:conclusion} we conclude and discuss possible directions for further research. 
