\section{Proposed Solution}

% The goal of our research is to figure out whether distillation~\cite{hamilton2021700} could be a solution to the intermediate data structures problem in linear algebra-based programs.
The goal of our research is to find out if linear algebra-based programs can be efficiently optimized by program distillation~\cite{hamilton2021700} through eliminating intermediate data structures and computations.
To answer this question, we have developed a library of matrix operations\footnote{\url{https://github.com/YaccConstructor/Distiller/blob/master/inputs/LinearAlgebra.pot}} in POT language: a simple functional language used by Hamilton in his distiller\footnote{\url{https://github.com/poitin/Distiller}}.  

% We use a quad-tree representation~\cite{qtree} for matrices because it avoids indexing and is natural for functional programming since it is expressible in terms of algebraic data types.
We use a quad-tree matrix representation~\cite{qtree} since it both avoids indexing and can be implemented via algebraic data type as depicted in listing~\ref{lst:qtree}, which itself is very natural for functional programming.
% Moreover, it makes it possible to represent both sparse and dense matrices naturally, and express basic operations over such a representation via recursive functions which traverse this tree-like structure.
Besides, it provides \at{similar compression rate to widely adopted CSR and COO}, and a natural way to represent both sparse and dense matrices, as well as makes it possible to express basic operations over the representation via recursive functions traversing the tree-like structure.
% Finally, this structure allows to conveniently exploit divide-and-conquer parallelism in matrices handling functions.
Finally, the quad-tree representation allows to natively exploit divide-and-conquer parallelism in matrices handling functions.

\begin{listing}

    \begin{minted}[]{Haskell}
  
    data QTree a = QNone  
                 | QVal a 
                 | QNode (QTree a) (QTree a)
                         (QTree a) (QTree a) 
    
    \end{minted}
    \caption{Quad-tree compressed representation}
    \label{lst:qtree}
    
    \end{listing}

\at{Since general purpose devices like CPUs and GPUs appear to be heavily underutilized when executing sparse routines due to low arithmetic intensity and memory boundness, custom hardware seems to be a promising solution towards mitigating these issues. 
Distillation provides all the needed optimizations for free for a functional language, so what remains is to translate a functional program into custom hardware. We have opted for two solutions here.
The first one is Reduceron~\cite{naylorRunciman2012} --- a processor, designed to perform highly performant reductions. 
The second one is FHW~\cite{Edwards2019FHWP} --- a project, which is constituted of several compilers that help to translate an arbitrary Haskell progam into System Verilog to eventually provide a bitstream for a custom hardware.
It leverages a parallel and pipelined dataflow representation, which is abstracted over, e.g., nodes for memory operations, which makes it possible to come up with a specialized memory for our data structures.
% We selected two different target hardware platforms.
% The first one is Reduceron~\cite{naylorRunciman2012} --- a general-purpose functional-language-specific processor.
% The second is program-specific hardware for arbitrary functional programs FHW~\cite{Edwards2019FHWP} which utilizes the flexibility of FPGA to create hardware for a particular program.
While the first case is more typical, the second might provide higher performance for specific tasks.}

% At the current stage,
For now, we propose to use program distillation as the first step of program optimization which, we hope, should reduce memory % traffic,
usage \at{and other unnecessary computations}, and then compile a distilled program to the two different hardware platforms by using the respective compiler with platform-specific optimizations.
\at{For evaluation, we have been implementing \db{\{propose or some of them are already implemented?!? =)) \}} a library of a subset of linear algebraic routines, only matrix-matrix operations for now.
Programs of interest are compositions of these basic functions.}
