% Short papers (up to 4 pages) and long papers (up to 10 pages). 
% All submissions must be single-spaced double-column pages using 10-point size font on 8.5x11 inch pages (IEEE conference style),
% including figures, tables, and references.
\documentclass[sigconf, anonymous, review,table]{acmart}
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{hyperref}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{multirow}

\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation emai}{June 03--05,
  2018}{Woodstock, NY}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}


%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}


%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Spla: Open-Source Generalized Sparse Linear Algebra Framework with Vendor-Agnostic GPUs Accelerated Computations}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Egor Orachev}
\email{egor.orachev@gmail.com}
\orcid{0000-0002-0424-4059}
\author{Semyon Grigorev}
\email{s.v.grigoriev@spbu.ru}
\orcid{0000-0002-7966-0698}
\affiliation{%
  \institution{St. Petersburg State University}
  \streetaddress{7-9 Universitetskaya Embankmen}
  \city{St. Petersburg}  
  \country{Russia}
  \postcode{199034}
}



%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Orachev, Grigorev}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
    Scalable high-performance graph analysis is an actual nontrivial challenge. Usage of sparse linear algebra operations as building blocks for graph analysis algorithms, which is a core idea of GraphBLAS standard, is a promising way to attack it.
    While it is known that sparse linear algebra operations can be efficiently implemented on GPGPU, full GraphBLAS implementation on GPGPU is a nontrivial task that is almost solved by GraphBLAST project. Though it is shown that utilization of GPGPUs for GraphBLAS implementation significantly improves performance, portability and scalability problems are not solved yet: GraphBLAST uses Nvidia stack and utilizes only one GPGPU.
    In this work we propose a Spla library that aimed to solve these problems: it uses OpenCL to be portable and designed to utilize multiple GPGPUs.
    Preliminary evaluation shows that while further optimizations are required, the proposed solution demonstrates performance comparable with GraphBLAST on some tasks. Moreover, our solution on embedded GPU outperforms SuiteSparse:GrpaphBLAS on the respective CPU on some graph analysis tasks.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%  
  
\begin{CCSXML}
<ccs2012>
 <concept>
   <concept_id>10002950.10003624.10003633.10010917</concept_id>
   <concept_desc>Mathematics of computing~Graph algorithms</concept_desc>
   <concept_significance>500</concept_significance>
 </concept>
 <concept>
   <concept_id>10010147.10010169.10010170</concept_id>
   <concept_desc>Computing methodologies~Parallel algorithms</concept_desc>
   <concept_significance>500</concept_significance>
 </concept>
 <concept>
   <concept_id>10010520.10010521.10010528.10010534</concept_id>
   <concept_desc>Computer systems organization~Single instruction, multiple data</concept_desc>
   <concept_significance>500</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Mathematics of computing~Graph algorithms}
\ccsdesc[500]{Computing methodologies~Parallel algorithms}
\ccsdesc[500]{Computer systems organization~Single instruction, multiple data}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{graphs, algorithms, graph analysis, sparse linear algebra, GraphBLAS, GPGPU, OpenCL}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle


\input{introduction}
\input{background_of_study}
\input{solution_description}
\input{implementation}
\input{evaluation}
\input{conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{spla_main}


\end{document}
