\section {Coq Listing}
This listing contains main theorems and definitions from our work.


   \begin{pyglist}[language=coq, numbers=none, numbersep=5pt]

 Inductive ter : Type := | T : Tt -> ter.
 Inductive var : Type := | V : Vt -> var.

 Lemma language_normal_form (G:grammar) (A: var) (u: word):
   u <> [] -> (language G A u <-> language (normalize G) A u).

  Inductive symbol : Type :=
     | Ts : ter -> symbol
     | Vs : var -> symbol.

 Definition word := list ter.
 Definition phrase := list symbol.
 Inductive rule : Type := | R : var -> phrase -> rule.
 Definition grammar := list rule.

 Inductive der (G : grammar) (A : var) : phrase -> Prop :=
  | vDer : der G A [Vs A]
  | rDer l : (R A l) el G -> der G A l
  | replN B u w v :
      der G A (u ++ [Vs B] ++ w) ->
      der G B v -> der G A (u ++ v ++ w).

 Definition language (G : grammar) (A : var) (w : phrase) :=
     der G A w /\ terminal w.

 Context {State T: Type}.

Record dfa: Type :=
  mkDfa {
    start: State;
    final: list State;
    next: State -> ter T -> State;
  }.

 Fixpoint final_state (next_d: dfa_rule) (s: State) (w: word): State :=
   match w with
   | nil => s
   | h :: t => final_state next_d (next_d s h) t
   end.

Record s_dfa : Type :=
   s_mkDfa {
     s_start: State;
     s_final: State;
     s_next: State -> (@ter T) -> State;
 }.

 Fixpoint split_dfa_list (st_d : State) (next_d : dfa_rule)
                         (f_list : list State): list (s_dfa) :=
    match f_list with
    | nil => nil
    | h :: t => (s_mkDfa st_d h next_d) :: split_dfa_list st_d next_d t
    end.

 Definition split_dfa (d: dfa) := split_dfa_list (start d) (next d) (final d).

 Lemma correct_split:
    forall dfa w,
      dfa_language dfa w <->
      exists sdfa, In sdfa (split_dfa dfa) /\ s_dfa_language sdfa w.

 Definition syntactic_analysis_is_possible :=
  forall (G : grammar) (A : var) (w : phrase),
  der G A w -> (R A w \in G) \/ (exists rhs, R A rhs \in G /\ derf G rhs w).

Definition convert_nonterm_rule_2 (r r1 r2: _) (state1 state2 : _) :=
   map (fun s3 => R (V (s1, r, s3)) [Vs (V (s1, r1, s2)); Vs (V (s2, r2, s3))])
       list_of_states.

 Definition convert_nonterm_rule_1 (r r1 r2: _) (s1 : _) :=
   flat_map (convert_nonterm_rule_2 r r1 r2 s1) list_of_states.

 Definition convert_nonterm_rule (r r1 r2: _) :=
   flat_map (convert_nonterm_rule_1 r r1 r2) list_of_states.

Definition convert_terminal_rule (next: _) (r: _) (t: _): list TripleRule :=
   map (fun s1 => R (V (s1, r, next s1 t)) [Ts t]) list_of_states.

Definition convert_rule (next: _) (r: _ ) :=
   match r with
   | R r [Vs r1; Vs r2] =>
       convert_nonterm_rule r r1 r2
   | R r [Ts t] =>
       convert_terminal_rule next r t
   | _  => []   (* Never called *)
   end.

 Definition convert_rules
   (rules: list rule) (next: _): list rule :=
   flat_map (convert_rule next) rules.

 (* Maps grammar and s_dfa
    to grammar over triples *)
 Definition convert_grammar grammar s_dfa :=
   convert_rules grammar (s_next s_dfa).

Inductive labeled_Vt : Type :=
  | start : labeled_Vt
  | lV : nat -> Vt -> labeled_Vt.

Definition label_var (label: nat) (v: @var Vt): @var labeled_Vt := V (lV label v).

Variable grammars: seq (var * grammar).

Theorem correct_union:
   forall word,
     language (grammar_union grammars)
     (V (start Vt)) (to_phrase word) <->
     exists s_l, language (snd s_l) (fst s_l) (to_phrase word) /\ In s_l grammars.


   \end{pyglist}
