% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\usepackage{minted}
\usepackage{verbments}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{url}

\begin{document}
%
\title{Bar-Hillel Theorem Mechanization in Coq\thanks{The research was supported by the Russian Science Foundation, grant
  No. 18-11-00100, and by the grant from JetBrains Research.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Sergey Bozhko\inst{1} \and
Leyla Khatbullina\inst{2} \and 
Semyon Grigorev\inst{3,4}\orcidID{0000-0002-7966-0698}}
%
\authorrunning{S. Bozko et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Max Planck Institute for Software Systems (MPI-SWS), Saarbrücken, Germany \\
\email{sbozhko@mpi-sws.com}
\and
St.Petersburg Electrotechnical University ``LETI'', St.Petersburg, Russia\\
\email{leila.xr@gmail.com}
\and
St.Petersburg State University, 7/9 Universitetskaya nab., St.Petersburg, Russia\\
\email{s.v.grigoriev@spbu.ru}
\and
JetBrains Research, Universitetskaya emb., 7-9-11/5A, St.Petersburg, Russia
\email{semen.grigorev@jetbrains.com}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Formal language theory has a deep connection with such areas as static code analysis, graph database querying, formal verification, and compressed data processing.
Many application problems can be formulated in terms of languages intersection.
The Bar-Hillel theorem states that context-free languages are closed under intersection with a regular set.
This theorem has a constructive proof and thus provides a formal justification of correctness of the algorithms for applications mentioned above. 
Mechanization of the Bar-Hillel theorem, therefore, is both a fundamental result of formal language theory and a basis for the certified implementation of the algorithms for applications.
In this work, we present the mechanized proof of the Bar-Hillel theorem in Coq.
We generalize results of Gert Smolka and Jana Hofmann and use them as the base for our work.

\keywords{Formal languages \and Coq \and Bar-Hillel theorem \and Closure \and Intersection \and Regular Language \and Context-free language.}
\end{abstract}
%
%
%
\input{Introduction.tex}
\input{Original_B-H.tex}
%\input{CNF.tex}
\input{Main.tex}
\input{RelatedWork.tex}
\input{Conclusion.tex}

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
 \bibliographystyle{splncs04}
 \bibliography{BH_in_Coq.bib}
%

\end{document}
