%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Trim Size: 9.75in x 6.5in
%% Text Area: 8in (include Runningheads) x 5in
%% ws-ijfcs.tex: 05-05-2015
%% Tex file to use with ws-ijfcs.cls written in Latex2E.
%% The content, structure, format and layout of this style file is the
%% property of World Scientific Publishing Co. Pte. Ltd.
%% Copyright 2015 by World Scientific Publishing Co.
%% All rights are reserved.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%

\documentclass{ws-ijfcs}
\usepackage{enumerate}
\usepackage{textcomp}
\usepackage{url}
\usepackage{graphicx}
\usepackage[numbers]{natbib}
\usepackage{bussproofs}
\usepackage{tikz}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[all,cmtip]{xy}
\usetikzlibrary{positioning, automata}
\usetikzlibrary{decorations.pathmorphing}
 \usetikzlibrary{snakes}


\tikzset{snake it/.style={decorate, decoration=snake}}
\urlstyle{same}
\begin{document}

\markboth{E. Shemetova, A. Okhotin, S. Grigorev}
{The Parallel Complexity of CFL-Reachability Problem:
Tractable Cases
}

%%%%%%%%%%%%%%%%%%%%% Publisher's Area please ignore %%%%%%%%%%%%%%%
%
\catchline{}{}{}{}{}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{The Parallel Complexity of CFL-Reachability Problem:
Tractable Cases\footnote{
This research was supported by the Russian Science Foundation, grant \textnumero 18-11-00100.}}
%using \LaTeX\footnote{For the title, try not to use more than 3
%lines. Typeset the title in 10 pt Roman, boldface with the first letter of
%important words capitalized.}}%

\author{Ekaterina Shemetova\footnote{
St. Petersburg State University, 
7/9 Universitetskaya nab., Saint Petersburg 199034, Russia.}
\footnote{
St. Petersburg Academic University, 
ul. Khlopina, 8, Saint Petersburg 194021, Russia.}
\footnote{
JetBrains Research,
Primorskiy prospekt 68-70, Building 1, St. Petersburg, 197374, Russia.}
}

\address{
\email{katyacyfra@gmail.com}}

\author{Alexander Okhotin\footnotemark[2] }

\address{
\email{alexander.okhotin@spbu.ru}
}

\author{Semyon Grigorev\footnotemark[2] \footnotemark[4] }

\address{
\email{s.v.grigoriev@spbu.ru}
}


\maketitle

\begin{history}
\received{(Day Month Year)}
%\revised{(Day Month Year)}
\accepted{(Day Month Year)}
%\comby{(xxxxxxxxxx)}
\end{history}

\begin{abstract}
Whereas it has been shown that  context-free language (CFL) reachability problem is P-complete, there are some subclasses of context-free languages, for which CFL-reachability lies in NC complexity class. We present two common classes which generalize known examples of such tractable subclasses: bounded-oscillation languages and context-free languages with a poly-slender storage languages. Polynomiality of the rational indices of languages in these classes is proved. Also closure properties of tractable subclasses in terms of polynomial rational index are investigated.
\end{abstract}

\keywords{CFL-reachability; parallel complexity; digraphs; regular languages; context-free languages; context-free path queries.}

\section{Introduction}
\label{intro}
The context-free language (CFL) reachability problem for a context-free grammar $G$ and directed edge-labeled graph $D$ consists of determining for pairs of nodes  $v$ and $u$ whether $v$ can reach $u$ via a path labeled by a string in $L(G)$.  That is, CFL-reachability is a kind of graph reachability problem with path constraints given by context-free languages. It is an important problem underlying some fundamental static code analysis like data flow analysis and program slicing \cite{RepsBasic}, alias analysis \cite*{Chatterjee, alias}, points-to analysis \cite{Incremental} and other \cite{Cai, android, typeflow}, and graph database query evaluation \cite{Azimov, GrigorevRagozina, HellingsCFPQ, RDF}.


Unlike context-free language recognition, which is in NC (when context-free grammar is fixed), CFL-reachability is P-complete \cite{ RepSeq, Yannakakis}. Practically, it means that there is no efficient parallel algorithm for solving this problem (unless P $\neq$ NC). 


While problem is not parallelizable in general, it is useful to develop more efficient parallel solutions for specific subclasses of context-free languages. For example, there are context-free languages which admit more efficient parallel algorithms in comparison with the general case of context-free recognition \cite{IBARRA, IBARRA2, Okhotin2014ComplexityOI}.  The same holds for CFL-reachability problem: there are some examples of context-free languages, for which CFL-reachability problem lies in NL complexity class (for example, linear and one-counter languages) \cite{labelledGraphs, LReach, Regularrealizability}. 


CFL-reachability problem has long been known to be P-complete \cite{PCompl}. A parallel complexity of this problem is studied by both static code analysis \cite{RepSeq, RepsBasic} and database communities \cite{ChainQ, Ullman, Yannakakis}. First investigations of such type were made in terms of Datalog queries, because some classes of Datalog queries (chain queries) can be represented via context-free grammars, while database can be considered as a graph. 
\begin{example}[Datalog chain query as a context-free grammar]
Consider the following Datalog query which determines all pairs of people $x$ and $y$ such that a person $x$ is a descendant of a person $y$: 


$Desc(x, y)$ :- $Child(x, y)$

$Desc(x, y)$ :- $Child(x, z), Desc(z, y)$


This query can be represented as a context-free grammar with the following rules: 


$Desc \rightarrow Child$ $\vert$ $Child$ $Desc$


$Child  \rightarrow child$


Thus, evaluating the above mentioned Datalog query is equivalent to the solving the CFL-reachability problem for a context-free grammar representation of this query and an edge-labeled graph representing database, where every person is represented by a node and edges are labeled with a word ``$child$''. 
\end{example}
Important decidability result is obtained in \cite{Vardi}: given a context-free grammar (query) and an arbitrary graph (database), it is undecidable whether CFL-reachability problem for them is in NC or P-complete. However, Ulman and Van Gelder in \cite{Ullman} introduce a notion of a  \textit{polynomial fringe property} and show that the CFL-reachability problem for context-free grammars having this property and an arbitrary graph is in NC. A context-free grammar $G$ has the \textit{polynomial fringe property} if and only if there is a polynomial $p$ such that, for each regular language $R$ recognized by an automaton with $n$ states, $L(G) \cap R$ is either empty or contains a word shorter than $p(n)$. It is undecidable whether a context-free grammar has the polynomial fringe property. Important results from \cite{Ullman} can be reinterpreted in terms of CFL-reachability as follows: 
\begin{enumerate}
\item CFL-reachability for linear languages and piecewise linear languages, and for arbitrary graphs is in NC, because corresponding grammars have the polynomial fringe property
\item The same holds for $D_1$ (the Dyck language on one kind of parentheses) and its GSM-mappings (one-counter languages)
\item CFL-reachability for $D_2$ (the Dyck language on two kinds of parentheses) is P-complete.
\end{enumerate}
The third result is important because any context-free language can be represented via a regular language and $D_2$, which are combined by means of an intersection and a homomorphism, so it is the direct consequence of P-completeness of CLF-reachability problem in general. Also, using the fact that $D_2$ is included in many interesting subclasses of context-free languages, such as visibly pushdown languages \cite{Okhotin2014ComplexityOI}, simple deterministic languages (defined by LL(1) grammars in Greibach normal form), we can state that CFL-reachability for these languages is P-complete. Afrati et al. \cite{ChainQ} investigate parallel complexity of Datalog simple chain queries and presents the Polynomial Stack Lemma which will be discussed in detail in Section~\ref{sec:poly}. 


The definition of polynomial fringe property coincides with the notion of a so called \textit{rational index}: for a context-free language $L(G)$ having the polynomial rational index is the same as for $G$ to have the polynomial fringe property. More precisely, rational index $\rho_L(n)$ is a function, which denotes the maximum length of the shortest word in $L(G) \cap R$, for arbitrary $R$ recognized by an $n$-state automaton. The notion of rational index was introduced in \cite{RatBasic} as a complexity measure for context-free languages and was investigated independently from the polynomial fringe property.  In particular, it has been proved that the rational index of $D_1$ is in $O( n^2)$ \cite{Dyck1}. 
Another important result concerns the rational index of languages, which generate all context-free languages (an example of such language is $D_2$). It states that the rational index of such languages is of the order $exp(\Theta(n^2/\ln n))$ \cite{CFRat} and, hence, this is the upper bound on the value of rational index for every context-free language. An example of a non-generating language with exponential rational index is given in \cite{Regularrealizability}. Also it has been shown that for every algebraic number $\gamma $ the language with the rational index in $\Theta (n^\gamma )$ exists \cite{GreibRat}. 


The CFL-reachability problem is the same as the intersection non-emptiness problem for a context-free language (pushdown automaton) and a regular language (finite automaton), because a labeled graph is a special kind of a nondeterministic finite automaton. Complexity of this problem is studied by Ganardi et al. \cite{ ganardi2016circuit} ,  Swernofsky et al. \cite{Intersection}, Vyalyi \cite{VyalyiRR}.


Computational complexity of the language reachability for different variants of languages (regular, context-free, context-sensitive) and graphs (acyclic graphs, trees, grid graphs) is discussed in detail by Barret et al. \cite{Barrett}, Holzer et al.\cite{labelledGraphs}, Komarath et al. \cite{LReach}. 


\begin{figure}
\centering
\begin{tikzpicture}
\draw[fill=green, opacity=0.3](0,1) -- (0,0) -- (6,0) --  (6,1);
\draw(0,1) -- (0,0) -- (6,0) --  (6,1);
\draw(0,1) -- (6,1) -- (3, 7) -- (0,1) ;
%\draw[thick, dashed, rounded corners=5mm, fill=yellow, opacity=0.3] (0,1) -- (3.1, 1.5) -- (4.5, 3) -- (6,1);
\draw[thick, dashed, rounded corners=5mm, fill=red, opacity=0.3] (6,1) -- (2.9, 1.5) -- (1.5, 3) -- (0,1);
\draw[thick, densely dotted, rounded corners=5mm, fill=blue, opacity=0.2](0,1) -- (2.5, 5) -- (2.5,4)-- (2.8, 2) -- (6,1);
\draw[thick, densely dotted, rounded corners=5mm , fill=yellow, opacity=0.3](6,1) -- (3.5, 5) -- (3.5,4)-- (3.2, 2) -- (0,1);
\node (reg) at (3, 0.5) {Regular languages}; 
\node (cfl) at (3, 5.5) {CFL}; 
\node (lin) at (1.5, 1.9) {Linear}; 
\node (one) at (4.15, 2.4) {One-counter}; 
\node (linsc) at (2, 3) {Piecewise};
\node (linpsc) at (2.25, 2.7)  {linear}; 
\end{tikzpicture}
\caption{The hierarchy of languages, for which CFL-reachability problem is in NC.}
\label{hierarchy}      
\end{figure}
Our focus is on investigating the parallel complexity of CFL-reachability. Especially we are interested in generalization of ``easy'' subclasses and discovering new examples of context-free languages, for which CFL-reachability is in NC. Effective subclasses can be useful in practice, because the general problem is not tractable \cite{ExperimentalCFPQ}. For example, in case of graph databases it is important to know the complexity of a given context-free path query. Also it is natural to ask which properties of subclasses imply parallel effectiveness. Why some languages have polynomial rational indices? What is the difference between them and other subclasses of context-free languages?


The hierarchy of subfamilies of context-free languages, for which the CFL-reachability problem is in NC, is presented in Figure~\ref{hierarchy}. Linear (and piecewise linear) languages and one-counter languages are incomparable families of context-free languages, but both have polynomial rational indices (polynomial fringe property). These subfamilies have one thing in common: both are defined by strong restrictions on the stack in a pushdown automaton. Our main idea is to generalize known tractable classes by investigating the restrictions on the PDA store.

\textbf{Our contributions.} Our results can be summarized as follows:
\begin{itemize}
\item We show that the CFL-reachability problem for bounded-oscillation languages of Ganty and Valput \cite{BoundOsc}, is in NC (see Section \ref{sec:osc}). This class generalizes the case of linear languages. 
\item Closure properties of the languages with polynomial rational indices are investigated in Section \ref{sec:closure}, particularly it is shown that the family of languages with polynomial rational is closed under Kleene star and insertion of a regular language.
\item In Section \ref{sec:poly} we introduce a new subclass of context-free languages --- context-free languages with a poly-slender pushdown store languages. These languages are the natural generalization of one-counter languages, and the CFL-reachability problem for them is in NC. Also we show that deciding poly-slenderness of a pushdown store language is in P.
\end{itemize}


\input{preliminaries.tex}
\input{boundedosc.tex}
\input{closure.tex}
\input{polystack.tex}



\section{Conclusions and open problems}
\label{sec:conc}
\begin{figure}
\centering
\begin{tikzpicture}
\draw[fill=green, opacity=0.3](0,1) -- (0,0) -- (6,0) --  (6,1);
\draw(0,1) -- (0,0) -- (6,0) --  (6,1);
\draw(0,1) -- (6,1) -- (3, 7) -- (0,1) ;
\draw[thick, dashed, rounded corners=5mm, fill=orange, opacity=0.3] (0,1) -- (3.1, 1.5) -- (4.5, 3) -- (6,1);
\draw[thick, dashed, rounded corners=5mm, fill=magenta, opacity=0.3] (6,1) -- (3.2, 1.7) -- (2.5, 3.4) -- (0,1);
\draw[thick, densely dotted, rounded corners=5mm, fill=blue, opacity=0.2](0,1) -- (2.5, 5) -- (2.5,4)-- (3.1, 2) -- (6,1);
\draw[thick, densely dotted, rounded corners=5mm , fill=yellow, opacity=0.3](6,1) -- (3.5, 5) -- (3.5,4)-- (3.2, 2.1) -- (0,1);
\draw[thick, dashed, rounded corners=5mm, fill=red, opacity=0.3] (0,1) -- (2.3, 2.4)  -- (3.3, 1.1) -- (6,1);
\draw[very thick, densely dotted, rounded corners=5mm] (0,1) -- (2.3, 4.9) -- (3.7, 4.9) -- (6,1);
\node (reg) at (3, 0.5) {Regular languages}; 
\node (cfl) at (3, 5.65) {CFL}; 
\node (pr) at (3, 4.2) {PR};
\node (prrho) at (3, 3.9) {$O(n^\gamma)$};
\node (cflrho) at (3, 5.2) {$exp(\Theta(\frac{n^2}{\ln n}))$}; 
\node (plin) at (2.35, 2.7) {PLin}; 
\node (lin) at (2.1, 1.5) {Lin}; 
\node (osc) at (2.2, 3.7) {BOsc}; 
\node (one) at (4.5, 1.8) {$O(n^3)$}; 
\node (onerho) at (4.5, 2.1) {OC}; 
\node (pssl) at (3.9,3.5) {PSSL}; 
\node (psslrho) at (3.9,3.1) {$O(n^{d})$}; 
\node (linrho) at (2, 1.2) {$O(n^2)$}; 
\node (linrho) at (2.2, 2.29) {$O(n^{2|N|+2})$}; 
\node (oscrho) at (2.1, 3.35) {$O(n^{k})$}; 
\end{tikzpicture}
\caption{The hierarchy of languages with polynomial rational indices and corresponding upper bounds on the value of rational index. PR --- the family of CFLs with a polynomial rational indices, BOsc --- bounded-oscillation languages, PSSL --- CFLs with poly-slender storage languages, PLin --- piecewise linear languages, OC --- one-counter languages, Lin --- linear languages, $n$ --- number of vertices in graph (NFA), $|N|$ --- the number of non-terminals of grammar in Chomsky normal form, $k$ --- the oscillation value, $d$ --- degree of polynomial density of a pushdown storage language, $\gamma$ --- algebraic number.}
\label{hierarchyfinal}      
\end{figure}
We have obtained two classes, which extend the classes in the recent literature \cite{ChainQ, labelledGraphs, LReach, Regularrealizability, Ullman}, for which CFL-reachability problem is in NC. The one is the class of bounded-oscillation languages, which generalizes the linear languages. The second class is context-free languages with a poly-slender pushdown store languages, which is generalization of the one-counter languages. Recall that regular languages have polynomial fringe property (and are accepted by PDA with a bounded stack height), also it is known that L-reachability for regular languages is in NL \cite{LReach, Yannakakis}. Thereby it has been demonstrated that some natural restrictions on the pushdown storage implies polynomial rational index for the corresponding context-free languages: low variability of stack height during the PDA run (bounded-oscillation PDA) and limited number of possible stack contents (languages with poly-slender pushdown store languages). The updated hierarchy of tractable subclasses and corresponding upper bounds on the rational indices are illustrated in Figure~\ref{hierarchyfinal}.


It will be interesting to know whether there is another kind of stack restriction which implies polynomial rational index. Or is there a context-free language which does not belong to any of the above mentioned classes? Are there any other properties (except polynomial rational index) which make the CFL-reachability problem solvable in NC? For example there is a Datalog query, which does not have a polynomial fringe property but its evaluation is in NC \cite{Kanellakis}. One can also approach this question from another direction by looking for simple subfamilies of context-free languages that would have P-complete CFL-reachability problem.


Also it is interesting to consider other machine models with stores having store restrictions examined in this work. Do these restrictions imply that certain properties can become decidable or decidable faster? Store languages of different machine models and their applications are studied in \cite{ Ibarra2018OnSL, IBARRA201928}.


We considered CFL-reachability problem for a fixed context-free languages and arbitrary graphs. What tractable cases can be obtained for a fixed graphs and an arbitrary context-free language? The known and trivial examples are acyclic graphs and trees. Can we have more complicated classes of graphs for which CFL-reachability problem is in NC? Interesting algebraic properties of such graphs (NFA) are given in \cite{ganardi2016circuit}, but automata-theoretic characterizations of these properties remain to be found.


%%For cite command type as \cite{1}; \cite{3,6} and \cite{2,4,6}.
%%For refcite command type as Refs.~[\refcite{1}];
%%[\refcite{1},\refcite{3}] and [\refcite{1}--\refcite{4}].


\section*{Acknowledgments}
This research was supported by the Russian Science Foundation, grant \textnumero 18-11-00100.



\bibliographystyle{ws-ijfcs}  
\bibliography{paper}
\end{document}