\section{The CFPQ\_Data collection}
1 We have collected graphs and grammars for analyzing CFPQ algorithms in one place.
2 We have collected graphs and grammars for analyzing CFPQ algorithms in one place.
3 We have collected graphs and grammars for analyzing CFPQ algorithms in one place.
4 We have collected graphs and grammars for analyzing CFPQ algorithms in one place.

\subsection{Graphs and grammars}
1 We have collected graphs and grammars for analyzing CFPQ algorithms in one place.
2 We have collected graphs and grammars for analyzing CFPQ algorithms in one place.
3 We have collected graphs and grammars for analyzing CFPQ algorithms in one place.
4 We have collected graphs and grammars for analyzing CFPQ algorithms in one place.

\textbf{RDF.}
1 We have collected RDF graphs and grammars for analyzing CFPQ algorithms in one place.
2 We have collected RDF graphs and grammars for analyzing CFPQ algorithms in one place.
3 We have collected RDF graphs and grammars for analyzing CFPQ algorithms in one place.
4 We have collected RDF graphs and grammars for analyzing CFPQ algorithms in one place.

\textbf{MemoryAliases.}
1 We have collected MemoryAliases graphs and grammars for analyzing CFPQ algorithms in one place.
2 We have collected MemoryAliases graphs and grammars for analyzing CFPQ algorithms in one place.
3 We have collected MemoryAliases graphs and grammars for analyzing CFPQ algorithms in one place.
4 We have collected MemoryAliases graphs and grammars for analyzing CFPQ algorithms in one place.

\textbf{LUBM.}
1 We have collected LUBM graphs and grammars for analyzing CFPQ algorithms in one place.
2 We have collected LUBM graphs and grammars for analyzing CFPQ algorithms in one place.
3 We have collected LUBM graphs and grammars for analyzing CFPQ algorithms in one place.
4 We have collected LUBM graphs and grammars for analyzing CFPQ algorithms in one place.

\textbf{Synthetic.}
1 We have collected WorstCase graphs and grammars for analyzing CFPQ algorithms in one place.
2 We have collected SparseGraph graphs and grammars for analyzing CFPQ algorithms in one place.
3 We have collected ScaleFree graphs and grammars for analyzing CFPQ algorithms in one place.
4 We have collected FullGraph graphs and grammars for analyzing CFPQ algorithms in one place.

\subsection{Baselines methods (Loaders)}
1 We have made convenient graph and grammar loaders for analyzing CFPQ algorithms in one place.
2 We have made convenient graph and grammar loaders for analyzing CFPQ algorithms in one place.
3 We have made convenient graph and grammar loaders for analyzing CFPQ algorithms in one place.
4 We have made convenient graph and grammar loaders for analyzing CFPQ algorithms in one place.

\subsection{Evaluation methods (Evaluators)}
1 We have made convenient evaluators for analyzing CFPQ algorithms in one place.
2 We have made convenient evaluators for analyzing CFPQ algorithms in one place.
3 We have made convenient evaluators for analyzing CFPQ algorithms in one place.
4 We have made convenient evaluators for analyzing CFPQ algorithms in one place.
