\section{Construction of Parse Forest Finite Representation}
The number of paths in input automaton graph is exponential and could be infinite,
thus, to be able to use parse forest in practice, some finite representation should 
be constructed. RNGLR-algorithm use Shared Packed Parse Forest (SPPF) presented by 
Rekers~\cite{SPPF} to compress the set of trees of possible derivations of input string. 
We inherit SPPF construction from RNGLR-algorithm, and in our case it merges derivation
trees for strings accumulated along the paths in the input graph. 

SPPF is a directed graph with the following structure: the \emph{root} (i.e. the vertex
with no incoming edges) corresponds to the start nonterminal of the grammar; vertices 
with no outgoing edges correspond to the terminals; the rest of the vertices could be 
divided into 2 classes: nonterminal and production. Each nonterminal node has a 
\emph{family} (a collection) of production nodes, each of which present one of the 
possible derivations of the nonterminal. Production nodes representing righthand side 
of the production have an ordered collection of terminal or nonterminal nodes. The 
lenght of a collection could be equal to the lenght of the rule, or less, if final 
symbols derive $\epsilon$: nullable symbols are ignored which help to reduce memory 
consumption. 

SPPF and GSS are constructed simultaneously. Each edge of GSS is associated with either 
a terminal node or a nonterminal node. When a GSS edge is added while push processing, 
terminal node is created and associated with the edge. Nonterminal nodes are associated
with edges added while reduction processing: if the edge has already been in GSS, 
production node is added to the family of a nonterminal node associated with the edge.
Subgraphs from the edges of the reduction path are added as children to the production
node. After the input is read to the end, vertices with accepting states are searched,
nodes accociated with outgoing edges of such vertices are merged to form the actual
SPPF. All unreachable vertices are deleted from the SPPF graph which leaves only 
correct derivation trees for correct input strings.