% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{amsmath,amsfonts,latexsym,amssymb,euscript,xr}
\usepackage{tikz} 
\usepackage{pgfplots} 
\usepackage{sidecap} 
\usepackage{soul}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{alltt}
\usepackage[misc,geometry]{ifsym}


% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled, linesnumbered, noend]{algorithm2e}
\renewcommand{\algorithmcfname}{Listing}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Modification of Valiant's Parsing Algorithm for the String-Searching Problem\thanks{The research was supported by the Russian Science Foundation, grant No. 18-11-00100.}}
%
\titlerunning{Valiant's Parsing for String-Searching}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Yuliya Susanina\inst{1,2}\orcidID{0000-0003-3904-3764} \Letter \and
Anna Yaveyn\inst{1}\orcidID{0000-0001-9733-5429} \and
Semyon Grigorev\inst{1,2}\orcidID{0000-0002-7966-0698}}
%
\authorrunning{Yuliya Susanina, Anna Yaveyn, and Semyon Grigorev}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Saint Petersburg State University, Universitetskaya nab. 7/9 \\ St. Petersburg, 199034 Russia \\
\and JetBrains Research, Primorskiy prospekt 68-70, Building 1 \\ St. Petersburg 197374, Russia\\
\email{jsusanina@gmail.com} \Letter,
\email{anya.yaveyn@yandex.ru},
\email{semyon.grigorev@jetbrains.com}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Some string-matching problems can be reduced to parsing: verification whether some sequence can be derived in the given grammar. 
To apply parser-based solutions to such area as bioinformatics, one needs to improve parsing techniques so that the processing of large amounts of data was possible.
The most asymptotically efficient parsing algorithm that can be applied to any context-free grammar is a matrix-based algorithm proposed by Valiant.
This paper presents a modification of the Valiant’s algorithm, which facilitates efficient utilization of modern hardware in highly-parallel implementation. 
Moreover, the modified version significantly decreases the number of excessive computations, accelerating the search of substrings.

\keywords{Context-free grammar \and Parsing \and Valiant's algorithm \and String-matching \and Secondary structure.}
\end{abstract}
%
%
%

\input{introduction.tex}
\input{background.tex}
\input{modification.tex}
\input{evaluation.tex}
\input{conclusion.tex}

%
\bibliographystyle{splncs04}
\bibliography{main}

\end{document}
