% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}


% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled, linesnumbered, noend]{algorithm2e}
\renewcommand{\algorithmcfname}{Listing}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

%\setlength{\textfloatsep}{0pt}
%\setlength{\intextsep}{0pt}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\newtheorem{defn}{Definition}

\begin{document}
%
\title{On the Valiant's Algorithm Parallelization\thanks{The research was supported by the Russian Science Foundation, grant No. 18-11-00100.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Yuliya Susanina\inst{1, 2} \and
Anna Yaveyn\inst{1} \and
Semyon Grigorev\inst{1, 2}\orcidID{0000-0002-7966-0698}}
%
\authorrunning{Yuliya Susanina, Anna Yaveyn, and Semyon Grigorev}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Saint Petersburg State University, 7/9 Universitetskaya nab.\\ St. Petersburg, 199034 Russia \\
\and JetBrains Research, Universitetskaya emb., 7-9-11/5A \\ St.Petersburg, Russia \\
\email{jsusanina@gmail.com},
\email{anya.yaveyn@yandex.ru},
\email{semen.grigorev@jetbrains.com}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Recent  research  has  shown  that the theory of formal languages can be used in bioinformatics. While using parsing for processing nucleotide sequences it became necessary to find an easily adaptable to the string-matching problem parsing algorithm, and as such field of application as bioinformatics requires working with large amount of data, it should be highly efficient. The asymptotically fastest parsing algorithm was proposed by Valiant and based on matrix multiplication. The original version of matrix-based algorithm is difficult to apply to the string-matching problem. In this paper we present  a modification  of Valiant's  algorithm dealing with the problem mentioned above. The modification has a succinct proof of correctness and is implemented.

\keywords{Parsing \and Matrix multiplication \and Context-free grammars.}
\end{abstract}
%
%
%
\input{Introduction.tex}
\input{Background.tex}
\input{ModifiedAlgo.tex}
\input{Evaluation.tex}
\input{Conclusion.tex}


%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
\bibitem{chomsky1}
Chomsky, N.: On certain formal properties of grammars. Information and control \textbf{2}(2), 137--167 (1959)

\bibitem{chomsky2}
Chomsky, N. and Schützenberger M. P.: The algebraic theory of context-free languages. In: Studies in Logic and the Foundations of Mathematics vol. 35. pp. 118--161 Elsevier (1963)

\bibitem{rivas}
Rivas, E. and Eddy, S. R: The language of RNA: a formal grammar that includes pseudoknots. Bioinformatics \textbf{16}(4), 334–-340 (2000)

\bibitem{knudsen}
Knudsen, B. and Hein, J.: RNA secondary structure prediction using stochastic context-free grammars and evolutionary history. Bioinformatics (Oxford, England) \textbf{15}(6), 446–-454 (1999)

\bibitem{yuan}
Yuan, C. and Lei, J. and Cole, J. and Sun, Y.: Reconstructing 16S rRNA genes in metagenomic data. Bioinformatics \textbf{31}(12), i35–-i43 (2015)

\bibitem{dowell}
Dowell, R. D. and Eddy, S. R.: Evaluation of several lightweight stochastic context-free grammars for RNA secondary structure prediction. BMC bioinformatics \textbf{5}(1), 71 (2004)

\bibitem{bernardy}
Bernardy, J.P., Claessen K.: Efficient divide-and-conquer parsing of practical context-free languages. In: ACM SIGPLAN Notices, vol. 48. no. 9, pp. 111-122 ACM (2013)

\bibitem{kasami}
Kasami, T.: An efficient recognition and syntax-analysis algorithm for context-free languages. Coordinated Science Laboratory Report no. R-257 (1966)

\bibitem{younger}
Younger, D.H.: Recognition and parsing of context-free languages in time n3. Information and control \textbf{10}(2), 189--208 (1967)

\bibitem{earley}
Earley, J.: An efficient context-free parsing algorithm. Communications of the ACM \textbf{13}(2), 94--102 (1970)

\bibitem{valiant}
Valiant, L. G.: General context-free recognition in less than cubic time. Journal of computer and system sciences \textbf{10}(2), 308--315 (1975)

\bibitem{okhotin}
Okhotin, A.: Parsing by matrix multiplication generalized to Boolean grammars. Theoretical Computer Science \textbf{516}, 101--120 (2014)

\bibitem{m4ri}
The M4RI Library---Version 20121224, \url{http://m4ri.sagemath.org}. Last accessed 5
Jan 2019


\end{thebibliography}
\end{document}
