% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{multirow} 
\usepackage{fancyvrb}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{ON SECONDARY STRUCTURE ANALYSIS BY USING FORMAL GRAMMARS AND ARTIFICIAL NEURAL NETWORKS\thanks{Supported by the Russian Science Foundation grant 18-11-00100}}
%
\titlerunning{Formal Grammars \& Neural Networks}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Polina Lunina\inst{1,2}\orcidID{0000-0002-7172-2647} \and
Semyon Grigorev\inst{1,2}\orcidID{0000-0002-7966-0698}}
%
\authorrunning{Polina Lunina, Semyon Grigorev}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Saint Petersburg State University, 7/9 Universitetskaya nab., St. Petersburg, 199034, Russia \and
JetBrains Research, Primorskiy prospekt 68-70, Building 1, St. Petersburg 197374, Russia\\
\email{lunina\_polina@mail.ru, semyon.grigorev@jetbrains.com, s.v.grigoriev@spbu.ru}\\
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
A way to combine formal grammars and artificial neural networks for biological sequences processing was recently proposed.
In this approach, an ordinary grammar encodes primitive features of the RNA secondary structure, parsing is utilized for features extraction and artificial neural network~--- for processing of the extracted features.
Parsing is a bottleneck of the solution: input sequences should first be parsed before processing with a trained model which is a time-consuming operation when working with huge biological databases.
In this work, we solve this problem by employing staged learning and limiting parsing to be used only during network training.
We also compare networks which represent the parsing result in two different ways: by a vector and a bitmap image.
Finally, we evaluate our solution on tRNA classification tasks.

\keywords{DNN \and CNN \and Machine Learning \and Secondary Structure \and Genomic Sequences \and Formal Grammars \and Parsing.}
\end{abstract}
%
%
%

\input{Introduction}
\input{PreviousWork}
\input{OurSolution}
\input{Evaluation}
\input{Conclusion}

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{main}

\end{document}
