%8 pages
%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf,table]{acmart}

\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{verbatim}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{colortbl}

\usepackage{tikz}
\usetikzlibrary{automata,positioning}

\definecolor{lightgray}{gray}{0.9}


%\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
	%	T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\usepackage{booktabs} % For formal tables


\newlength\Origarrayrulewidth

% horizontal rule equivalent to \cline but with 2pt width
\newcommand{\Cline}[1]{%
	\noalign{\global\setlength\Origarrayrulewidth{\arrayrulewidth}}%
	\noalign{\global\setlength\arrayrulewidth{3pt}}\cline{#1}%
	\noalign{\global\setlength\arrayrulewidth{\Origarrayrulewidth}}%
}

% draw a vertical rule of width 2pt on both sides of a cell
\newcommand\Thickvrule[1]{%
	\multicolumn{1}{!{\vrule width 2pt}c!{\vrule width 2pt}}{#1}%
}

% draw a vertical rule of width 2pt on the left side of a cell
\newcommand\Thickvrulel[1]{%
	\multicolumn{1}{!{\vrule width 2pt}c}{#1}%
}

% draw a vertical rule of width 2pt on the right side of a cell
\newcommand\Thickvruler[1]{%
	\multicolumn{1}{c!{\vrule width 2pt}}{#1}%
}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
	\providecommand\BibTeX{{%
			\normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2020}
\acmYear{2020}
\acmConference[GRADES-NDA'20]{Joint Workshop on Graph Data Management Experiences \& Systems (GRADES) and Network Data Analytics (NDA)}{June 14, 2020}{Portland, OR, USA}
\acmBooktitle{Joint Workshop on Graph Data Management Experiences \& Systems (GRADES) and Network Data Analytics (NDA), June 14, 2020, Portland, OR, USA}
\acmPrice{15.00}
\acmDOI{}
\acmISBN{XXX-X-XXXXX-XXX-X}


%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

\newtheorem{mytheorem}{Theorem}

\newcommand{\ltz}{$< 1$}

\algtext*{EndWhile}% Remove "end while" text
\algtext*{EndIf}% Remove "end if" text
\algtext*{EndFor}% Remove "end for" text
\algtext*{EndFunction}% Remove "end function" text

%%
%% end of the preamble, start of the body of the document source.
\begin{document}
	
	%%
	%% The "title" command has an optional parameter,
	%% allowing the author to define a "short title" to be used in page headers.
	\title[]{Context-Free Path Querying with Single-Path Semantics by Matrix Multiplication}
	
	%%
	%% The "author" command and its associated commands are used to define
	%% the authors and their affiliations.
	%% Of note is the shared affiliation of the first two authors, and the
	%% "authornote" and "authornotemark" commands
	%% used to denote shared contribution to the research.
	\author{Arseniy Terekhov}
	\email{simpletondl@yandex.ru}
	\affiliation{%
		\institution{Saint Petersburg State University}
		\streetaddress{7/9 Universitetskaya nab.}
		\city{St. Petersburg}
		\country{Russia}
		\postcode{199034}
	}
	
	
	\author{Artyom Khoroshev}
	\email{arthoroshev@gmail.com}
	\affiliation{%
		\institution{ITMO University}
		\streetaddress{49 Kronverksky prospekt}
		\city{St. Petersburg}
		\country{Russia}
		\postcode{197101}
	}
	
	
	\author{Rustam Azimov}
	\email{rustam.azimov19021995@gmail.com}
	%\email{semyon.grigorev@jetbrains.com}
	%orcid{0000-0002-7966-0698}
	\affiliation{
		\institution{Saint Petersburg State University}
		\streetaddress{7/9 Universitetskaya nab.}
		\city{St. Petersburg}
		\country{Russia}
		\postcode{199034}
	}
	\affiliation{
		\institution{JetBrains Research}
		\streetaddress{Primorskiy prospekt 68-70, Building 1}
		\city{St. Petersburg}
		\country{Russia}
		\postcode{199034}
	}
	
	
	\author{Semyon Grigorev}
	\email{s.v.grigoriev@spbu.ru}
	\email{semyon.grigorev@jetbrains.com}
	\orcid{0000-0002-7966-0698}
	\affiliation{
		\institution{Saint Petersburg State University}
		\streetaddress{7/9 Universitetskaya nab.}
		\city{St. Petersburg}
		\country{Russia}
		\postcode{199034}
	}
	\affiliation{
		\institution{JetBrains Research}
		\streetaddress{Primorskiy prospekt 68-70, Building 1}
		\city{St. Petersburg}
		\country{Russia}
		\postcode{199034}
	}
	
	%%
	%% By default, the full list of authors will be used in the page
	%% headers. Often, this list is too long, and will overlap
	%% other information printed in the page headers. This command allows
	%% the author to define a more concise list
	%% of authors' names for this purpose.
	\renewcommand{\shortauthors}{Terekhov, Khoroshev, Azimov, Grigorev}
	
	%
	% The abstract is a short summary of the work to be presented in the article.
	\begin{abstract}
		A recent study showed that the applicability of context-free path querying (CFPQ) algorithms with relational query semantics integrated with Neo4j database is limited because of low performance and high memory consumption. In this work we implement a matrix-based CFPQ algorithm by using appropriate high-performance libraries for linear algebra and integrate it with RedisGraph graph database. Also, we introduce a new CFPQ algorithm with single-path query semantics that allows us to extract one found path for each pair of nodes. Finally, we provide the evaluation of our algorithms for both semantics which shows that matrix-based CFPQ implementation for RedisGraph database is performant enough for real-world data analysis.
	\end{abstract}
	
	%
	% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
	% Please copy and paste the code instead of the example below.
	%
	\begin{CCSXML}
		<ccs2012>
		<concept>
		<concept_id>10002951.10002952.10003197.10010825</concept_id>
		<concept_desc>Information systems~Query languages for non-relational engines</concept_desc>
		<concept_significance>500</concept_significance>
		</concept>
		<concept>
		<concept_id>10003752.10003766.10003771</concept_id>
		<concept_desc>Theory of computation~Grammars and context-free languages</concept_desc>
		<concept_significance>500</concept_significance>
		</concept>
		<concept>
		<concept_id>10010147.10010169.10010170.10010174</concept_id>
		<concept_desc>Computing methodologies~Massively parallel algorithms</concept_desc>
		<concept_significance>500</concept_significance>
		</concept>
		<concept>
		<concept_id>10003752.10003753.10003761.10003762</concept_id>
		<concept_desc>Theory of computation~Parallel computing models</concept_desc>
		<concept_significance>300</concept_significance>
		</concept>
		<concept>
		<concept_id>10010520.10010521.10010528.10010534</concept_id>
		<concept_desc>Computer systems organization~Single instruction, multiple data</concept_desc>
		<concept_significance>300</concept_significance>
		</concept>
		</ccs2012>
	\end{CCSXML}
	
	\ccsdesc[500]{Information systems~Query languages for non-relational engines}
	\ccsdesc[500]{Theory of computation~Grammars and context-free languages}
	\ccsdesc[500]{Computing methodologies~Massively parallel algorithms}
	\ccsdesc[300]{Theory of computation~Parallel computing models}
	\ccsdesc[300]{Computer systems organization~Single instruction, multiple data}
	%
	% Keywords. The author(s) should pick words that accurately describe the work being
	% presented. Separate the keywords with commas.
	\keywords{Context-free path querying, transitive closure, graph databases, RedisGraph database, linear algebra, context-free grammar, GPGPU, CUDA, Boolean matrix, matrix multiplication}
	
	%%
	%% This command processes the author and affiliation and title
	%% information and builds the first part of the formatted document.
	\maketitle

\input{introduction}
\input{motivating}
\input{preliminaries}
\input{matrix_based_algorithm}
\input{cfpq_single_path}
\input{implementation}
\input{dataset}
\input{evaluation}
\input{conclusion}

\begin{acks}
The reported study was funded by RFBR, project number 19-37-90101.
\end{acks}

%
% The next two lines define the bibliography style to be used, and the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{cfpq_for_redisgraph}

%
% If your work has an appendix, this is the place to put it.
%\vspace{10cm}

%\appendix

%\input{appendix}

%\subsection{Part One}


\end{document}
